---
title: "Step 1 Prepare data to make count history matrix"
author: "Zachary Amir"
date: "`r Sys.time()`"
output: html_document
---

## Introduction

This code will prepare camera trap data to create count history matrices and the associated covariates. This is a necessary first step towards preparing the data to implement Zachary Amir's co-abundance models on the High Performance Computers (HPC).

This code works off the spatially re-sampled captures and covariates that are generated in the 4-step Southeast Asian camera trap data standardization pipeline. To learn more about this Southeast Asian camera trap data standardization pipeline, please contact Zachary Amir (z.amir@uq.edu.au) or Matthew Luskin (m.luskin@uq.edu.au) to request access to the following relevant [GitHub repository](https://github.com/EcologicalCascadesLab/AsianCaptureHistories).

```{r global-options, include=FALSE, warning=FALSE, error=FALSE}
## We can knit this into a nice document! but dont include all the code unless otherwise specified (i.e. include=TRUE)
knitr::opts_chunk$set(include=FALSE, warning=FALSE, error = FALSE)

## start fresh
rm(list = ls())

## load libraries
library(tidyverse)  # For basic data wrangling
library(plyr)       # For data summaries
library(traitdata)  # For species trait data (e.g. body mass, trophic guild, home ranges)

```

```{r import captures}

#### Re-sampled captures and metadata are saved per survey in a folder in the cap hist DB on DropBox
### This raw data is not available for public usage, but a derived version will be hosted on DropBox. 

### captures first 

## specify where the data lives
wd = "/Users/zachary_amir/Dropbox/CT capture histories database/Asian ECL raw CT data/Step5_HPC_import_resampled_captures/"

## import and combine here
files = list.files(wd)
files = files[!grepl("OLD", files)] # remove old data folder

## save the scale we are interested in 
scale = "5km" #

## thin to the relevant scale were interested in 
files = files[grepl(scale, files)] 

## store caps here-
caps = list()

for(i in 1:length(files)){
  
  # select a file 
  f = files[i]
  
  # make the path 
  f_path = paste(wd, f, sep = "")
  
  # read it
  d = read.csv(f_path)
  
  ## rename cell_id so it is scale agnostic
  names(d)[grepl("cell_id", names(d))] = "cell_id"
  
  ## extract ID tag from file name
  id = strsplit(f, "_spatially_resampled_captures")[[1]][1]
  
  # remove the number
  id = gsub("^\\d+_", "", id)
  
  ## save it! 
  caps[[i]] = d
  names(caps)[i] = id
  
  ## let us know if there are any duplicate files in our repo
  if(any(duplicated(names(caps)))){
    print(paste("There are repeated files from the same survey with the tag:", id,
                "Make sure to remove it before importing!"))
  }# end conditional
  
} # end per file 
rm(f,f_path, d, id, i, files, wd)

## combine list into a df
caps = do.call(rbind, caps)
rownames(caps) = NULL


```

```{r import metdata}

### Same deal as the captures, this data lives in a separate dropbox directory not linked to R Project, so we need to specify that directly specifically. 

## specify where the data lives
wd = "/Users/zachary_amir/Dropbox/CT capture histories database/Asian ECL raw CT data/Step5_HPC_import_resampled_metadata/"

## import and combine here
files = list.files(wd)
files = files[!grepl("OLD", files)] # remove old data folder

## thin to the same scale as above
files = files[grepl(scale, files)]

## store caps here-
meta = list()

for(i in 1:length(files)){
  
  # select a file 
  f = files[i]
  
  # make the path 
  f_path = paste(wd, f, sep = "")
  
  # read it
  d = read.csv(f_path)
  
  ## rename cell_id so it is scale agnostic
  names(d)[grepl("cell_id", names(d))] = "cell_id"
  
  ## extract ID tag from file name
  id = strsplit(f, "_spatially_resampled_metadata")[[1]][1]
  
  # remove the number
  id = gsub("^\\d+_", "", id)
  
  ## save it! 
  meta[[i]] = d
  names(meta)[i] = id
  
  ## let us know if there are any duplicate files in our repo
  if(any(duplicated(names(meta)))){
    print(paste("There are repeated files from the same survey with the tag:", id,
                "Make sure to remove it before importing!"))
  }# end conditional
  
} # end per file 
rm(f,f_path, d, id, i, files, wd)

## combine list into a df
meta = do.call(rbind, meta)
rownames(meta) = NULL

```

```{r Inspect and prepare data}

## ensure all cell_ids match!
setdiff(meta$cell_id, caps$cell_id)
setdiff(caps$cell_id, meta$cell_id) # full match, good to go! 

## format classes
caps$Date = as.Date(caps$Date, format = "%Y-%m-%d")
meta$Sampling_begin = as.Date(meta$Sampling_begin, format = "%Y-%m-%d")
meta$Sampling_end = as.Date(meta$Sampling_end, format = "%Y-%m-%d")

## need survey year in covariates 
meta$year = str_extract(meta$survey_id, stringr::regex("(\\d+)(?!.*\\d)"))
sort(table(meta$year)) #looks good! 


#### Remove data that does not meet inclusion criteria

## Remove Brodie's data W of wallace line --> doesnt have relevant species
meta = meta[! grepl("E_Indonesia", meta$Landscape),]

# and thin caps to match
caps = caps[caps$survey_id %in% meta$survey_id,]


#### To be conservative, lets thin the data surveys with enough data

## summarize info about surveys
surv_summary = ddply(caps, .(survey_id), summarize,
                     num_SU = length(unique(cell_id)), # total number of sampling units (SUs)
                     max_cam = max(num_cams_active_at_date), # maximum number of cams active in a single sampling unit
                     dur = difftime(max(Date), min(Date), units= "days")) # duration of the survey 
## calculate trap nights 
surv_summary$trap_nights = surv_summary$num_SU * surv_summary$dur

## Extract surveys that do not meet inclusion criteria. 
rm = surv_summary$survey_id[as.numeric(surv_summary$trap_nights) < 100] # less than 100 trap nights 
rm2 = surv_summary$survey_id[(surv_summary$num_SU) < 3] # less than 3 sampling units in a survey 
rm = unique(c(rm,rm2))

## how much do we lose when we remove this data?
dat_rm = round(dim(caps[caps$survey_id %in% rm,])[1]  / dim(caps)[1] * 100, 3)

## remove the data from caps
caps = caps[!caps$survey_id %in% rm,]

## and metadata
meta = meta[meta$survey_id %in% caps$survey_id,]

## display survey information below


```

```{r Import non-resampled data for referencing}

### this will be important for determining the number of deployments used total and per survey 

## specify the workind diretory where og data lives
og_wd = "/Users/zachary_amir/Dropbox/CT capture histories database/Asian ECL raw CT data/Step4_output_pre-resampling/"

# list all relevant files 
files = list.files(og_wd, recursive = T)
# remove TEAM and OLD data 
files = files[!grepl("TEAM", files)]
files = files[!grepl("OLD", files)]

## import captures
og_caps = read.csv(paste(og_wd, files[grepl("cap", files)], sep = ""))
## import metadata
og_meta = read.csv(paste(og_wd, files[grepl("meta", files)], sep = ""))

## verify surveys match
og_meta = og_meta[og_meta$survey_id %in% meta$survey_id, ]
og_caps = og_caps[og_caps$survey_id %in% caps$survey_id, ]

## clean up! 
rm(files, og_wd)


## Also need SU and num polygons per landscape
land_sum = ddply(meta, .(Landscape), summarize,
                 num_su = (length(unique(cell_id))),
                 num_polygon = length(unique(Polygon5km)))



```


## Camera trap data included for analysis  

The camera trap data used in this analysis comes from **`r length(unique(meta$source))` data providers** that sampled **`r length(unique(meta$Landscape))` spatially distinct landscapes** from a total of **`r length(unique(meta$survey_id))` temporally distinct survey_ids** that have been defined to be a maximum of 100 days long. This data is comprised of **`r length(unique(og_meta$deployment_id))` single-season camera deployments**. The single-season surveys ranged in duration from `r as.numeric(min(surv_summary$dur[!surv_summary$survey_id %in% rm]))` to `r as.numeric(max(surv_summary$dur[!surv_summary$survey_id %in% rm]))` days, with an average of `r round(as.numeric(mean(surv_summary$dur[!surv_summary$survey_id %in% rm])),2)` and SD of `r round(as.numeric(sd(surv_summary$dur[!surv_summary$survey_id %in% rm])),2)` days. In terms of trap-nights (*i.e.*, number of deployments multiplied by duration), the average sampling effort per survey had `r as.numeric(min(surv_summary$trap_nights[!surv_summary$survey_id %in% rm]))` to `r as.numeric(max(surv_summary$trap_nights[!surv_summary$survey_id %in% rm]))` trap-nights, with an average of `r round(as.numeric(mean(surv_summary$trap_nights[!surv_summary$survey_id %in% rm])),2)` and SD of `r round(as.numeric(sd(surv_summary$trap_nights[!surv_summary$survey_id %in% rm])),2)` trap-nights. 

We have spatially re-sampled the camera trap captures and covariates to the **`r scale` spatial scale**, meaning that any cameras within a `r scale` hexagon had their captures aggregated. There are a total of **`r paste(length(unique(meta$cell_id)), scale, "sampling units")`** available for analysis, and each **landscape contains between `r min(land_sum$num_su)` to `r max(land_sum$num_su)`** sampling units, with an average of `r round(mean(land_sum$num_su),2)` and SD of `r round(sd(land_sum$num_su),2)` per landscape. But remember, sampling units are temporal, so the same location can be repeated multiple times thru time. Therefore, to assess sizes, its better to inspect the number of `r scale` polygons sampled. **Each landscapes contains between `r min(land_sum$num_polygon)` to `r max(land_sum$num_polygon)` polygons**, with an average of `r round(mean(land_sum$num_polygon), 2)` and SD of `r round(sd(land_sum$num_polygon),2)` per landscape.

As part of ensuring we are using robust data, a landscape needs be at least 9 km2 to provide spatial variation and a survey needs to contain at least 100 trap-nights. Therefore, a total of __`r length(rm)` surveys have been removed__ that do not meet these standards, and this represents __`r paste(round(dat_rm,1), "%", sep = "")` of the captures__. For the remaining `r length(unique(meta$survey_id))` valid surveys, the __median number of `r scale` sampling units is `r median(surv_summary$num_SU[!surv_summary$survey_id %in% rm])` with a range from `r paste(min(surv_summary$num_SU[!surv_summary$survey_id %in% rm]), max(surv_summary$num_SU[!surv_summary$survey_id %in% rm]), sep = " - ")`__, and the __median sampling effort per survey is `r as.numeric(median(surv_summary$trap_nights[!surv_summary$survey_id %in% rm]))` trap-nights, with a range from `r paste(as.numeric(min(surv_summary$trap_nights[!surv_summary$survey_id %in% rm])), as.numeric(max(surv_summary$trap_nights[!surv_summary$survey_id %in% rm])), sep = " - ")`__.

## Determine which species to analyze

In these code chunks, I will determine which species have **sufficient captures (n = 100)**, standardize similar species to the genera level (*e.g.*, clouded leopards in the genus *Neofelis*), and remove non-relevant 'species' (e.g. blanks). 

I will also gather species traits from the `library(speciestraits)` databases. Each species will have 1) a trophic level (*i.e.*, carnivore, omnivore, herbivore), 2) body mass (in grams), and 3) home range size (in km2). Finally, I have also added dietary preferences for the four large carnivores species that are the focus of this analysis. 

```{r select and standardize species to include}

## clean up environment from previous reporting text
rm(dat_rm, rm, rm2, surv_summary, land_sum)

##### Which species will be included and how to standardize species?

## replace space with underscore to facilitate clean code/data
caps$Species = gsub(" ", "_", caps$Species)

# which species have sufficent (n=100) detections?
sp_count = ddply(caps, .(Species), summarize,
                 num_cap = length(Species))
keep = sp_count[sp_count$num_cap >= 100,]

## what did we get?
keep[order(keep$num_cap),] ## Awesome!! 
full_length = length(keep$Species) # 106 species before any cleaning 


## Standardize muntjacs, tragulus, Hystrix (?), Neofelis, Capricornis, Tupaia, maybe Canis_lupus
## remove from analysis: Ghost Homo_sapiens, unID'd, remove, Aves, bird, unkown, bird_pheasent, etc

### Clean up and combine some speices names

# Hystrix
sort(table(caps$Species[grepl("Hystrix", caps$Species)])) ## dont change crassispinis or brachyura b/c we have enough! 
# caps$Species[caps$Species %in% c("Hystrix_brachyura", "Hystrix")] = "Hystrix_genus"

# Muntjac
sort(table(caps$Species[grepl("Munt", caps$Species)])) # ecologically very similar species, but what about spatial distribution?
## where do all these muntjacs come from?
Ma = caps$cell_id[caps$Species == "Muntiacus_atherodes"]
unique(meta$Landscape[meta$cell_id %in% Ma]) # Sarawak and Sabah.
Mr = caps$cell_id[caps$Species == "Muntiacus_reevesi"]
unique(meta$Landscape[meta$cell_id %in% Mr]) # Sarawak, Sabah, Sumatra, and S. Thailand...
Mv = caps$cell_id[caps$Species == "Muntiacus_vaginalis"]
unique(meta$Landscape[meta$cell_id %in% Mv]) # Thailand, Vietnam, China
Mm = caps$cell_id[caps$Species == "Muntiacus_muntjak"]
sort(unique(meta$Landscape[meta$cell_id %in% Mm])) # Laos, P_Malaysia, Sarawak, Sabah, Sumatra, Thailand (many)
M = caps$cell_id[caps$Species == "Muntiacus"]
sort(unique(meta$Landscape[meta$cell_id %in% M])) # Laos, Sarawak, Thailand (many), Vietnam.
## These are all quite spatially distinct, so there is good motivation to combine these species into one
# Make sure to save independent detections of all for supplementary materials later! 
rm(Ma, Mr, Mv, Mm, M)

## combine
caps$Species[grepl("Munt", caps$Species)] = "Muntiacus_genus"

# clouded leopards
sort(table(caps$Species[grepl("Neof", caps$Species)])) #just geographical differences, but way more bornean than mainland! 
caps$Species[grepl("Neof", caps$Species)] = "Neofelis_genus"

# mousedeer
sort(table(caps$Species[grepl("Trag", caps$Species)])) #Amazed that these are even differentiated 
caps$Species[grepl("Trag", caps$Species)] = "Tragulus_genus"

# serow
sort(table(caps$Species[grepl("Capric", caps$Species)])) #I think these were recently combined by taxonomists anyway 
caps$Species[grepl("Capric", caps$Species)] = "Capricornis_genus"

# treeshrew
sort(table(caps$Species[grepl("Tupai", caps$Species)])) #Amazed that these are even differentiated 
caps$Species[grepl("Tupai", caps$Species)] = "Tupaia_genus"

#dogs
sort(table(caps$Species[grepl("Canis_lup", caps$Species)])) #Just call these dogs, we dont have pics to know if domestic or not 
caps$Species[grepl("Canis_lup", caps$Species)] = "Canis_lupus_familiaris"

# mongoose
sort(table(caps$Species[grepl("Herpe", caps$Species)])) # Just stick with H_urva and H_brachyrus 
caps$Species[caps$Species %in% c("Herpestes", "Herpestes_semitorquatus",
                                 "Herpestidae")] = "Herpestes_genus" # this will get excluded later anyway b/c we have more specific species, which is better. 

# Murids
unique(caps$Species[grepl("Muri", caps$Species)]) #only the family muridae
caps$Species[grepl("Muri", caps$Species)] = "Muridae_spp" #spp denotes family here. 

#squirells 
unique(caps$Species[grepl("Sciur", caps$Species)]) #only the family sciuridae
caps$Species[grepl("Sciur", caps$Species)] = "Sciuridae_spp"


#Combine northern and southern pig tail macaques into one species Macaca_leonina Macaca_nemestrina 
sort(table(caps$Species[grepl("Macaca_", caps$Species)]))
caps$Species[caps$Species %in% c("Macaca_leonina", "Macaca_nemestrina" )] = "Macaca_nemestrina"

#chickens
unique(caps$Species[grepl("Gallu", caps$Species)]) #analyze at genus level. 
caps$Species[startsWith(caps$Species, "Gallus")] = "Gallus_genus"

## attach genus to the other Genera names that need it to stay consistent
caps$Species[caps$Species %in% c("Amaurornis", "Macaca", "Hystrix",
                                 "Maxomys","Melogale", "Varanus")] = paste(caps$Species[caps$Species %in% c("Amaurornis", "Macaca",
                                                                                                            "Hystrix","Maxomys","Melogale", "Varanus")], "genus", sep = "_")

### Make a new species count w/ updated species names
sp_count = ddply(caps, .(Species), summarize,
                 num_cap = length(Species))
keep = sp_count[sp_count$num_cap >= 100,]

## what did we get?
sort(keep$Species) #reduced to 91 species

# remove some of the not useful fluff
keep = keep[!keep$Species %in% c("Aves", "Ghost", "Mammalia",
                                 "Unknown","Vehicle"),] 

#remove all  people 
keep = keep[!grepl("Homo", keep$Species ),] 

# remove species not ID to the species or genus level 
keep = keep[!endsWith(keep$Species, "_spp"),]
# and the others. 
keep = keep[! keep$Species %in% c("Scincidae","Small_mammal", "Herpestidae",
                                  "Phasianidae", "Rodentia","Scandentia"),]

length(keep$Species) # 73 species after removing fluff. 
rm(sp_count)

### Looks good, but will remove more species once body weights are accounted for. 
sort(keep$Species)

## save it as a vector, not dataframe
keep = keep$Species

```

```{r initalize species trait data, include=FALSE}

## access panthera dataset
pan = pantheria
# replace space with underscore for species names to match our style
pan$scientificNameStd = gsub(" ", "_", pan$scientificNameStd)

## access mammal_diet dataset
mam = mammal_diet2
mam = as.data.frame(mam) # remove silly tibble
# replace space with underscore for species names to match our style
mam$scientificNameStd = gsub(" ", "_", mam$scientificNameStd)
## remove NA species names
mam = mam[!is.na(mam$scientificNameStd),]

## import Sam Lee's cleaned avonet dataframe from dropbox 
wd = "/Users/zachary_amir/Dropbox/Zach PhD/Ch3 Trophic release project/SEA_TC_GitHub_data_storage/data/"

bird = read.csv(paste(wd, "bird_guild_information_20221102.csv", sep = ""))

# replace space with underscore for bird species names
bird$Species = gsub(" ", "_", bird$Species)

## check which species we have matches for 
intersect(bird$Species, keep) # 12 matches and all birds, great! 
# thin bird to relevant species
bird = bird[bird$Species %in% c(keep, "Gallus_gallus"),]
bird$Species[bird$Species == "Gallus_gallus"] = "Gallus_genus" #std to match

## copy "Amaurornis_phoenicurus" and provide for genus Amaurornis catch-all
am = bird[bird$Species == "Amaurornis_phoenicurus",]
am$Species = "Amaurornis_genus" 
bird = rbind(bird,am)
rm(am)

## access a dataset for varanus salvator 
rep = lizard_traits
# thin to our singular herp species
rep = rep[rep$scientificNameStd == "Varanus salvator",] # good! 
rep = rep[!is.na(rep$scientificNameStd),] # idk why, but it comes with lots of NA
# make sure species name is consistent!
rep$scientificNameStd = gsub(" ", "_", rep$scientificNameStd)
# thin to relevant info
rep = select(rep, scientificNameStd, diet) # missing body wieght and home range 

## adding custom body weight as determined by Shine et al 1996, "Commercial harvesting of giant lizards: The biology of water monitors Varanus salvator in southern Sumatra"
rep$Body_mass = 3.24
names(rep)[2] = "TrophicLevel"

## add a second row of rep for Varanus catch-all genus
rep = rbind(rep,rep)
rep$scientificNameStd[2] = "Varanus_genus"

## rbind w/ our birds
names(bird)[1:2] = c("scientificNameStd","TrophicLevel")
rep_bird = rbind(rep, bird)

# convert kg masses to grams to match mammal format 
rep_bird$Body_mass = rep_bird$Body_mass * 1000
names(rep_bird)[3] = "AdultBodyMass_g"

## dont know home range for these guys, so just leave as NA
rep_bird$HomeRange_km2 = NA

## thin mam and pan to relevant info while incorporating known typo
mam = select(mam[mam$scientificNameStd %in% c(keep, "Taperus_indicus", 
                                              "Canis_lupus"),], scientificNameStd, TrophicLevel) # diet from here
pan = select(pan[pan$scientificNameStd %in% c(keep, "Taperus_indicus", 
                                              "Canis_lupus"),], scientificNameStd, AdultBodyMass_g, HomeRange_km2) # mass and home range from here

# and merge together
mam_pan = distinct(merge(mam, pan, by = "scientificNameStd"))

# There are two duplicated rows, remove them! 
a = mam_pan[mam_pan$scientificNameStd == "Callosciurus_notatus" & !is.na(mam_pan$AdultBodyMass_g),] # remove dup w/out mass
b = mam_pan[mam_pan$scientificNameStd == "Atherurus_macrourus" & 
              mam_pan$TrophicLevel == "Herbivore",] # remove dup w/ omnivore tag for porcupine
c = rbind(a,b)
## remove bad species
mam_pan = mam_pan[!mam_pan$scientificNameStd %in% c("Atherurus_macrourus", "Callosciurus_notatus"),]
## and rbind good records of those species
mam_pan = rbind(mam_pan, c)
rm(a,b,c)


# Now combine them all!
traits = rbind(mam_pan, rep_bird)

## check which species are missing 
setdiff(keep, traits$scientificNameStd) # 12, though mostly genera or a few typos 

## fix obvious typos first
traits$scientificNameStd[traits$scientificNameStd == "Taperus_indicus"] = "Tapirus_indicus"
traits$scientificNameStd[traits$scientificNameStd == "Canis_lupus"] = "Canis_lupus_familiaris"

## add 10 missing species to traits object
traits[64:(64+9), "scientificNameStd"] = setdiff(keep, traits$scientificNameStd)

### re-load mam and pam to include other species
# pan
pan = pantheria
pan$scientificNameStd = gsub(" ", "_", pan$scientificNameStd)
pan = select(pan, scientificNameStd, AdultBodyMass_g, HomeRange_km2)
# mam
mam = mammal_diet2
mam = as.data.frame(mam) # remove silly tibble
mam$scientificNameStd = gsub(" ", "_", mam$scientificNameStd)
mam = mam[!is.na(mam$scientificNameStd),]
mam = select(mam, scientificNameStd, TrophicLevel)


## add info per species/guild
#serow
traits[traits$scientificNameStd == "Capricornis_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Capricornis_sumatraensis"]
traits[traits$scientificNameStd == "Capricornis_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Capricornis_sumatraensis", c("AdultBodyMass_g", "HomeRange_km2")]

# mongooses 
traits[traits$scientificNameStd == "Herpestes_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Herpestes_urva"]
traits[traits$scientificNameStd == "Herpestes_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Herpestes_urva", c("AdultBodyMass_g", "HomeRange_km2")]

#muntjacs
traits[traits$scientificNameStd == "Muntiacus_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Muntiacus_muntjak"])
traits[traits$scientificNameStd == "Muntiacus_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Muntiacus_muntjak", c("AdultBodyMass_g", "HomeRange_km2")]

#clouded leopards
traits[traits$scientificNameStd == "Neofelis_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Neofelis_nebulosa"]
traits[traits$scientificNameStd == "Neofelis_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Neofelis_nebulosa", c("AdultBodyMass_g", "HomeRange_km2")]

#mousedeer
traits[traits$scientificNameStd == "Tragulus_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Tragulus_napu"]
traits[traits$scientificNameStd == "Tragulus_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Tragulus_napu", c("AdultBodyMass_g", "HomeRange_km2")]

#gaur
traits[traits$scientificNameStd == "Bos_gaurus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Bos_javanicus"]
traits[traits$scientificNameStd == "Bos_gaurus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Bos_javanicus", c("AdultBodyMass_g", "HomeRange_km2")]

#porcupines
traits[traits$scientificNameStd == "Hystrix_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Hystrix_brachyura"])
traits[traits$scientificNameStd == "Hystrix_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Hystrix_brachyura", c("AdultBodyMass_g", "HomeRange_km2")]

#maxomys
traits[traits$scientificNameStd == "Maxomys_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Maxomys_rajah"] # widespread in Sundaland
traits[traits$scientificNameStd == "Maxomys_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Maxomys_rajah", c("AdultBodyMass_g", "HomeRange_km2")]

#Tupaia_genus
traits[traits$scientificNameStd == "Tupaia_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Tupaia_glis"])
traits[traits$scientificNameStd == "Tupaia_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Tupaia_glis", c("AdultBodyMass_g", "HomeRange_km2")]

#Macaca
traits[traits$scientificNameStd == "Macaca_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Macaca_nemestrina"]) # maybe silly b/c already present?
traits[traits$scientificNameStd == "Macaca_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Macaca_nemestrina", c("AdultBodyMass_g", "HomeRange_km2")]

#Melogale
traits[traits$scientificNameStd == "Melogale_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Melogale_moschata"]) 
traits[traits$scientificNameStd == "Melogale_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Melogale_moschata", c("AdultBodyMass_g", "HomeRange_km2")]

### two civets are missing weights!! 
# Hose civet
traits$AdultBodyMass_g[traits$scientificNameStd == "Diplogale_hosei"] = 2500 #https://www.ecologyasia.com/verts/mammals/hose's-civet.htm
# large spotted civet
traits$AdultBodyMass_g[traits$scientificNameStd == "Viverra_megaspila"] = 6600 #https://en.wikipedia.org/wiki/Large-spotted_civet

### Inspect
anyNA(traits[, c("scientificNameStd", "TrophicLevel", "AdultBodyMass_g")]) # must be F, tho home range could be true 

## double check
setdiff(keep, traits$scientificNameStd) # full match! 
setdiff(traits$scientificNameStd, keep) # full match! 
## should be good to go! 

## keep it clean! 
rm(mam,pan, bird,rep, rep_bird, mam_pan)

```

```{r clean species traits data }

## each macaque species has a different guild (omni, herb, and carn), 
## standardize to omnivore
traits$TrophicLevel[grepl("Macaca", traits$scientificNameStd)] = "Omnivore"
## make all bears omnivores, not herbivore as currently recognized
traits$TrophicLevel[traits$scientificNameStd %in% c("Helarctos_malayanus", "Ursus_thibetanus")] = "Omnivore"
## make all Lophura herbivores
traits$TrophicLevel[grepl("Lophura", traits$scientificNameStd)] = "Herbivore"
## caught a typo from the rep data base to be corrected here
traits$TrophicLevel[traits$TrophicLevel == "Carnivorous"] = "Carnivore"
## Orangutans should be herbivores, not omnivores
traits$TrophicLevel[traits$scientificNameStd == "Pongo_pygmaeus"] = "Herbivore"

# make body mass numeric 
traits$AdultBodyMass_g = as.numeric(traits$AdultBodyMass_g)

## We do NOT want to analyze anything < 1kg in body mass, so remove them here
# which do we lose?
sort(traits$scientificNameStd[traits$AdultBodyMass_g < 1000]) 
## double checked old results for those >500g, and doesnt seem like we will miss much here... REMOVE
traits = traits[traits$AdultBodyMass_g > 1000,]
## also remove anything over 1000 kg, which is too large to be preyed upon 
sort(traits$scientificNameStd[traits$AdultBodyMass_g > 1000000])  # just elephants
traits = traits[traits$AdultBodyMass_g < 1000000,]

## We are also NOT interested in examining domestic animals, so remove them here
traits = traits[! traits$scientificNameStd %in% c("Canis_lupus_familiaris", "Bos_taurus"), ]

## There is also no point analyzing species that are ID'd at the genus level AND species level --> extra noise
traits = traits[! traits$scientificNameStd %in% c("Herpestes_genus", "Hystrix_genus", "Macaca_genus", "Varanus_genus" ), ]

## also do not include species that are primarily arboreal --> orangutans, langur, tufted squirrell any anyone else?
traits = traits[! traits$scientificNameStd %in% c("Pongo_pygmaeus", "Presbytis_rubicunda", "Rheithrosciurus_macrotis"), ]

# thin keep to match 
keep = keep[keep %in% traits$scientificNameStd]

## add small and large to trophic levels 
traits$TrophicGuild = traits$TrophicLevel

# inspect omnivores first --> 50 kg cut off for omnivores
traits$scientificNameStd[traits$TrophicLevel == "Omnivore"] 
traits$TrophicGuild[traits$TrophicLevel == "Omnivore" & 
                      traits$AdultBodyMass_g < 50000] = "Small_Omnivore"
traits$TrophicGuild[traits$TrophicLevel == "Omnivore" & 
                      traits$AdultBodyMass_g > 50000] = "Large_Omnivore"
# inspect
sort(table(traits$TrophicGuild[traits$TrophicLevel == "Omnivore"]))


## herbivores --> 100 kg cut off
traits$scientificNameStd[traits$TrophicLevel == "Herbivore"] # good mix
traits$TrophicGuild[traits$TrophicLevel == "Herbivore" & 
                      traits$AdultBodyMass_g < 100000] = "Small_Herbivore"
traits$TrophicGuild[traits$TrophicLevel == "Herbivore" & 
                      traits$AdultBodyMass_g > 100000] = "Large_Herbivore"
# inspect
sort(table(traits$TrophicGuild[traits$TrophicLevel == "Herbivore"]))

## Carnivores --> 15kg cut off (but just less than 15 to accomodate clouded leopards)
traits$scientificNameStd[traits$TrophicLevel == "Carnivore"] # 
traits$TrophicGuild[traits$TrophicLevel == "Carnivore" & 
                      traits$AdultBodyMass_g < 14940] = "Small_Carnivore"
traits$TrophicGuild[traits$TrophicLevel == "Carnivore" & 
                      traits$AdultBodyMass_g > 14940] = "Large_Carnivore"
# inspect
sort(table(traits$TrophicGuild[traits$TrophicLevel == "Carnivore"]))


table(traits$TrophicGuild) # looks good! 
## many more small than large, but this is to be expected. 


```

```{r determine large carnivore dietary preferences }

### The goal here is to determine if large carnivores have different effects 
## when looking at 'preferred' prey species vs non-preferred species. 
# Ideally, just add a yes-no column to the guilds dataframe for each large carnivore. 

# Diet changes so much based on habitat, interactions, and time of year... 
# All species seem to prefer to some degree Sus scrofa based on readings. 

#### Tigers
# https://zslpublications.onlinelibrary.wiley.com/doi/full/10.1111/j.1469-7998.2011.00871.x
# suggests preferred weight range of 60-250 kg, and wil eat the largest available prey.... So anything >= 60 kg to accomodate tapir, gaur, and cattle
### Change minimum wieght to 17 kg to include muntjac and cite eco-evo paper about it. 
## and can cite Allan Rabinowitz about it: https://thesiamsociety.org/wp-content/uploads/2020/04/NHBSS_037_2k_Rabinowitz_TheDensityAndB.pdf
traits$tiger_pref = "No"
traits$tiger_pref[traits$AdultBodyMass_g > 17000] = "Yes"
traits$tiger_pref[traits$scientificNameStd == "Panthera_tigris"] = "No" # no cannibals allowed. 
traits$tiger_pref[traits$TrophicGuild == "Large_Carnivore"] = "No" # no other large carnivores
# who is preferred?
sort(traits$scientificNameStd[traits$tiger_pref == "Yes"]) # 10 species, ok
# black bear made it in, and this MS says they are preyed upon by tigers in Laos: https://onlinelibrary.wiley.com/doi/full/10.1002/ece3.9067
# Might as well leave sun bear in here too. 
#who is missing?
sort(traits$scientificNameStd[traits$tiger_pref == "No"]) # seems ok, 33 species



#### Leopards 
#  https://doi.org/10.1111/j.1469-7998.2006.00139.x
# suggest preferred weight range of 10-40 kg. 
# https://www.sciencedirect.com/science/article/pii/S0006320712005149 
# langurs/primates, pigs when tigers are absent. 
# https://royalsocietypublishing.org/doi/10.1098/rsos.171187
# suggest large prey (Banteng/Gaur) is possible in the absence of tigers, but not common and restricted to males. 
# https://thesiamsociety.org/wp-content/uploads/2020/04/NHBSS_037_2k_Rabinowitz_TheDensityAndB.pdf
# Muntjac (for tigers too), then primates, then pig == sambar == porcupines == hog badger 
# https://thesiamsociety.org/wp-content/uploads/2020/04/NHBSS_047_1l_Grassman_EcologyAndBehavi.pdf
# Suggests hog badger is important too. 

traits$leopard_pref = "No"
traits$leopard_pref[traits$AdultBodyMass_g > 10000 & traits$AdultBodyMass_g <= 40000] = "Yes"
traits$tiger_pref[traits$scientificNameStd == "Panthera_pardus"] = "No" # no cannibals allowed. 
# who is preferred?
traits$scientificNameStd[traits$leopard_pref == "Yes"] # 4 species, mostly carnivores
## exclude clouded leopards and dholes, leave dogs tho (evidence from india)
traits$leopard_pref[traits$scientificNameStd %in% c("Cuon_alpinus","Neofelis_genus")] = "No"
## also add primates and pigs based off readings
traits$leopard_pref[grepl("Macac", traits$scientificNameStd)] = "Yes" # only include macaques. 
## include pigs and hog badger 
traits$leopard_pref[traits$scientificNameStd %in% c("Sus_scrofa","Arctonyx_collaris")] = "Yes"
# who is preferred now?
sort(traits$scientificNameStd[traits$leopard_pref == "Yes"]) # 7 species, better! 
#who is missing?
sort(traits$scientificNameStd[traits$leopard_pref == "No"]) # seems ok, 38 species --> nothing obvious to include here. 


#### Dholes
# https://zslpublications.onlinelibrary.wiley.com/doi/10.1111/jzo.12171
## Hayward style pref wieght range of 130–190 kg, and highlights sambar as important. 
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9388674/
## scat analysis that emphasizes large deer (sambar + chital), pigs and muntjac were moderetly avoided but still consumed.
# https://doi.org/10.1016/j.mambio.2013.08.007
## lit review of scats and abundance: 40 and 60 kg weight range, sambar preferred most. 
### Matthew suggests a drop down to 10 kg, but wait for citation from him --> would only add binturongs anyway... 
traits$dhole_pref = "No"
traits$dhole_pref[traits$AdultBodyMass_g > 40000 & traits$AdultBodyMass_g < 190000 &
                    traits$TrophicGuild != "Large_Carnivore"] = "Yes"
# who is preferred?
traits$scientificNameStd[traits$dhole_pref == "Yes"] # Some species here dont make sense: bears --> remove! 
# remove bears, no evidence for this! 
traits$dhole_pref[traits$scientificNameStd %in% c("Helarctos_malayanus", "Ursus_thibetanus")] = "No"
## Adding muntjac because they are consumed according to lit (eaten by everything)
traits$dhole_pref[traits$scientificNameStd %in% c("Muntiacus_genus")] = "Yes"
# who is preferred?
traits$scientificNameStd[traits$dhole_pref == "Yes"] # 5 species now... just ok. 
#who is missing?
sort(traits$scientificNameStd[traits$dhole_pref == "No"]) # seems ok, 38 species --> nothing obvious to include here. 


#### Clouded leopards
# https://onlinelibrary.wiley.com/doi/full/10.1002/ece3.9067 
# Serow! Ungulates in particular. No evidence for primates in this citation,
# but evidence from elsewhere for primates, and generally evidence for a varied and wide diet. 
# https://www.sciencedirect.com/science/article/pii/S235198941930976X#bib14
# Evidence for small carnivores as prey --> vivverids (but not other cats) fit! 
# http://cloudedleopardpartners.org/wp-content/uploads/2021/03/2015-First-description-of-population-density-_-habitat-use-of-the-mainland-clouded-leopard-within-a-logged-primary-forest-in-South-East-Asia-PopulationEco.pdf
# Evidence for prey < 10 kg, tho old outdated methods. 
# Justification 
traits$CL_pref = "No"
traits$CL_pref[traits$AdultBodyMass_g > 7000 & traits$AdultBodyMass_g < 190000 &
                 traits$TrophicGuild != "Large_Carnivore"] = "Yes"
# who is preferred?
traits$scientificNameStd[traits$CL_pref == "Yes"] # Remove some of the larger carnivores (e.g. bears, cats), but leaving binturong b/c evidence exists! https://www.researchgate.net/publication/271701967_First_record_of_a_clouded_leo-pard_predating_on_a_binturong
traits$CL_pref[traits$scientificNameStd %in% c("Catopuma_temminckii", "Helarctos_malayanus", "Ursus_thibetanus")] = "No"

# who is preferred?
traits$scientificNameStd[traits$CL_pref == "Yes"] # 12 species now... more than tigers! b/c of small carnivores
#who is missing?
sort(traits$scientificNameStd[traits$CL_pref == "No"]) # seems ok, 32 species --> nothing obvious to include here. 

## What is the median home range for species where its known?
summary(traits$HomeRange_km2) # 3.5 km is the median, 12.4 km is the mean 
# what is the SD?
sd(traits$HomeRange_km2, na.rm = T) #19.2 km 

## whqt about home range for our predators?
summary(traits$HomeRange_km2[traits$scientificNameStd %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")]) 
## order of magnitude difference! median is 33, mean is 39
## one NA that is for clouded leopards. 
sd(traits$HomeRange_km2[traits$scientificNameStd %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")], na.rm = T) #SD is 23.5


## quickly replace space with dash in og caps for reporting
og_caps$Species = gsub(" ", "_", og_caps$Species)

```

Before any standardization there are a total of `r full_length` 'species' with 100 captures ready for analysis, but after removing domestic species (*e.g.*, *Canis lupus familiaris* & *Bos taurus*), primarily arboreal species (*e.g.*, *Pongo pygmaeus* & *Presbytis rubicunda*), species less than 1kg or greater than 1000 kg, or species lacking any trait data, we are left with a __total of `r length(keep)` species ready for analysis__. Therefore, the total number of independent captures for these species is `r nrow(og_caps[og_caps$Species %in% keep,])`.

A total of `r length(unique(traits$TrophicGuild))` trophic guilds were constructed based on trophic level and body mass. Unfortunately, __home-range data was lacking for `r length(traits$scientificNameStd[is.na(traits$HomeRange_km2)])` species__, but for the known home-range values. For the remaining values, the median value is `r median(traits$HomeRange_km2, na.rm = T)` with a range of `r paste(min(traits$HomeRange_km2, na.rm = T), max(traits$HomeRange_km2, na.rm = T), sep = " - " )` square kilometers. 

There are a **total of `r length(traits$scientificNameStd[traits$TrophicLevel == "Herbivore"])` herbivores**, split between **`r length(traits$scientificNameStd[traits$TrophicGuild == "Small_Herbivore"])` small herbivores** and **`r length(traits$scientificNameStd[traits$TrophicGuild == "Large_Herbivore"])` large herbivores** using a 100 kg cut-off. 

There are a **total of `r length(traits$scientificNameStd[traits$TrophicLevel == "Omnivore"])` omnivores**, split between **`r length(traits$scientificNameStd[traits$TrophicGuild == "Small_Omnivore"])` small omnivores** and **`r length(traits$scientificNameStd[traits$TrophicGuild == "Large_Omnivore"])` large omnivores** using a 50 kg cut-off. 

There are a **total of `r length(traits$scientificNameStd[traits$TrophicLevel == "Carnivore"])` carnivores**, split between **`r length(traits$scientificNameStd[traits$TrophicGuild == "Small_Carnivore"])` small carnivores** and **`r length(traits$scientificNameStd[traits$TrophicGuild == "Large_Carnivore"])` large carnivores** using a 15 kg cut-off. 

__Tigers (*Panthera tigris*)__ were generated `r nrow(og_caps[og_caps$Species == "Panthera tigris",])` independent detections from `r length(unique((og_caps$Landscape[og_caps$Species == "Panthera tigris"])))` landscapes. They were deemed to prefer any other species that is not a large carnivore and above 17 kg. This includes a total of **`r length(traits$scientificNameStd[traits$tiger_pref == "Yes"])` species composed of `r paste(traits$scientificNameStd[traits$tiger_pref == "Yes"], collapse = ", ")`**. 

__Leopards (*Panthera pardus*)__ were generated `r nrow(og_caps[og_caps$Species == "Panthera pardus",])` independent detections from `r length(unique((og_caps$Landscape[og_caps$Species == "Panthera pardus"])))` landscapes. They were deemed to prefer any other species that is not a large carnivore and with a weight range of 10-40 kg. Also, all primates were included as preferred prey. This includes a total of **`r length(traits$scientificNameStd[traits$leopard_pref == "Yes"])` species composed of `r paste(traits$scientificNameStd[traits$leopard_pref == "Yes"], collapse = ", ")`**. 

__Dholes (*Cuon alpinus*)__ were generated `r nrow(og_caps[og_caps$Species == "Cuon alpinus",])` independent detections from `r length(unique((og_caps$Landscape[og_caps$Species == "Cuon alpinus"])))` landscapes. They were deemed to prefer any other species that is not a large carnivore and with a weight range of 40-190 kg. Also, all bear species were removed as preferred prey and muntjac deer were included. This includes a total of **`r length(traits$scientificNameStd[traits$dhole_pref == "Yes"])` species composed of `r paste(traits$scientificNameStd[traits$dhole_pref == "Yes"], collapse = ", ")`**. 

__Clouded leopards (*Neofelis* genus)__ were generated `r nrow(og_caps[grepl("Neofelis", og_caps$Species),])` independent detections from `r length(unique((og_caps$Landscape[grepl("Neofelis", og_caps$Species)])))` landscapes. They were deemed to prefer any other species that is not a large carnivore and with a weight range of 7-190 kg. Also, all bear species and golden cats were removed as preferred prey. This includes a total of **`r length(traits$scientificNameStd[traits$CL_pref == "Yes"])` species composed of `r paste(traits$scientificNameStd[traits$CL_pref == "Yes"], collapse = ", ")`**. 


## Prepare data to be saved

Now that relevant species have been determined and species traits are sorted, I will finalize preparing the data by generating a sampling occasion index to create count history matrices for each species using the formatted captures and metadata files. Instead of creating a detection history matrix based on the date a photo was taken, I make all cameras start on the same 'day' (based around start/end dates) to increase model speed and efficiency.

```{r Generate a sampling occasion index in the captues}

## Instead of creating a detection history matrix based on the date a photo was taken
## make all cameras start on the same 'day' to increase model speed and efficiency. 

### Create a data frame with each sampling unit as a row, with start and stop dates
s = distinct(dplyr::select(meta, cell_id, Sampling_begin, Sampling_end))  

## split the dataframe by cell_id, and add the full sequence of dates between start/stop dates. 
s2 = ddply(s, .(cell_id), summarize,
           Date = as.character(seq.Date(from = min(Sampling_begin), to = (max(Sampling_end)), by = 1)))


#That worked, now bring back the start and stops. Make sure you dont loose records
dim(s2) #232760
t = merge(s2, s, by = "cell_id")
dim(t) #232760, good! 


## Add sequence from 1-n for each sampling unit
res = t[0,]
res$seq = numeric()

for(i in unique(t$cell_id)){
  
  d = t[t$cell_id == i,]
  
  d$seq = as.numeric(seq(from= 1, 
                         to = unique(as.numeric(difftime(d$Sampling_end+1, # Add one to start seq on 1 instead of 0
                                                         d$Sampling_begin, units = "days"))), by = 1))
  
  res = rbind(res, d)
}
rm(d,i)

#inspect
str(res)
dim(res) #264568 , same as before! Good! 

## Make sure all cams got accounted for
setdiff(res$cell_id, unique(caps$cell_id))
setdiff(unique(caps$cell_id), res$cell_id) # no difference

## merge the sequence with the captures
t = merge(caps, res, by = c("cell_id", "Date"))
dim(caps) #240177
dim(t) #240058, not exactly the same...

## make a statement in the text below. 
lost =  round((nrow(caps) - nrow(t))/nrow(caps) * 100, 2)


## Captures are all good with updated observation sequence! 
# head(t)
caps = t

## keep your global environment clutter free
rm(t, s, s2, res, full_length)

```

By adding the observation sequence to the captures, we lost `r paste(lost*100, "%", sep = "")` of the captures. This is an acceptable amount. 

```{r Determine species to be analyzed as a group}

## first, determine which species will need to be analyzed as groups rather than individuals 
# because of very large group sizes skewing results
big = ddply(caps[caps$Species %in% keep,], .(Species), summarize,
            max_count = max(total_indiv_records),
            mean_count = mean(total_indiv_records),
            mode_count = median(total_indiv_records))

big$Species[big$mean_count >= 2] #this is probably it... 
# big$Species[big$max_count >= 10] #but this is concerning 

## go w/ mean >= 2 for now and inspect for later
group_sp = big$Species[big$mean_count >= 2]
# adding muntjacs and ALL macaque species because they were probelmatic in the past 
group_sp = c(group_sp, "Muntiacus_genus", "Macaca_fascicularis" )
rm(big)


```

I have also determined which species need to be analyzed as groups instead of inviduals because of very large group sizes cause N-mixture models to crash and create unrealistic estimates. __This includes `r length(group_sp)` species: `r paste(sort(group_sp), collapse = ", ")`__.

```{r Prepare metadata and check for colinearity}

### Found a downstream error! Sampling units w/ too many active cameras were causing models to crash! 
# create a new column in the metadata that counts the number of active cams per SU
meta$cams_included_count = str_count(meta$cameras_included, " - ") + 1
hist(meta$cams_included_count) # classic Poisson

## > 7 cams per SU was causing probelms earlier... 
# how many SUs do we lose by exclusing them?
nrow(meta[meta$cell_id %in% meta$cell_id[meta$cams_included_count > 7],]) / nrow(meta) * 100
# 4% of SUs, thats fine. 

## now begin gathering covairates at the right spatial scale 
flii = names(meta)[grepl("FLLI", names(meta))]
flii = flii[grepl(scale, flii)]
elev = names(meta)[grepl("altitude", names(meta))]
elev = elev[grepl(scale, elev)]
hfp = names(meta)[grepl("human_footprint", names(meta))]
hfp = hfp[grepl(scale, hfp)]
## add more if you want! 
cols = c(flii, elev, hfp)
rm(flii, elev, hfp)


### Verify key covariates  are not strongly correlated
check = meta[, cols]
cov.corr = cor(check[,sapply(check, is.numeric)])
cov.corr= as.data.frame(cov.corr)
# cov.corr[abs(cov.corr) < abs(.6)] = "good" # stick w/ .6
# rm(check, cov.corr)

## make sure to grab other relevant variables
other = c("cell_id", "survey_id", "Landscape", "source", "cams_included_count", "year")

### To ensure that data remains spatially secure regarding sensetive species locations
## remove any spatial variables and thin to only include bare minimum variables. 
meta = meta[, c(cols, other)]


### Also thin down captures to the salient info 
caps = select(caps, cell_id, survey_id, seq, Date, num_cams_active_at_date, Species, total_indiv_records, independent_events)
# and thin to only include relevant species
setdiff(traits$scientificNameStd, caps$Species) # all species in traits are present in caps  
caps = caps[caps$Species %in% traits$scientificNameStd, ]


```

The key spatial variables that will get included in each species' N-mixture model to account for habitat filtering are: **`r paste(cols, collapse = ", ")`**. The correlation coefficient between forest integrity and human footprint is **`r round(cov.corr["Avg_FLLI_5km", "Avg_human_footprint_5km"], 3)`**. The correlation coefficient between elevation and human footprint is **`r round(cov.corr["Avg_altitude_5km", "Avg_human_footprint_5km"], 3)`**. The correlation coefficient between forest integrity and elevation is **`r round(cov.corr["Avg_FLLI_5km", "Avg_altitude_5km"], 3)`**.

## Save data 

Generating the count histories and covariates for each species on a regular computer takes hours, so instead I will send the information to the High Performance Computers (HPC) to run much faster. The data is saved in my personal Dropbox folder, but will move to GitHub eventually. 

```{r Save everything!}

## grab today's date 
day<-str_sub(Sys.Date(),-2)
month<-str_sub(Sys.Date(),-5,-4)
year<-str_sub(Sys.Date(),-10,-7)
date = paste(year,month,day, sep = "")

## grab the WD where we want to save data 
wd = "/Users/zachary_amir/Dropbox/Zach PhD/Ch3 Trophic release project/SEA_TC_GitHub_data_storage/data/step1_output/"


# ### Save all relevant files
# saveRDS(keep, paste(wd, "/species_vector_", length(keep), "_species_", scale, "_scale_", date, ".RDS", sep = ""))
# 
# saveRDS(group_sp, paste(wd, "/group_living_", length(group_sp), "_species_", scale, "_scale_",  date, ".RDS", sep = ""))
# 
# write.csv(caps, paste(wd, "/clean_captures_to_make_UMFs_", scale, "_scale_",  date,".csv", sep = ""), row.names = F)
# 
# write.csv(meta, paste(wd, "/clean_metadata_to_make_UMFs_", scale, "_scale_",  date,".csv", sep = ""), row.names = F)
# 
# ## Also save trait data so I dont have to remake it later
# write.csv(traits, paste(wd, "/clean_", nrow(traits), "_species_trait_data_", date, ".csv", sep = ""), row.names = F)




```

## HPC matrix and covariate creation code

The next section of analysis relies on the HPC, where we convert our covariates and metadata into a count history matrix with the proper associated observation- and site-level covariates. We use the HPC to iterate the process per species with more computational power which vastly speeds up the process. 

The separate R script is called: __scripts/HPC_code/HPC_matrix_generator_SEA_TC.R__.

To make that code run on the HPC, I use the SLURM script called: __scripts/SLURM_code/SLURM_generate_matricies.txt__.


