---
title: "Step 1_Prepare data for bundling"
author: "Zachary Amir"
date: Sys.time()
output: html_document
---

## Introduction

This code will generate count histories and associated covariates and bundle the data to implement Zachary Amir's co-abundance models on the High Performance Computers (HPC).

This code works off the spatially re-sampled captures and covariates that are generated in the 4-step cam trap data standardization pipeline. To learn more about this camera trap history data standardization pipeline, please contact Zachary Amir (z.amir@uq.edu.au) or Matthew Luskin (m.luskin@uq.edu.au) to request access to the following GitHub repository: https://github.com/EcologicalCascadesLab/AsianCaptureHistories 

```{r setup, include=FALSE, echo=FALSE}
## We can knit this into a nice document! but dont include all the code unless otherwise specified (i.e. include=TRUE)
knitr::opts_chunk$set(include=FALSE, warning=FALSE, error = FALSE)

## start fresh
rm(list = ls())

## load libraries
library(tidyverse)
library(here)
library(plyr)
library(traitdata)

```

```{r import and inspect data, include=FALSE}

#### Resampled captures and metadata are saved per survey in a folder in the cap hist DB on DropBox
### This raw data is not available for public usage, but a derived version will be hosted on DropBox. 
##

### captures first 

## import and combine here
files = list.files("/Users/zachary_amir/Dropbox/CT capture histories database/Asian ECL raw CT data/Step5_HPC_import_resampled_captures/")
files = files[!grepl("OLD", files)] # remove old data folder

## save the scale we are interested in 
scale = "5km" # "3km"

## thin to the relevant scale were interested in -- 3km 
files = files[grepl(scale, files)] # additional testing 

## store caps here-
caps = list()

for(i in 1:length(files)){
  
  # select a file 
  f = files[i]
  
  # make the path 
  f_path = paste("/Users/zachary_amir/Dropbox/CT capture histories database/Asian ECL raw CT data/Step5_HPC_import_resampled_captures/", f, sep = "")
  
  # read it
  d = read.csv(f_path)
  
  ## rename cell_id so it is scale agnostic
  names(d)[grepl("cell_id", names(d))] = "cell_id"
  
  ## extract ID tag from file name
  id = strsplit(f, "_spatially_resampled_captures")[[1]][1]
  
  # remove the number
  id = gsub("^\\d+_", "", id)
  
  ## save it! 
  caps[[i]] = d
  names(caps)[i] = id
  
  ## let us know if there are any duplicate files in our repo
  if(any(duplicated(names(caps)))){
    print(paste("There are repeated files from the same survey with the tag:", id,
                "Make sure to remove it before importing!"))
  }# end conditional
  
} # end per file 
rm(f,f_path, d, id, i, files)

## combine list into a df
caps = do.call(rbind, caps)
rownames(caps) = NULL


### metadata next

## import and combine here
files = list.files("/Users/zachary_amir/Dropbox/CT capture histories database/Asian ECL raw CT data/Step5_HPC_import_resampled_metadata/")
files = files[!grepl("OLD", files)] # remove old data folder

## thin to the same scale as above
files = files[grepl(scale, files)]


## store caps here-
meta = list()

for(i in 1:length(files)){
  
  # select a file 
  f = files[i]
  
  # make the path 
  f_path = paste("/Users/zachary_amir/Dropbox/CT capture histories database/Asian ECL raw CT data/Step5_HPC_import_resampled_metadata/", f, sep = "")
  
  # read it
  d = read.csv(f_path)
  
  ## rename cell_id so it is scale agnostic
  names(d)[grepl("cell_id", names(d))] = "cell_id"
  
  ## extract ID tag from file name
  id = strsplit(f, "_spatially_resampled_metadata")[[1]][1]
  
  # remove the number
  id = gsub("^\\d+_", "", id)
  
  ## save it! 
  meta[[i]] = d
  names(meta)[i] = id
  
  ## let us know if there are any duplicate files in our repo
  if(any(duplicated(names(meta)))){
    print(paste("There are repeated files from the same survey with the tag:", id,
                "Make sure to remove it before importing!"))
  }# end conditional
  
} # end per file 
rm(f,f_path, d, id, i, files)

## combine list into a df
meta = do.call(rbind, meta)
rownames(meta) = NULL


## ensure all cell_ids match!
setdiff(meta$cell_id, caps$cell_id)
setdiff(caps$cell_id, meta$cell_id) # full match, good to go! 

## format classes
caps$Date = as.Date(caps$Date, format = "%Y-%m-%d")
meta$Sampling_begin = as.Date(meta$Sampling_begin, format = "%Y-%m-%d")
meta$Sampling_end = as.Date(meta$Sampling_end, format = "%Y-%m-%d")




#### Remove data that does not meet inclusion criteria

## Remove Brodie's data W of wallace line --> doesnt have relevant species
meta = meta[! grepl("E_Indonesia", meta$Landscape),]

# and thin caps to match
caps = caps[caps$survey_id %in% meta$survey_id,]


#### To be conservative, lets thin the data surveys with enough data

## summarize info about surveys
surv_summary = ddply(caps, .(survey_id), summarize,
                     num_SU = length(unique(cell_id)), # total number of sampling units (SUs)
                     max_cam = max(num_cams_active_at_date), # maximum number of cams active in a single sampling unit
                     dur = difftime(max(Date), min(Date), units= "days")) # duration of the survey 
## calculate trap nights 
surv_summary$trap_nights = surv_summary$num_SU * surv_summary$dur

## inspect SUs
summary(surv_summary$num_SU) # good spread, 
# median is 9 @ 3km settings
# median is 8 @ 5km settings

# how many surveys w/ < 3 SUs? ( ~ 75% of 1st quarter)
length(surv_summary$survey_id[surv_summary$num_SU < 3]) / length(surv_summary$survey_id) * 100
## lose 15% just based on SU.
# A landscape should be at least 9 km2 to provide spatial variation, so lets stick with 3 as the minimum SUs.

## inspect effort
summary(as.numeric(surv_summary$trap_nights)) # good spread, 
# median is 765 @ 3km 
# median is 591 @ 5km 

# how many surveys w/ < 100 tn ( ~ 33% of 1st quarter)
length(surv_summary$survey_id[surv_summary$trap_nights < 100]) / length(surv_summary$survey_id) * 100
## lose 12% just based on effort, probably overlap w/ above. 
# A survey should at least have 100 trapnights of effort as minimum for inclusion. 

## Extract surveys that do not meet inclusion criteria. 
rm = surv_summary$survey_id[as.numeric(surv_summary$trap_nights) < 100] # less than 100 trap nights 
rm2 = surv_summary$survey_id[(surv_summary$num_SU) < 3] # less than 3 sampling units in a survey 
rm = unique(c(rm,rm2))

## make a statement about it! 
print(paste("The number of rows that will be excluded from analysis because it is a survey w/ less than 100 trap nights or less than 3 sampling units in a survey are:", 
            dim(caps[caps$survey_id %in% rm,])[1],
            "and this represents:", 
            round(dim(caps[caps$survey_id %in% rm,])[1]  / dim(caps)[1] * 100, 3),
            "% of the full data."))
# nice! minimal losses. 
## only 3.4% and 8882 rows removed from caps @ 3km res
## only 5.4% and 13594 rows removed from caps @ 5km res


## remove the data from caps
caps = caps[!caps$survey_id %in% rm,]

## and metadata
meta = meta[meta$survey_id %in% caps$survey_id,]

## keep environment clean
rm(surv_summary, rm,rm2)


## need survey year in covariates 
meta$year = str_extract(meta$survey_id, stringr::regex("(\\d+)(?!.*\\d)"))
sort(table(meta$year)) #looks good! 


```

### Determine which species to analyze

Here I will determine which species have sufficient detections (n = 100), standardize similar species to the genera level (e.g. Tragulus), and remove non-relevant species (e.g. blanks)

```{r select and standardize species to include, include=FALSE}

##### Which species will be included and how to standardize species?

## replace space with underscore to facilitate clean code/data
caps$Species = gsub(" ", "_", caps$Species)

# which species have sufficent (n=100) detections?
sp_count = ddply(caps, .(Species), summarize,
                 num_cap = length(Species))
keep = sp_count[sp_count$num_cap >= 100,]

## what did we get?
keep[order(keep$num_cap),] ## Awesome!! 
length(keep$Species) # 106 species before any cleaning 


## Standardize muntjacs, tragulus, Hystrix (?), Neofelis, Capricornis, Tupaia, maybe Canis_lupus
## remove from analysis: Ghost Homo_sapiens, unID'd, remove, Aves, bird, unkown, bird_pheasent, etc


### Clean up and combine some speices names

# Hystrix
sort(table(caps$Species[grepl("Hystrix", caps$Species)])) ## dont change crassispinis or brachyura b/c we have enough! 
# caps$Species[caps$Species %in% c("Hystrix_brachyura", "Hystrix")] = "Hystrix_genus"

# Muntjac
sort(table(caps$Species[grepl("Munt", caps$Species)])) # ecologically very similar species, but what about spatial distribution?
## where do all these muntjacs come from?
Ma = caps$cell_id[caps$Species == "Muntiacus_atherodes"]
unique(meta$Landscape[meta$cell_id %in% Ma]) # Sarawak and Sabah.
Mr = caps$cell_id[caps$Species == "Muntiacus_reevesi"]
unique(meta$Landscape[meta$cell_id %in% Mr]) # Sarawak, Sabah, Sumatra, and S. Thailand...
Mv = caps$cell_id[caps$Species == "Muntiacus_vaginalis"]
unique(meta$Landscape[meta$cell_id %in% Mv]) # Thailand, Vietnam, China
Mm = caps$cell_id[caps$Species == "Muntiacus_muntjak"]
sort(unique(meta$Landscape[meta$cell_id %in% Mm])) # Laos, P_Malaysia, Sarawak, Sabah, Sumatra, Thailand (many)
M = caps$cell_id[caps$Species == "Muntiacus"]
sort(unique(meta$Landscape[meta$cell_id %in% M])) # Laos, Sarawak, Thailand (many), Vietnam.
## These are all quite spatially distinct, so there is good motivation to combine these species into one
# Make sure to save independent detections of all for supplementary materials later! 
rm(Ma, Mr, Mv, Mm, M)

## combine
caps$Species[grepl("Munt", caps$Species)] = "Muntiacus_genus"

# clouded leopards
sort(table(caps$Species[grepl("Neof", caps$Species)])) #just geographical differences, but way more bornean than mainland! 
caps$Species[grepl("Neof", caps$Species)] = "Neofelis_genus"

# mousedeer
sort(table(caps$Species[grepl("Trag", caps$Species)])) #Amazed that these are even differentiated 
caps$Species[grepl("Trag", caps$Species)] = "Tragulus_genus"

# serow
sort(table(caps$Species[grepl("Capric", caps$Species)])) #I think these were recently combined by taxonomists anyway 
caps$Species[grepl("Capric", caps$Species)] = "Capricornis_genus"

# treeshrew
sort(table(caps$Species[grepl("Tupai", caps$Species)])) #Amazed that these are even differentiated 
caps$Species[grepl("Tupai", caps$Species)] = "Tupaia_genus"

#dogs
sort(table(caps$Species[grepl("Canis_lup", caps$Species)])) #Just call these dogs, we dont have pics to know if domestic or not 
caps$Species[grepl("Canis_lup", caps$Species)] = "Canis_lupus_familiaris"

# mongoose
sort(table(caps$Species[grepl("Herpe", caps$Species)])) # Just stick with H_urva and H_brachyrus 
caps$Species[caps$Species %in% c("Herpestes", "Herpestes_semitorquatus",
                                 "Herpestidae")] = "Herpestes_genus" # this will get excluded later anyway b/c we have more specific species, which is better. 

# Murids
unique(caps$Species[grepl("Muri", caps$Species)]) #only the family muridae
caps$Species[grepl("Muri", caps$Species)] = "Muridae_spp" #spp denotes family here. 

#squirells 
unique(caps$Species[grepl("Sciur", caps$Species)]) #only the family sciuridae
caps$Species[grepl("Sciur", caps$Species)] = "Sciuridae_spp"


#Combine northern and southern pig tail macaques into one species Macaca_leonina Macaca_nemestrina 
sort(table(caps$Species[grepl("Macaca_", caps$Species)]))
caps$Species[caps$Species %in% c("Macaca_leonina", "Macaca_nemestrina" )] = "Macaca_nemestrina"

#chickens
unique(caps$Species[grepl("Gallu", caps$Species)]) #analyze at genus level. 
caps$Species[startsWith(caps$Species, "Gallus")] = "Gallus_genus"

## attach genus to the other Genera names that need it to stay consistent
caps$Species[caps$Species %in% c("Amaurornis", "Macaca", "Hystrix",
                                 "Maxomys","Melogale", "Varanus")] = paste(caps$Species[caps$Species %in% c("Amaurornis", "Macaca",
                                                                                                            "Hystrix","Maxomys","Melogale", "Varanus")], "genus", sep = "_")


### Make a new species count w/ updated species names
sp_count = ddply(caps, .(Species), summarize,
                 num_cap = length(Species))
keep = sp_count[sp_count$num_cap >= 100,]

## what did we get?
sort(keep$Species) #reduced to 91 species

# remove some of the not useful fluff
keep = keep[!keep$Species %in% c("Aves", "Ghost", "Mammalia",
                                 "Unknown","Vehicle"),] 

#remove all  people 
keep = keep[!grepl("Homo", keep$Species ),] 

# remove species not ID to the species or genus level 
keep = keep[!endsWith(keep$Species, "_spp"),]
# and the others. 
keep = keep[! keep$Species %in% c("Scincidae","Small_mammal", "Herpestidae",
                                  "Phasianidae", "Rodentia","Scandentia"),]

length(keep$Species) # 73 species after removing fluff. 
rm(sp_count)

### Looks good, but will remove more species once body weights are accounted for. 
sort(keep$Species)

## save it as a vector, not dataframe
keep = keep$Species

```

### Determine species traits

Here I am generating all species data from the library(speciestraits) databases, cleaning where necessary, and generating trait data for combined genera based off representative species. Each species should have 1) a trophic level (i.e., carnivore, omnivore, herbivore), and body mass (in grams). I will also add information about the dietary preferences for the 4 main large carnivores for each species, where the column will be a "Yes" or "No" if the literature supports it being a preferred prey species. I will try to include missing information from outside sources, but if none is available, species without trait data will be discarded.

```{r generate species trait data, include=FALSE}

## access panthera dataset
pan = pantheria
# replace space with underscore for species names to match our style
pan$scientificNameStd = gsub(" ", "_", pan$scientificNameStd)

## access mammal_diet dataset
mam = mammal_diet2
mam = as.data.frame(mam) # remove silly tibble
# replace space with underscore for species names to match our style
mam$scientificNameStd = gsub(" ", "_", mam$scientificNameStd)
## remove NA species names
mam = mam[!is.na(mam$scientificNameStd),]


## import Sam Lee's cleaned avonet dataframe
bird = read.csv(here::here("data_GitHub/bird_guild_information_20221102.csv"))

# replace space with underscore for bird species names
bird$Species = gsub(" ", "_", bird$Species)

## check which species we have matches for 
intersect(bird$Species, keep) # 12 matches and all birds, great! 
# thin bird to relevant species
bird = bird[bird$Species %in% c(keep, "Gallus_gallus"),]
bird$Species[bird$Species == "Gallus_gallus"] = "Gallus_genus" #std to match

## copy "Amaurornis_phoenicurus" and provide for genus Amaurornis catch-all
am = bird[bird$Species == "Amaurornis_phoenicurus",]
am$Species = "Amaurornis_genus" 
bird = rbind(bird,am)
rm(am)

## access a dataset for varanus salvator 
rep = lizard_traits
# thin to our singular herp species
rep = rep[rep$scientificNameStd == "Varanus salvator",] # good! 
rep = rep[!is.na(rep$scientificNameStd),] # idk why, but it comes with lots of NA
# make sure species name is consistent!
rep$scientificNameStd = gsub(" ", "_", rep$scientificNameStd)
# thin to relevant info
rep = select(rep, scientificNameStd, diet) # missing body wieght and home range 

## adding custom body weight as determined by Shine et al 1996, "Commercial harvesting of giant lizards: The biology of water monitors Varanus salvator in southern Sumatra"
rep$Body_mass = 3.24
names(rep)[2] = "TrophicLevel"

## add a second row of rep for Varanus catch-all genus
rep = rbind(rep,rep)
rep$scientificNameStd[2] = "Varanus_genus"

## rbind w/ our birds
names(bird)[1:2] = c("scientificNameStd","TrophicLevel")
rep_bird = rbind(rep, bird)

# convert kg masses to grams to match mammal format 
rep_bird$Body_mass = rep_bird$Body_mass * 1000
names(rep_bird)[3] = "AdultBodyMass_g"

## dont know home range for these guys, so just leave as NA
rep_bird$HomeRange_km2 = NA

## thin mam and pan to relevant info while incorporating known typo
mam = select(mam[mam$scientificNameStd %in% c(keep, "Taperus_indicus", 
                                              "Canis_lupus"),], scientificNameStd, TrophicLevel) # diet from here
pan = select(pan[pan$scientificNameStd %in% c(keep, "Taperus_indicus", 
                                              "Canis_lupus"),], scientificNameStd, AdultBodyMass_g, HomeRange_km2) # mass and home range from here

# and merge together
mam_pan = distinct(merge(mam, pan, by = "scientificNameStd"))

# There are two duplicated rows, remove them! 
a = mam_pan[mam_pan$scientificNameStd == "Callosciurus_notatus" & !is.na(mam_pan$AdultBodyMass_g),] # remove dup w/out mass
b = mam_pan[mam_pan$scientificNameStd == "Atherurus_macrourus" & 
              mam_pan$TrophicLevel == "Herbivore",] # remove dup w/ omnivore tag for porcupine
c = rbind(a,b)
## remove bad species
mam_pan = mam_pan[!mam_pan$scientificNameStd %in% c("Atherurus_macrourus", "Callosciurus_notatus"),]
## and rbind good records of those species
mam_pan = rbind(mam_pan, c)
rm(a,b,c)


# Now combine them all!
traits = rbind(mam_pan, rep_bird)

## check which species are missing 
setdiff(keep, traits$scientificNameStd) # 12, though mostly genera or a few typos 

## fix obvious typos first
traits$scientificNameStd[traits$scientificNameStd == "Taperus_indicus"] = "Tapirus_indicus"
traits$scientificNameStd[traits$scientificNameStd == "Canis_lupus"] = "Canis_lupus_familiaris"

## add 10 missing species to traits object
traits[64:(64+9), "scientificNameStd"] = setdiff(keep, traits$scientificNameStd)

### re-load mam and pam to include other species
# pan
pan = pantheria
pan$scientificNameStd = gsub(" ", "_", pan$scientificNameStd)
pan = select(pan, scientificNameStd, AdultBodyMass_g, HomeRange_km2)
# mam
mam = mammal_diet2
mam = as.data.frame(mam) # remove silly tibble
mam$scientificNameStd = gsub(" ", "_", mam$scientificNameStd)
mam = mam[!is.na(mam$scientificNameStd),]
mam = select(mam, scientificNameStd, TrophicLevel)


## add info per species/guild
#serow
traits[traits$scientificNameStd == "Capricornis_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Capricornis_sumatraensis"]
traits[traits$scientificNameStd == "Capricornis_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Capricornis_sumatraensis", c("AdultBodyMass_g", "HomeRange_km2")]

# mongooses 
traits[traits$scientificNameStd == "Herpestes_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Herpestes_urva"]
traits[traits$scientificNameStd == "Herpestes_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Herpestes_urva", c("AdultBodyMass_g", "HomeRange_km2")]

#muntjacs
traits[traits$scientificNameStd == "Muntiacus_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Muntiacus_muntjak"])
traits[traits$scientificNameStd == "Muntiacus_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Muntiacus_muntjak", c("AdultBodyMass_g", "HomeRange_km2")]

#clouded leopards
traits[traits$scientificNameStd == "Neofelis_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Neofelis_nebulosa"]
traits[traits$scientificNameStd == "Neofelis_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Neofelis_nebulosa", c("AdultBodyMass_g", "HomeRange_km2")]

#mousedeer
traits[traits$scientificNameStd == "Tragulus_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Tragulus_napu"]
traits[traits$scientificNameStd == "Tragulus_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Tragulus_napu", c("AdultBodyMass_g", "HomeRange_km2")]

#gaur
traits[traits$scientificNameStd == "Bos_gaurus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Bos_javanicus"]
traits[traits$scientificNameStd == "Bos_gaurus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Bos_javanicus", c("AdultBodyMass_g", "HomeRange_km2")]

#porcupines
traits[traits$scientificNameStd == "Hystrix_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Hystrix_brachyura"])
traits[traits$scientificNameStd == "Hystrix_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Hystrix_brachyura", c("AdultBodyMass_g", "HomeRange_km2")]

#maxomys
traits[traits$scientificNameStd == "Maxomys_genus", "TrophicLevel"] = mam$TrophicLevel[mam$scientificNameStd == "Maxomys_rajah"] # widespread in Sundaland
traits[traits$scientificNameStd == "Maxomys_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Maxomys_rajah", c("AdultBodyMass_g", "HomeRange_km2")]

#Tupaia_genus
traits[traits$scientificNameStd == "Tupaia_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Tupaia_glis"])
traits[traits$scientificNameStd == "Tupaia_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Tupaia_glis", c("AdultBodyMass_g", "HomeRange_km2")]

#Macaca
traits[traits$scientificNameStd == "Macaca_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Macaca_nemestrina"]) # maybe silly b/c already present?
traits[traits$scientificNameStd == "Macaca_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Macaca_nemestrina", c("AdultBodyMass_g", "HomeRange_km2")]

#Melogale
traits[traits$scientificNameStd == "Melogale_genus", "TrophicLevel"] = unique(mam$TrophicLevel[mam$scientificNameStd == "Melogale_moschata"]) 
traits[traits$scientificNameStd == "Melogale_genus", c("AdultBodyMass_g", "HomeRange_km2")] = pan[pan$scientificNameStd == "Melogale_moschata", c("AdultBodyMass_g", "HomeRange_km2")]

### two civets are missing weights!! 
# Hose civet
traits$AdultBodyMass_g[traits$scientificNameStd == "Diplogale_hosei"] = 2500 #https://www.ecologyasia.com/verts/mammals/hose's-civet.htm
# large spotted civet
traits$AdultBodyMass_g[traits$scientificNameStd == "Viverra_megaspila"] = 6600 #https://en.wikipedia.org/wiki/Large-spotted_civet

### Inspect
anyNA(traits[, c("scientificNameStd", "TrophicLevel", "AdultBodyMass_g")]) # must be F, tho home range could be true 
# traits[is.na(traits$AdultBodyMass_g),]

## double check
setdiff(keep, traits$scientificNameStd) # full match! 
setdiff(traits$scientificNameStd, keep) # full match! 
## should be good to go! 

rm(mam,pan, bird,rep, rep_bird, mam_pan)



## each macaque species has a different guild (omni, herb, and carn), 
## standardize to omnivore
traits$TrophicLevel[grepl("Macaca", traits$scientificNameStd)] = "Omnivore"
## make all bears omnivores, not herbivore as currently recognized
traits$TrophicLevel[traits$scientificNameStd %in% c("Helarctos_malayanus", "Ursus_thibetanus")] = "Omnivore"
## make all Lophura herbivores
traits$TrophicLevel[grepl("Lophura", traits$scientificNameStd)] = "Herbivore"
## caught a typo from the rep data base to be corrected here
traits$TrophicLevel[traits$TrophicLevel == "Carnivorous"] = "Carnivore"
## Orangutans should be herbivores, not omnivores
traits$TrophicLevel[traits$scientificNameStd == "Pongo_pygmaeus"] = "Herbivore"

# make body mass numeric 
traits$AdultBodyMass_g = as.numeric(traits$AdultBodyMass_g)

## We do NOT want to analyze anything < 1kg in body mass, so remove them here
# which do we lose?
sort(traits$scientificNameStd[traits$AdultBodyMass_g < 1000]) 
## double checked old results for those >500g, and doesnt seem like we will miss much here... REMOVE
traits = traits[traits$AdultBodyMass_g > 1000,]
## also remove anything over 1000 kg, which is too large to be preyed upon 
sort(traits$scientificNameStd[traits$AdultBodyMass_g > 1000000])  # just elephants
traits = traits[traits$AdultBodyMass_g < 1000000,]

## We are also NOT interested in examining domestic animals, so remove them here
traits = traits[! traits$scientificNameStd %in% c("Canis_lupus_familiaris", "Bos_taurus"), ]

## There is also no point analyzing species that are ID'd at the genus level AND species level --> extra noise
traits = traits[! traits$scientificNameStd %in% c("Herpestes_genus", "Hystrix_genus", "Macaca_genus", "Varanus_genus" ), ]

## also do not include species that are primarily arboreal --> orangutans, langur, tufted squirrell any anyone else?
traits = traits[! traits$scientificNameStd %in% c("Pongo_pygmaeus", "Presbytis_rubicunda", "Rheithrosciurus_macrotis"), ]

# thin keep to match 
keep = keep[keep %in% traits$scientificNameStd]



## add small and large to trophic levels 
traits$TrophicGuild = traits$TrophicLevel

# inspect omnivores first --> 50 kg cut off for omnivores
traits$scientificNameStd[traits$TrophicLevel == "Omnivore"] 
traits$TrophicGuild[traits$TrophicLevel == "Omnivore" & 
                      traits$AdultBodyMass_g < 50000] = "Small_Omnivore"
traits$TrophicGuild[traits$TrophicLevel == "Omnivore" & 
                      traits$AdultBodyMass_g > 50000] = "Large_Omnivore"
# inspect
sort(table(traits$TrophicGuild[traits$TrophicLevel == "Omnivore"]))


## herbivores --> 100 kg cut off
traits$scientificNameStd[traits$TrophicLevel == "Herbivore"] # good mix
traits$TrophicGuild[traits$TrophicLevel == "Herbivore" & 
                      traits$AdultBodyMass_g < 100000] = "Small_Herbivore"
traits$TrophicGuild[traits$TrophicLevel == "Herbivore" & 
                      traits$AdultBodyMass_g > 100000] = "Large_Herbivore"
# inspect
sort(table(traits$TrophicGuild[traits$TrophicLevel == "Herbivore"]))

## Carnivores --> 15kg cut off (but just less than 15 to accomodate clouded leopards)
traits$scientificNameStd[traits$TrophicLevel == "Carnivore"] # 
traits$TrophicGuild[traits$TrophicLevel == "Carnivore" & 
                      traits$AdultBodyMass_g < 14940] = "Small_Carnivore"
traits$TrophicGuild[traits$TrophicLevel == "Carnivore" & 
                      traits$AdultBodyMass_g > 14940] = "Large_Carnivore"
# inspect
sort(table(traits$TrophicGuild[traits$TrophicLevel == "Carnivore"]))


table(traits$TrophicGuild) # looks good! 
## many more small than large, but this is to be expected. 

#
##
###
#### Add dietary preferences of the large carnivores

### The goal here is to determine if large carnivores have different effects 
## when looking at 'preferred' prey species vs non-preferred species. 
# Ideally, just add a yes-no column to the guilds dataframe for each large carnivore. 

# Diet changes so much based on habitat, interactions, and time of year... 
# All species seem to prefer to some degree Sus scrofa based on readings. 

#### Tigers
# https://zslpublications.onlinelibrary.wiley.com/doi/full/10.1111/j.1469-7998.2011.00871.x
# suggests preferred weight range of 60-250 kg, and wil eat the largest available prey.... So anything >= 60 kg to accomodate tapir, gaur, and cattle
### Change minimum wieght to 17 kg to include muntjac and cite eco-evo paper about it. 
## and can cite Allan Rabinowitz about it: https://thesiamsociety.org/wp-content/uploads/2020/04/NHBSS_037_2k_Rabinowitz_TheDensityAndB.pdf
traits$tiger_pref = "No"
traits$tiger_pref[traits$AdultBodyMass_g > 17000] = "Yes"
traits$tiger_pref[traits$scientificNameStd == "Panthera_tigris"] = "NA" # no cannibals allowed. 
# who is preferred?
sort(traits$scientificNameStd[traits$tiger_pref == "Yes"]) # 10 species, ok
# black bear made it in, and this MS says they are preyed upon by tigers in Laos: https://onlinelibrary.wiley.com/doi/full/10.1002/ece3.9067
# Might as well leave sun bear in here too. 
#who is missing?
sort(traits$scientificNameStd[traits$tiger_pref == "No"]) # seems ok, 33 species

## Make sure there are no tigers preferentially prey upon other big cats
traits$tiger_pref[grepl("Panthera", traits$scientificNameStd)] = "No" 


#### Leopards 
#  https://doi.org/10.1111/j.1469-7998.2006.00139.x
# suggest preferred weight range of 10-40 kg. 
# https://www.sciencedirect.com/science/article/pii/S0006320712005149 
# langurs/primates, pigs when tigers are absent. 
# https://royalsocietypublishing.org/doi/10.1098/rsos.171187
# suggest large prey (Banteng/Gaur) is possible in the absence of tigers, but not common and restricted to males. 
# https://thesiamsociety.org/wp-content/uploads/2020/04/NHBSS_037_2k_Rabinowitz_TheDensityAndB.pdf
# Muntjac (for tigers too), then primates, then pig == sambar == porcupines == hog badger 
# https://thesiamsociety.org/wp-content/uploads/2020/04/NHBSS_047_1l_Grassman_EcologyAndBehavi.pdf
# Suggests hog badger is important too. 

traits$leopard_pref = "No"
traits$leopard_pref[traits$AdultBodyMass_g > 10000 & traits$AdultBodyMass_g <= 40000] = "Yes"
traits$tiger_pref[traits$scientificNameStd == "Panthera_pardus"] = "NA" # no cannibals allowed. 
# who is preferred?
traits$scientificNameStd[traits$leopard_pref == "Yes"] # 4 species, mostly carnivores
## exclude clouded leopards and dholes, leave dogs tho (evidence from india)
traits$leopard_pref[traits$scientificNameStd %in% c("Cuon_alpinus","Neofelis_genus")] = "No"
## also add primates and pigs based off readings
traits$leopard_pref[grepl("Macac", traits$scientificNameStd)] = "Yes" # only include macaques. 
## include pigs and hog badger 
traits$leopard_pref[traits$scientificNameStd %in% c("Sus_scrofa","Arctonyx_collaris")] = "Yes"
# who is preferred now?
sort(traits$scientificNameStd[traits$leopard_pref == "Yes"]) # 7 species, better! 
#who is missing?
sort(traits$scientificNameStd[traits$leopard_pref == "No"]) # seems ok, 38 species --> nothing obvious to include here. 


#### Dholes
# https://zslpublications.onlinelibrary.wiley.com/doi/10.1111/jzo.12171
## Hayward style pref wieght range of 130–190 kg, and highlights sambar as important. 
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9388674/
## scat analysis that emphasizes large deer (sambar + chital), pigs and muntjac were moderetly avoided but still consumed.
# https://doi.org/10.1016/j.mambio.2013.08.007
## lit review of scats and abundance: 40 and 60 kg weight range, sambar preferred most. 
### Matthew suggests a drop down to 10 kg, but wait for citation from him --> would only add binturongs anyway... 
traits$dhole_pref = "No"
traits$dhole_pref[traits$AdultBodyMass_g > 40000 & traits$AdultBodyMass_g < 190000 &
                    traits$TrophicGuild != "Large_Carnivore"] = "Yes"
# who is preferred?
traits$scientificNameStd[traits$dhole_pref == "Yes"] # Some species here dont make sense: bears --> remove! 
# remove bears, no evidence for this! 
traits$dhole_pref[traits$scientificNameStd %in% c("Helarctos_malayanus", "Ursus_thibetanus")] = "No"
## Adding muntjac because they are consumed according to lit (eaten by everything)
traits$dhole_pref[traits$scientificNameStd %in% c("Muntiacus_genus")] = "Yes"
# who is preferred?
traits$scientificNameStd[traits$dhole_pref == "Yes"] # 5 species now... just ok. 
#who is missing?
sort(traits$scientificNameStd[traits$dhole_pref == "No"]) # seems ok, 38 species --> nothing obvious to include here. 


#### Clouded leopards
# https://onlinelibrary.wiley.com/doi/full/10.1002/ece3.9067 
# Serow! Ungulates in particular. No evidence for primates in this citation,
# but evidence from elsewhere for primates, and generally evidence for a varied and wide diet. 
# https://www.sciencedirect.com/science/article/pii/S235198941930976X#bib14
# Evidence for small carnivores as prey --> vivverids (but not other cats) fit! 
# http://cloudedleopardpartners.org/wp-content/uploads/2021/03/2015-First-description-of-population-density-_-habitat-use-of-the-mainland-clouded-leopard-within-a-logged-primary-forest-in-South-East-Asia-PopulationEco.pdf
# Evidence for prey < 10 kg, tho old outdated methods. 
# Justification 
traits$CL_pref = "No"
traits$CL_pref[traits$AdultBodyMass_g > 7000 & traits$AdultBodyMass_g < 190000 &
                 traits$TrophicGuild != "Large_Carnivore"] = "Yes"
# who is preferred?
traits$scientificNameStd[traits$CL_pref == "Yes"] # Remove some of the larger carnivores (e.g. bears, cats), but leaving binturong b/c evidence exists! https://www.researchgate.net/publication/271701967_First_record_of_a_clouded_leo-pard_predating_on_a_binturong
traits$CL_pref[traits$scientificNameStd %in% c("Catopuma_temminckii", "Helarctos_malayanus", "Ursus_thibetanus")] = "No"

# who is preferred?
traits$scientificNameStd[traits$CL_pref == "Yes"] # 12 species now... more than tigers! b/c of small carnivores
#who is missing?
sort(traits$scientificNameStd[traits$CL_pref == "No"]) # seems ok, 32 species --> nothing obvious to include here. 

## What is the median home range for species where its known?
summary(traits$HomeRange_km2) # 3.5 km is the median, 12.4 km is the mean 
# what is the SD?
sd(traits$HomeRange_km2, na.rm = T) #19.2 km 

## whqt about home range for our predators?
summary(traits$HomeRange_km2[traits$scientificNameStd %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")]) 
## order of magnitude difference! median is 33, mean is 39
## one NA that is for clouded leopards. 
sd(traits$HomeRange_km2[traits$scientificNameStd %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")], na.rm = T) #SD is 23.5


```

### Prepare data to be bundled

Now that relevant species have been determined and species traits are sorted out, I can begin formatting the data to run in the co-abundance models. This will involve making a count history matrix for each species and generating key covariates to be included as well. Will be using the captures and metadata files imported from the first code chunk to do this.

First, I will generate a sampling occasion index in the captures. Instead of creating a detection history matrix based on the date a photo was taken, I make all cameras start on the same 'day' (based around start/end dates) to increase model speed and efficiency.

```{r Generate a sampling occasion index in the captues, echo=F, include=FALSE}

## Instead of creating a detection history matrix based on the date a photo was taken
## make all cameras start on the same 'day' to increase model speed and efficiency. 

### Create a data frame with each sampling unit as a row, with start and stop dates
s = distinct(dplyr::select(meta, cell_id, Sampling_begin, Sampling_end))  

## split the dataframe by cell_id, and add the full sequence of dates between start/stop dates. 
s2 = ddply(s, .(cell_id), summarize,
           Date = as.character(seq.Date(from = min(Sampling_begin), to = (max(Sampling_end)), by = 1)))


#That worked, now bring back the start and stops. Make sure you dont loose records
dim(s2) #232760
t = merge(s2, s, by = "cell_id")
dim(t) #232760, good! 


## Add sequence from 1-n for each sampling unit
res = t[0,]
res$seq = numeric()

for(i in unique(t$cell_id)){
  
  d = t[t$cell_id == i,]
  
  d$seq = as.numeric(seq(from= 1, 
                         to = unique(as.numeric(difftime(d$Sampling_end+1, # Add one to start seq on 1 instead of 0
                                                         d$Sampling_begin, units = "days"))), by = 1))
  
  res = rbind(res, d)
}
rm(d,i)

#inspect
str(res)
dim(res) #264568 , same as before! Good! 

## Make sure all cams got accounted for
setdiff(res$cell_id, unique(caps$cell_id))
setdiff(unique(caps$cell_id), res$cell_id) # no difference

## merge the sequence with the captures
t = merge(caps, res, by = c("cell_id", "Date"))
dim(caps) #240177
dim(t) #240058, not exactly the same...

## How much did we lose?
print(paste("By adding the observation sequence to the captures, we lost ", 
             round((nrow(caps) - nrow(t))/nrow(caps) * 100, 2), "% of the data", sep = ""))
## this is an acceptable amount 

## Captures are all good with updated observation sequence! 
# head(t)
caps = t

## keep your global environment clutter free
rm(t, s, s2, res)

```

Generating the data bundles on a regular computer will take hours, so instead I will send the information to the High Performance Computers (HPC) to run much faster. Before sending off the data, I will determine which species need to be analyzed as groups (because of very large group sizes, e.g. dholes and macaques).

```{r save data to send to HPC for bundling, echo=FALSE}

## first, determine which species will need to be analyzed as groups rather than individuals 
# because of very large group sizes skewing results
big = ddply(caps[caps$Species %in% keep,], .(Species), summarize,
            max_count = max(total_indiv_records),
            mean_count = mean(total_indiv_records),
            mode_count = median(total_indiv_records))

big$Species[big$mean_count >= 2] #this is probably it... 
# big$Species[big$max_count >= 10] #but this is concerning 

## go w/ mean >= 2 for now and inspect for later
group_sp = big$Species[big$mean_count >= 2]
# adding muntjacs and ALL macaque species because they were probelmatic in the past 
group_sp = c(group_sp, "Muntiacus_genus", "Macaca_fascicularis" )
rm(big)


### Found a downstream error! Sampling units w/ too many active cameras were causing models to crash! 
# create a new column in the metadata that counts the number of active cams per SU
meta$cams_included_count = str_count(meta$cameras_included, " - ") + 1
hist(meta$cams_included_count) # classic Poisson

## > 7 cams per SU was causing probelms earlier... 
# how many SUs do we lose by exclusing them?
nrow(meta[meta$cell_id %in% meta$cell_id[meta$cams_included_count > 7],]) / nrow(meta) * 100
# 4% of SUs, thats fine. 


# ### Also verify key covariates we are intersted in are not strongly correlated 
# check = select(meta, Avg_FLLI_3km, Avg_human_footprint_3km, Avg_altitude_3km, Avg_hunting_accessibility_3km)
# cov.corr = cor(check[,sapply(check, is.numeric)])
# cov.corr= as.data.frame(cov.corr)
# cov.corr[abs(cov.corr) < abs(.6)] = "good" # stick w/ .6
# rm(check, cov.corr)

### To ensure that data remains spatially secure regarding sensetive species locations
## remove any spatial variables and thin to only include bare minimum variables. 
sort(names(meta))
# meta = select(meta, cell_id, survey_id, Landscape, source, cams_included_count, year, 
#               ## Environmental vars 
#               Avg_altitude_3km, Avg_ecoregion_intactness_3km, Avg_FLLI_3km, 
#               Avg_forest_cover_3km, 
#               ## Anthropogenic vars 
#               Avg_human_footprint_3km,Avg_hunting_accessibility_3km, Avg_nighttime_Lights_3km,
#               Avg_Islam_Pop_3km, Avg_Christ_Pop_3km, Avg_Budd_Pop_3km, 
#               ## degraded land cover
#               Avg_OP_percent_3km, Avg_LM_percent_3km, Avg_LO_percent_3km, 
#               Avg_RP_percent_3km,
#               ## intact land cover
#               Avg_LEF_percent_3km, Avg_LMEF_percent_3km, Avg_UMEF_percent_3km, 
#               Avg_LDF_percent_3km, Avg_LMDF_percent_3km)
## The 5km version! I should really br doing this based on the scale used... 
## maybe like grab all vars at that scale, select them, then rename so it is scale agnositc like cell_id?
meta = select(meta, cell_id, survey_id, Landscape, source, cams_included_count, year, 
              ## Environmental vars 
              Avg_altitude_5km, Avg_ecoregion_intactness_5km, Avg_FLLI_5km, 
              Avg_forest_cover_5km, 
              ## Anthropogenic vars 
              Avg_human_footprint_5km,Avg_hunting_accessibility_5km, Avg_nighttime_Lights_5km,
              Avg_Islam_Pop_5km, Avg_Christ_Pop_5km, Avg_Budd_Pop_5km, 
              ## degraded land cover
              Avg_OP_percent_5km, Avg_LM_percent_5km, Avg_LO_percent_5km, 
              Avg_RP_percent_5km,
              ## intact land cover
              Avg_LEF_percent_5km, Avg_LMEF_percent_5km, Avg_UMEF_percent_5km, 
              Avg_LDF_percent_5km, Avg_LMDF_percent_5km)

summary(meta)
## for the sake of our examples and exploring, change all NAs to zero 
meta[is.na(meta)] = 0

# ### Check how much correlation between each section of variables 
# ## Enviro
# check = select(meta, Avg_altitude_3km, Avg_ecoregion_intactness_3km, Avg_FLLI_3km,Avg_forest_cover_3km)
# cov.corr = cor(check[,sapply(check, is.numeric)])
# cov.corr= as.data.frame(cov.corr)
# cov.corr[abs(cov.corr) < abs(.6)] = "good" # stick w/ .6
# cov.corr
# ## FLII and forest cover are correlated --> remove forest cover! 
# meta$Avg_forest_cover_3km = NULL
# 
# ## anthro
# check = select(meta, Avg_human_footprint_3km,Avg_hunting_accessibility_3km, Avg_nighttime_Lights_3km,
#               Avg_Islam_Pop_3km, Avg_Christ_Pop_3km, Avg_Budd_Pop_3km)
# cov.corr = cor(check[,sapply(check, is.numeric)])
# cov.corr= as.data.frame(cov.corr)
# cov.corr[abs(cov.corr) < abs(.6)] = "good" # stick w/ .6
# cov.corr
# # nightime lights strongly correlated w/ HFP --> remove! 
# # also islam and buddist are correlated, but leave that in for now... 
# meta$Avg_nighttime_Lights_3km = NULL
# 
# ## degraded land
# check = select(meta, Avg_OP_percent_3km, Avg_LM_percent_3km, Avg_LO_percent_3km, 
#               Avg_RP_percent_3km)
# cov.corr = cor(check[,sapply(check, is.numeric)])
# cov.corr= as.data.frame(cov.corr)
# cov.corr[abs(cov.corr) < abs(.6)] = "good" # stick w/ .6
# cov.corr 
# ## all good! 
# 
# ## intact land
# check = select(meta, Avg_LEF_percent_3km, Avg_LMEF_percent_3km, Avg_UMEF_percent_3km, 
#               Avg_LDF_percent_3km, Avg_LMDF_percent_3km)
# cov.corr = cor(check[,sapply(check, is.numeric)])
# cov.corr= as.data.frame(cov.corr)
# cov.corr[abs(cov.corr) < abs(.6)] = "good" # stick w/ .6
# cov.corr
# ## all good! 
# rm(check, cov.corr)

### and do the same for caps
caps = select(caps, cell_id, survey_id, seq, Date, num_cams_active_at_date, Species, total_indiv_records, independent_events)
# and thin to only include relevant species
setdiff(traits$scientificNameStd, caps$Species) # all species in traits are present in caps  
caps = caps[caps$Species %in% traits$scientificNameStd, ]


## grab today's date 
day<-str_sub(Sys.Date(),-2)
month<-str_sub(Sys.Date(),-5,-4)
year<-str_sub(Sys.Date(),-10,-7)
date = paste(year,month,day, sep = "")

# ### Save all relevant files 
# saveRDS(keep, paste("data_GitHub/species_vector_", length(keep), "_species_", scale, "_scale_", date, ".RDS", sep = ""))
# 
# saveRDS(group_sp, paste("data_GitHub/group_living_", length(group_sp), "_species_", scale, "_scale_",  date, ".RDS", sep = ""))
# 
# write.csv(caps, paste("data_GitHub/clean_captures_to_make_UMFs_", scale, "_scale_",  date,".csv", sep = ""), row.names = F)
# 
# write.csv(meta, paste("data_GitHub/clean_metadata_to_make_UMFs_", scale, "_scale_",  date,".csv", sep = ""), row.names = F)

### Also save trait data so I dont have to remake it later
# write.csv(traits, paste("data_GitHub/clean_", nrow(traits), "_species_trait_data_", date, ".csv", sep = ""), row.names = F)

## clean up
rm(day, month, year, date)


### If starting from fresh, you can import the same data here-
keep = readRDS("data_GitHub/species_vector_44_species_5km_scale_20240913.RDS")
group_sp = readRDS("data_GitHub/group_living_9_species_5km_scale_20240913.RDS")
caps = read.csv("data_GitHub/clean_captures_to_make_UMFs_5km_scale_20240913.csv")
meta = read.csv("data_GitHub/clean_metadata_to_make_UMFs_5km_scale_20240913.csv")
traits = read.csv("data_Github/clean_44_species_trait_data_20240820.csv")


```

### HPC matrix and covariate creation code

The next section of analysis relies on the HPC, where we convert our covariates and metadata into a count history matrix with the proper associated observation- and site-level covariates. We use the HPC to iterate the process per species with more computational power which vastly speeds up the process. 

The separate R script is called: scripts/HPC_code/HPC_matrix_generator_SEA_TC.R.

To make that code run on the HPC, I use the SLURM script called: scripts/SLURM_code/SLURM_generate_matricies.txt

To avoid any confusion, I will not include that code here because it will not run locally.

### Generate data bundles

Import matrix data that was formatted on the HPC via loop. Objects were saved as .RDS files because they were list(), and I will save all lists in a singular bigger list. A list of lists if you will.

```{r import HPC data via loop, echo=FALSE}

### Before we import, just update the UMF file name to reflect the spatial scale

# # List of filenames (assuming they are in your current directory)
# files <- list.files("data_GitHub_UMFs", pattern = ".*_UMF_list_.*", full.names = T)
# 
# # Function to replace part of the filename
# new_filenames <- sub("(_UMF_list_)", paste("_", scale, "\\1", sep = ""), files)
# 
# # Rename files
# file.rename(files, new_filenames)
# 
# ## clean up
# rm(new_filenames)


## Store imported data here
umfs = list()

## list all files that should be imported 
files = list.files(here::here("data_GitHub_UMFs"))
files = files[!grepl("OLD", files)] # remove any old data 

for(i in 1:length(files)){
  
  #import the file
  f = files[i]
  path = paste("data_GitHub_UMFs/", f, sep = "")
  u = readRDS(here::here(path))
  
  #save it in a list 
  umfs[[i]] = u
  names(umfs)[i] = paste0(str_split(f, "_")[[1]][2:3], collapse = "_")
  
}
rm(i,u,f,path, files)

sort(names(umfs)) # looks good

## check if all species are present by importing old species vector
setdiff(keep, names(umfs)) # present and accounted for! 

# ### inspect present data
# head(umfs$Panthera_pardus$siteCovs) #nice
# head(umfs$Sus_barbatus$ObsCovs) #nice
# head(umfs$Ursus_thibetanus$y) #nice
# ## should be good to go! 

```

Now that the data is loaded in our environment, we will create co-abundance data bundles that will be ready to run on the HPC. Bundles will be generated for species pairs, where one species is explicitly assigned as a dominant species (predictor/independent variable), while the other is the subordinate species (response/dependent variable). Bundles will be generated based on spatial associations (i.e. the critters need to be present in at least one landscape together), and retained based on our preferred interactions we wish to examine. In particular, this means large carnivores acting as dominant over all other guilds, and large+small herbivores+ominivores acting as dominant to large carnivores.

```{r Ultra-Loop to generate data bundles based on guild-pairs, echo=FALSE }

## bind landscapes to captures to facilitate finding where critters are detected
l = distinct(select(meta, survey_id, Landscape))
caps = merge(caps, l, by = "survey_id")
rm(l)

#create a list to store results
bdata_list = list()

for(i in 1:length(umfs)){ # repeat for each first speices
  
  ## select a single species and thier data 
  # 1st species is the dominant! 
  sp_dom = names(umfs)[i]
  sp_dom_dat = umfs[[i]]
  
  ## grab the trophic guild of the dom species
  dom_tg = traits$TrophicGuild[traits$scientificNameStd == sp_dom]
  
  ## subset all other species
  all_else = names(umfs)
  all_else = all_else[all_else != sp_dom] # remove first species
  
  ### Thin all_else based upon the dominant species traits to avoid making too many models
  
  # if the dominant species is large+small herbivores+ominivores,
  if(dom_tg %in%  c("Small_Omnivore", "Large_Omnivore",
                    "Small_Herbivore","Large_Herbivore")){
    
    ## Then sub_species should only be large carnivores
    all_else = all_else[all_else %in%
                          traits$scientificNameStd[traits$TrophicGuild == "Large_Carnivore"]]

  } # end sub large carnivore conditional 
  
  # if the dominant species is a small carnivore,
  if(dom_tg == "Small_Carnivore"){
    
    ## skip it! 
    next
  }
  
  temp = list() # store results per sp1 here 
  for(l in 1:length(all_else)){ # repeat for each second species
    
    ## select a second species and their data
    # 2nd species is the subordinate! 
    sp_sub = all_else[l]
    sp_sub_dat = umfs[[sp_sub]]
    
    ##### Add a conditional to verify there are a combined total of 100 detections
    ##### shared across the landscapes where BOTH species were detected. 
    
    ## Grab landscapes present for each species
    dom_land = unique(caps$Landscape[caps$Species == sp_dom])
    sub_land = unique(caps$Landscape[caps$Species == sp_sub])
    ## and only save the ones where they intersect 
    lands = intersect(dom_land, sub_land)
    
    ## gather number of detections at shared landscapes for both species
    dom_det = nrow(caps[caps$Species == sp_dom &
                          caps$Landscape %in% lands,])
    sub_det = nrow(caps[caps$Species == sp_sub &
                          caps$Landscape %in% lands,])
    
    ## if there are less than 100 detections of both species from shared landscapes, 
    if(dom_det + sub_det < 100){
      # let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have less than 100 detections in shared landscapes! This pair was skipped."))
      # and skip the probelmatic pair
      next
      
    } # end 100 detection conditional 
    
    ## this conditional will also catch species pairs that dont spatially overlap, 
    ## where a conditonal does that below and isnt needed anymore, but leaving for saftey. 
    
    #
    ##
    ###
    #### Remove landscapes where neither species was detected and convert to matrix
    ###
    ##
    #
    
    ## First, determine which landscapes species are extirpated in
    exp = distinct(select(caps[caps$Species %in% c(sp_sub, sp_dom),], Landscape, Species))
    exp$presence = "yes"
    
    ## need to add landscapes where Species was NOT detected via loop 
    exp2 = list()
    # e = unique(exp$Species)[1]
    for(e in unique(exp$Species)){ # repeat for each species
      
      # select the relevant landscapes
      b = unique(caps$Landscape[caps$Species == e])
      #and make it a dataframe
      a = data.frame("Landscape" = unique(caps$Landscape[!caps$Landscape %in% b]))
      
      # save species name and make it a no
      a$Species = e
      a$presence = "no"
      
      exp2[[e]] = a
      
    }
    rm(a, b, e)
    
    # combine list into a df
    exp2 = do.call(rbind, exp2)
    rownames(exp2)= NULL
    
    # rbind presence w/ absences
    exp = rbind(exp, exp2)
    rm(exp2)
    
    ## grab all landscapes where neither species was detected
    non_det_land <- exp %>%
      group_by(Landscape) %>%
      filter(all(presence == "no"))
    
    ## extract all the sampling unit names from these landscapes
    non_det_SU = meta$cell_id[meta$Landscape %in% unique(non_det_land$Landscape)]
    
    ## extract both count matricies
    data.dom<- sp_dom_dat$y #dominant
    data.sub<- sp_sub_dat$y #subordinate
    
    ## and thin both to remove empty SUs
    data.dom = data.dom[!data.dom$SU %in% non_det_SU,]
    data.sub = data.sub[!data.sub$SU %in% non_det_SU,]
    
    ## make sure we match! 
    if(nrow(data.dom) != nrow(data.sub)){
      
      # if we dont, let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have different sized matricies! This pair was skipped."))
      # and skip the probelmatic one
      next
      
    } # end exact matrix size validation
    
    ### Now verify that both species are detected together in at least one landscape
    pres = unique(exp$Landscape[exp$presence == "yes" & exp$Species == unique(exp$Species)[1] &
                                  exp$Landscape %in% exp$Landscape[exp$presence == "yes" & 
                                                                     exp$Species ==  unique(exp$Species)[2]]])
    ## if there are no landscapes where both are detected,
    if(length(pres) == 0){
      
      ## let us know 
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "do not spatially overlap! This pair was skipped."))
      ## and skip the problematic one
      next
      
    }
    
    ## convert matrix dataframes back to matrix 
    dom_rows =data.dom[,1] #save sampling unit as rownames 
    data.dom=data.dom[,-1] #remove sampling unit col
    data.dom=as.matrix(data.dom) #turn into a matrix
    data.dom = apply(data.dom, 2, as.numeric) # make sure its numeric
    rownames(data.dom) = dom_rows # and re-instate the rownames 
    # repeat for spp 2
    sub_rows=data.sub[,1]
    data.sub=data.sub[,-1]
    data.sub=as.matrix(data.sub)
    data.sub = apply(data.sub, 2, as.numeric)
    rownames(data.sub) = sub_rows
    
    #
    ##
    ###
    #### Site-level covairtes
    ###
    ##
    #

    # Make sure we have the same number of sampling units
    if(dim(sp_dom_dat$siteCovs)[1] != dim(sp_sub_dat$siteCovs)[1]){
      
       # if we dont, let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have a different number of sampling units! This pair was skipped."))
      # and skip the probelmatic one
      next
     
    }else{
      
      # thin to match relevant sampling units
      covs = sp_dom_dat$siteCovs[sp_dom_dat$siteCovs$cell_id %in% rownames(data.dom),]
      
    } # end cov size conditional 

    
    ### Add conditional if no sites match each other
    if(dim(covs)[1] == 0){
      print(paste0(sp_dom, " & ", sp_sub, " dont spatially overlap, no combo created."))
      next
    }# end lack of spatial overlap bypass  
    
    
    #Extract Landscape for random effect, stored as a number
    d = data.frame("Landscape"= sort(unique(covs$Landscape)),
                   "land.num" = seq(from = 1, to = length(unique(covs$Landscape))))
    covs = merge(covs, d, by = "Landscape")

    #Extract years for random effect, stored as a number
    d = data.frame("year"= sort(unique(covs$year)),
                   "year.num" = seq(from = 1, to = length(unique(covs$year))))
    covs = merge(covs, d, by = "year")

     #Extract data source for random effect, stored as a number
    d = data.frame("source"= sort(unique(covs$source)),
                   "source.num" = seq(from = 1, to = length(unique(covs$source))))
    covs = merge(covs, d, by = "source")
    rm(d)
    
    
    ## Make sure covs match order of matrix
    covs = covs[order(match(covs$cell_id, rownames(data.dom))),]
    
    #verify
    # match(covs$cell_id, rownames(data.dom))
    
    #
    ##
    ###
    #### Observation-level covariates
    ###
    ##
    #
    
    # Make sure we have the same sampling units!
    if(length(setdiff(sp_dom_dat$ObsCovs$SU, sp_sub_dat$ObsCovs$SU)) == 0 &
       length(setdiff(sp_sub_dat$ObsCovs$SU, sp_dom_dat$ObsCovs$SU)) == 0){
      
      ## thin obs down to relevant SUs
      obs = sp_dom_dat$ObsCovs[sp_dom_dat$ObsCovs$SU %in% rownames(data.dom),]
      
      ## and convert to a matrix 
      rownames(obs)=obs[,1] #set rownames as sampling units
      obs=obs[,-1] #remove sampling unit col
      obs=as.matrix(obs) #turn into a matrix, but only keep one 
      
    }else{
      
      ## if obs SUs dont match, let us know! 
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                "have a different number of sampling units in ObsCovs! This pair was skipped."))
      # and skip the probelmatic one
      next
    }
    
    ## Ensure obs matches the order of matrix 
    obs = obs[order(match(rownames(obs), rownames(data.dom))),]
    # #verify
    # match(rownames(obs), rownames(data.dom))
    
    #
    ##
    ###
    ####
    ##### Prepare the informed ZIP parameters
    ####
    ###
    ##
    #
    
    ## Establish which sites have extirpated dominant species
    
    # first get max number of counts per row
    p.dom = apply(data.dom, 1, max, na.rm = TRUE)

    # subset for only sampling units with detections
    p.dom = names(p.dom[p.dom >= 1])
    
    ### Add conditional if no sites match each other
    if(length(p.dom) == 0){
      print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
      next
    }# end lack of spatial overlap bypass  
    
    # gather all sampling units in the Landscapes where species was detected
    d.dom = select(covs[covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.dom]), ], 
                   cell_id, Landscape) # only need cell_id and landscape
    
    # gather all sites in landscapes without ANY detections
    e.dom = select(covs[! covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.dom]), ], 
                   cell_id, Landscape)
    
    # this should match number of rows in matrix 
    if((length(e.dom$cell_id) + length(d.dom$cell_id)) != dim(data.dom)[1] &
      (length(e.dom$cell_id) + length(d.dom$cell_id)) != dim(data.sub)[1]){
      
      print(paste0("problem with ", sp_dom, " & ", sp_sub))
      
    } # end data check 
    
    
    ## Add if-else statement for dominant species not extirpated anywhere. 
    if(length(e.dom$cell_id) > 0){
      
      # combine, where 0 = extirpated, and 1 = present. 
      z.dom = rbind(data.frame("occu" = 0, "cell_id" = e.dom$cell_id),
                    data.frame("occu" = 1, "cell_id" = d.dom$cell_id))
      
    }else{
      
      z.dom = data.frame("occu" = 1, "cell_id" = d.dom$cell_id)
      
    } # end no extirpation conditional 
    
    # make sure its in the same order as the matrix
    z.dom = z.dom[order(match(z.dom$cell_id, rownames(data.dom))),]
    

    ## Establish which sites have extirpated subordianate species-
    
    # first get max number of counts per row
    p.sub = apply(data.sub, 1, max, na.rm = TRUE)
    
    # subset for only sites with detections
    p.sub = names(p.sub[p.sub >= 1])
    
    ### Add conditional if no sites match each other
    if(length(p.sub) == 0){
      print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
      next
    }# end lack of spatial overlap bypass  
    
    # gather all sampling units in the Landscapes where species was detected
    d.sub = select(covs[covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.sub]), ], 
                   cell_id, Landscape) # only need cell_id and landscape
    
    # gather all sites in landscapes without ANY detections
    e.sub = select(covs[! covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.sub]), ], 
                   cell_id, Landscape)
    
    
    # this should match number of rows in matrix 
    if((length(e.sub$cell_id) + length(d.sub$cell_id)) != dim(data.dom)[1] &
       (length(e.sub$cell_id) + length(d.sub$cell_id)) != dim(data.sub)[1]){
      
      print(paste0("problem with ", sp1, " & ", sp2))
      
    } # end data check 
   
    # combine, where 0 = extirpated, and 1 = present. 
    ## Add if-else statement for subordinate species not extirpated anywhere. 
    if(length(e.sub$cell_id) > 0){
      
      z.sub = rbind(data.frame("occu" = 0, "cell_id" = e.sub$cell_id),
                    data.frame("occu" = 1, "cell_id" = d.sub$cell_id))
      
    }else{
      
      z.sub = data.frame("occu" = 1, "cell_id" = d.sub$cell_id)
    } # end no extirpation conditional 
    
    # make sure its in the same order as the matrix
    z.sub = z.sub[order(match(z.sub$cell_id, rownames(data.sub))),]
    
    ## verify z.dom and z.sub have the same number of sampling units
    if(nrow(z.sub) != nrow(z.dom)){
      print(paste0(sp1, " & ", sp2, " have differing rows in their iZIP parameter! Inspect here!"))
      
      ## and end the loop, this is critical! 
      break
    }
    
    
    ####### BUNDLE all the data for the bayesian model 
    
    #these are the variable names used in the bayes mod! 
    bdata = list(y.dom = data.dom,                           # Count history matrix for dominant spp
                 y.sub = data.sub,                           # Count history matrix for subordinate spp
                 Z.dom = z.dom$occu,                         # Dominant presence/absence for all sites
                 Z.sub = z.sub$occu,                         # Subordinate presence/absence for all sites 
                 nsites = dim(data.dom)[1],                  # Number of sampling locations (i.e. rows in matrix)
                 nreps = dim(data.dom)[2],                   # Number of sampling occasions (i.e. cols in matrix)
                 # flii = covs$Avg_FLLI_3km,                   # Site covaraite 1
                 # hfp = covs$Avg_human_footprint_3km,         # Site covaraite 2
                 # elev = covs$Avg_altitude_3km,               # Site covaraite 3
                 flii = covs$Avg_FLLI_5km,                   # Site covaraite 1 @ 5km
                 hfp = covs$Avg_human_footprint_5km,         # Site covaraite 2 @ 5km
                 elev = covs$Avg_altitude_5km,               # Site covaraite 3 @ 5km
                 cams = obs,                                 # Observation covaraite 1
                 narea = length(unique(covs$land.num)),      # number of levels in landscape RE
                 area = covs$land.num,                       # Landscape random effect (as.num)
                 nsource = length(unique(covs$source)),      # number of levels in source RE 
                 source = covs$source.num,                   # Data source random effect (as.num)
                 nyear = length(unique(covs$year.num)),      # number of levels in year RE
                 year = covs$year.num)                       # year random effect (as.num)
    
    ## save the bundle 
    temp[[l]] = bdata
    names(temp)[l] = paste("SUB-", sp_sub, "~DOM-",sp_dom, sep = "" )
    
  } # end 2nd species
  
  ## remove null values in list for species combos that dont overlap 
  temp[sapply(temp, is.null)] <- NULL
  
  ## save all results for all species
  bdata_list[[i]] = temp
  
  
} # end 1st species
rm(i,l,sp_dom,sp_sub, sp_dom_dat, sp_sub_dat, temp, z.dom, z.sub, rm_SU, dom_tg,
   bdata, all_else, data.dom, data.sub, covs, exp, non_det_SU, non_det_land, pres,
   ob, obs_dom, obs_sub, o, obs, ob_name, sp, a, dom_det, sub_det, lands, env, check,
   d.dom,e.dom,p.dom,d.sub,e.sub,p.sub, dom_land, dom_rows, sub_land, sub_rows)



## flatten the list of lists into a single list
bdata_list = unlist(bdata_list, recursive = FALSE)
length(bdata_list) #260 combos after removing species that have less than 100 detections in shared landscapes and reducing nonsense species --> Nice! should allow faster testing as we reduce species. 
## interesting, 262 combos @ 5 km and also better iZIP?
names(bdata_list) # looks good! 

```

### Reorganize list for efficent analysis

After some preliminary testing, I realized that all co-abundance models should NOT run on the same amount of RAM... Most complete w/ 100 GB, but some need 250 or even 500 GB to complete. Rather than letting heavy models fail for a slow burn, I will reorganize the list into 100, 250, or 500 GB lists. This process will be achieved by inspecting the OE (output or error) files extracted from the HPC when each job runs. 

NOTE, these files are not hosted on GitHub due to storage limitations. You can complete the analysis as outlined in the materials and methods section by implementing the SLURM code called: SLURM_co-abundance_array_HALF_LONG_3Chain_100gb, 250gb, & 500GB.txt. 

```{r Use OE files to determine GB per list, echo=FALSE}

## First, examine how many OE files we have 
file = list.files("/Users/zachary_amir/Dropbox/Zach PhD/Ch3 Trophic release project/SEA_trophic_cascades_co-abundance/results/HALF_LONG_testing_Oct_2023/HALF_LONG_OE_files/")

## Then extract the unique job numbers 
jobs = unique(str_extract(file, pattern = "(?<=-|_)[0-9]+(?=-|_)")) #Thx chat GPT for the RegEx
unique(jobs) # good, should only have three. 

## and save job numbers, GB,  and species pairs  
jobs = data.frame("number" = jobs, 
                  "RAM" = c("100GB", "250GB", "500GB"), # Based on personal knowledge of job submissions
                  "Species_Pair" = as.character(NA))
jobs
## End goal is to fill this species_pair column

# Loop to the rescue as usual!

# Only need to work with .out files, all relevant info is here. 
out = file[grepl(".out", file)]

for(i in 1:length(bdata_list)){
  
  # grab one sp pair
  sp = names(bdata_list)[i]

  ## import each out file to find the correct one 
  for(f in 1:length(out)){
    
    ## import an out file 
    o = read.delim(paste("/Users/zachary_amir/Dropbox/Zach PhD/Ch3 Trophic release project/SEA_trophic_cascades_co-abundance/results/HALF_LONG_testing_Oct_2023/HALF_LONG_OE_files/", 
                         out[f], sep = ""), sep = "\t")
    
    ## grab the job number too 
    job_num = str_extract(out[f], pattern = "(?<=-|_)[0-9]+(?=-|_)")
    
    ## and the relevant GB that matches the job num
    gb = unique(jobs$RAM[jobs$number == job_num])
    
    ## grab the 8th word from the first line of the out file to grab the species pair 
    sp_out = strsplit(o[1,], "\\s+")[[1]]
    sp_out = sp_out[8]
    
    ## Verify the out file has the correct species pair by skipping the iteration if it is wrong! 
    if(sp != sp_out){next}
    
    ## now verify the out file actually completed the job,
    # by checking for "Finished running co-abundance model" on line 9
    if(grepl("Finished running co-abundance model", o[9,])){
      
      # if it is a match, make a new row to add to the jobs DF 
      new = data.frame("number" = job_num, 
                       "RAM" = gb, 
                       "Species_Pair" = sp)
      
      # and combine with jobs 
      jobs = rbind(jobs, new)
      
      
    } # end finished mod conditional 
    
  } # end per out file 
    
} # end per sp_pair
rm(o,f,i,gb,file,job_num,out,sp,sp_out,new)

## remove NA species pairs from making the df
jobs_final = jobs[!is.na(jobs$Species_Pair),]

## verify all jobs are accounted for 
setdiff(names(bdata_list), jobs_final$Species_Pair) #all here! 

## inspect to make sure no mods have multiple job numbers
inspect = ddply(jobs_final, .(Species_Pair), summarize,
                num_job = length(unique(number)))
sp = inspect$Species_Pair[inspect$num_job > 1] # 5 mods were completed twice! 
## verify
jobs_final[jobs_final$Species_Pair %in% sp,] #Only want the smaller job number

# remove duplicated by taking the min number 
jobs_final = jobs_final %>%
  group_by(Species_Pair) %>%
  filter(number == min(number))

## verify all species are present still! 
setdiff(names(bdata_list), jobs_final$Species_Pair) #all here! 

## make sure there is no NA
anyNA(jobs_final) # MUST BE F

### Begin to add preferred prey column

## First, split apart species pair into dom and sub species 
jobs_final <- within(jobs_final, {
  sub_sp <- sapply(strsplit(Species_Pair, "~"), function(x) x[1])
  dom_sp <- sapply(strsplit(Species_Pair, "~"), function(x) x[2])
})
jobs_final$dom_sp = gsub("DOM-", "", jobs_final$dom_sp)
jobs_final$sub_sp = gsub("SUB-", "", jobs_final$sub_sp)

## inspect
setdiff(keep, jobs_final$dom_sp) # makes sense, small carnivores are not dominant! 
setdiff(keep, jobs_final$sub_sp) # all here! 
length(unique(jobs_final$sub_sp)) == length(keep) # MUST BE TRUE

## now add a new column if the model is preferred or not 
jobs_final$preference = "community" # not preferred
# top-downs
jobs_final$preference[jobs_final$dom_sp == "Panthera_tigris" &
                        jobs_final$sub_sp %in% 
                        traits$scientificNameStd[traits$tiger_pref == "Yes"]] = "preferred"
jobs_final$preference[jobs_final$dom_sp == "Panthera_pardus" &
                        jobs_final$sub_sp %in% 
                        traits$scientificNameStd[traits$leopard_pref == "Yes"]] = "preferred"
jobs_final$preference[jobs_final$dom_sp == "Cuon_alpinus" &
                        jobs_final$sub_sp %in% 
                        traits$scientificNameStd[traits$dhole_pref == "Yes"]] = "preferred"
jobs_final$preference[jobs_final$dom_sp == "Neofelis_genus" &
                        jobs_final$sub_sp %in% 
                        traits$scientificNameStd[traits$CL_pref == "Yes"]] = "preferred"
# bottom-ups
jobs_final$preference[jobs_final$sub_sp == "Panthera_tigris" &
                        jobs_final$dom_sp %in% 
                        traits$scientificNameStd[traits$tiger_pref == "Yes"]] = "preferred"
jobs_final$preference[jobs_final$sub_sp == "Panthera_pardus" &
                        jobs_final$dom_sp %in% 
                        traits$scientificNameStd[traits$leopard_pref == "Yes"]] = "preferred"
jobs_final$preference[jobs_final$sub_sp == "Cuon_alpinus" &
                        jobs_final$dom_sp %in% 
                        traits$scientificNameStd[traits$dhole_pref == "Yes"]] = "preferred"
jobs_final$preference[jobs_final$sub_sp == "Neofelis_genus" &
                        jobs_final$dom_sp %in% 
                        traits$scientificNameStd[traits$CL_pref == "Yes"]] = "preferred"
## how many do we have?
table(jobs_final$preference) # 66 preferred, 196 not. 66 + 194 = 262, all accounted for! 

## clean up
rm(inspect, jobs, sp)

```


### Save the final list of bundled data

Now that all species pairs are established, the data is properly formatted, and everything is clean, its time to save the data! We will send it to the HPC to run on the R script called: data/HPC_code/HPC_co-abundance_model_final.R

Alternatively, I could save the list as a single object rather than splitting it up. 


```{r save bundled data!, echo=FALSE}

## grab today's date
day<-str_sub(Sys.Date(),-2)
month<-str_sub(Sys.Date(),-5,-4)
year<-str_sub(Sys.Date(),-10,-7)
date = paste(year,month,day, sep = "")
rm(day,month,year)


# # Save data per preference AND GB requirements!
# # pref 100 GB
# save = bdata_list[names(bdata_list) %in%
#                     jobs_final$Species_Pair[jobs_final$RAM == "100GB"] &
#                     names(bdata_list) %in%
#                             jobs_final$Species_Pair[jobs_final$preference == "preferred"]]
# saveRDS(save, here::here(paste("data_GitHub_CoA_bundles/Bundled_data_for_Bayes_co-abundance_mods_preferred_100GB_", length(save), "_species_pairs_", date, ".RDS", sep = "")))
# # pref 250 GB
# save = bdata_list[names(bdata_list) %in%
#                     jobs_final$Species_Pair[jobs_final$RAM == "250GB"] &
#                     names(bdata_list) %in%
#                             jobs_final$Species_Pair[jobs_final$preference == "preferred"]]
# saveRDS(save, here::here(paste("data_GitHub_CoA_bundles/Bundled_data_for_Bayes_co-abundance_mods_preferred_250GB_", length(save), "_species_pairs_", date, ".RDS", sep = "")))
# # pref 500 GB
# save = bdata_list[names(bdata_list) %in%
#                     jobs_final$Species_Pair[jobs_final$RAM == "500GB"] &
#                     names(bdata_list) %in%
#                             jobs_final$Species_Pair[jobs_final$preference == "preferred"]]
# saveRDS(save, here::here(paste("data_GitHub_CoA_bundles/Bundled_data_for_Bayes_co-abundance_mods_preferred_500GB_", length(save), "_species_pairs_", date, ".RDS", sep = "")))
# # NON-pref 100 GB
# save = bdata_list[names(bdata_list) %in%
#                     jobs_final$Species_Pair[jobs_final$RAM == "100GB"] &
#                     names(bdata_list) %in%
#                             jobs_final$Species_Pair[jobs_final$preference == "community"]]
# saveRDS(save, here::here(paste("data_GitHub_CoA_bundles/Bundled_data_for_Bayes_co-abundance_mods_community_100GB_", length(save), "_species_pairs_", date, ".RDS", sep = "")))
# # NON-pref 250 GB
# save = bdata_list[names(bdata_list) %in%
#                     jobs_final$Species_Pair[jobs_final$RAM == "250GB"] &
#                     names(bdata_list) %in%
#                             jobs_final$Species_Pair[jobs_final$preference == "community"]]
# saveRDS(save, here::here(paste("data_GitHub_CoA_bundles/Bundled_data_for_Bayes_co-abundance_mods_community_250GB_", length(save), "_species_pairs_", date, ".RDS", sep = "")))
# # NON-pref 500 GB
# save = bdata_list[names(bdata_list) %in%
#                     jobs_final$Species_Pair[jobs_final$RAM == "500GB"] &
#                     names(bdata_list) %in%
#                             jobs_final$Species_Pair[jobs_final$preference == "community"]]
# saveRDS(save, here::here(paste("data_GitHub_CoA_bundles/Bundled_data_for_Bayes_co-abundance_mods_community_500GB_", length(save), "_species_pairs_", date, ".RDS", sep = "")))


### However, if you are just going for testing (MIDDLE settings) with preferred models, 
## we can save all preferred models as one list. 
sp_pairs = names(bdata_list)[names(bdata_list) %in% jobs_final$Species_Pair[jobs_final$preference == "preferred"]]
save = bdata_list[names(bdata_list) %in% sp_pairs]

# saveRDS(save, here::here(paste("data_GitHub_CoA_bundles/Bundled_data_for_Bayes_co-abundance_mods_preferred_", length(save), "_species_pairs_", date, ".RDS", sep = "")))
rm(sp_pairs, jobs_final, date)


```


### Generate counterfactual data bundles 

After submitting the manuscript and analysis to Science and receving reviews from 3 expert reviewers, 2 reviewers revealed an appetite for causation that was lacking from the current analysis. Moreover, when presenting my findings at conferences and to collegues, similar sentiments were voiced. Structural Equation Models (SEM) was a logical starting place, but the limitations of the data (ie lacking key variables like hunting pressure or resource availability) and inability to include a count history matrix as the response variable (and therefore propagate uncertainty) left me unsatisfied. 

Instead, the new goal here is to examine the counter-factual to understand if prey abundance increases would not have happened if not for large carnivore decline/extirpation (top-down test), AND if predator abundance increases would not have happened if not for prey abundance increases (bottom-up test). Additionally, we can run more variations of the same co-abundance model by sub-setting the co-occuring species data into locations that match in their measured covariates to help control for unmeasured variables. All counter-factual testing will occurr only with the 33 preferred prey species-pairs. 

This approach builds towards the top of the 'ladder of causation' proposed by Judea Pearl's Book of Why. The first rung, association/observation, can be represented simply with a non-directional association between two species (e.g. correlation between RAI of two species). The second run, intervention, typically involves a manipulation, but can also be expressed through causal diagrams, where one variable directionally affects another variable, such as predators suppress prey or prey bolster predators. Finally, the third rung, counter-factual, forces the researcher to imagine an alternative reality where causal forces are reversed or eliminated, such as testing if prey abundance increases in a world with no large carnivore extirpation. 

I will be running several counter-factual tests, listed here:
1) Would prey abundance increase in landscapes where large carnivores remain abundant (top-down test), AND would predator abundance increase in landscapes where preferred prey remain scarce (bottom-up test). This isolates the effect of declines while being present rather than species extirpations, and is tested by selecting landscapes where predators remain common (upper 3rd quartile of site-level abundance) and prey remain scarce (bottom 1st quartile of site-level abundance). 
2) What if prey abundance increases would not have happened if not for large carnivore extirpation (top-down test), AND if predator abundance decreases would not have happened if not for prey abundance extirpations (bottom-up test). This isolates the effect of species extirpations rather than declines, and is tested by only selecting landscapes where species co-occur w/ sufficient detections.
3) Would we observe trophic release in prey or bottom-up bolstering in predators when controlling for landscape variation that may be driving observed trends. This is tested by implementing good and clear site-matching, but in this test, just the Thai Eastern Forest Complex (for now). 

4:6) Would we observe trophic release in prey and bottom-up bolstering in predators when controlling for our observed site covariates? This is tested by running three different counterfactual tests where we let one of the three covariates (HFP, FLII, elev) vary while holding the other two around their mean (mean plus and minus 1/2 SD). 

```{r prepare data for counterfactual test 1}

##### Null abundance estiamtes were creareated for all species to test SEMs
#### in this R script: HPC_N-mixture_mod_abundance_generation_for_SEM.R
### Estimates were made from pcount(num_active_cams + (1|source) ~ (1|Landscape) + (1|year), umf) 
## using count data from landscapes where species was detected.  

# import null abundance
abund = read.csv("/Users/zachary_amir/Dropbox/SEM of wildlife communities/MetaSEMs/SEM_estimates_abundamce_per_SU_dataframe_20240524.csv")
head(abund) # currentlty long format, but should convert to wide and merge w/ meta 
abund$X = NULL # damn row names

## do this via loop, cant figure out pivot
abund_wide = distinct(data.frame(cell_id = abund$cell_id, Landscape = abund$Landscape))

for(i in 1:length(unique(abund$species))){
  
  # grab data from one species 
  sp_dat = abund[abund$species == unique(abund$species)[i], ]
  
  # Extract the relevant info abundance
  sp_abund = select(sp_dat, cell_id, Landscape, abundance)
  
  ## rename last col 
  names(sp_abund)[3] = paste(unique(abund$species)[i], "abundance", sep = "_")
  
  # and merge w/ abund_wide
  abund_wide = merge(abund_wide, sp_abund, by = c("cell_id", "Landscape"))
  
} # end per species
rm(sp_dat, sp_abund, i)

## inspect
head(abund_wide)
anyNA(abund_wide) # MUST BE F, will be zero where species does not occurr.

## Now that we have calculated abundance at 3 km and testing new models @ 5km 



## prep to merge w/ the meta 
names(abund_wide)[1] = "cell_id" # rename for easier merging
setdiff(abund_wide$cell_id, meta$cell_id)
setdiff(meta$cell_id, abund_wide$cell_id) # all good
setdiff(abund_wide$Landscape, meta$Landscape)
setdiff(meta$Landscape, abund_wide$Landscape) # all good!
# merge
meta_abund = merge(meta, abund_wide, by= c("cell_id", "Landscape"))
# quick exploration below 
{
  ## Now that were all here, just check a few quick correlations/tests
  cor(meta_abund$Panthera_tigris_abundance, meta_abund$Muntiacus_genus_abundance) # weak and negative
  t.test(meta_abund$Panthera_tigris_abundance, meta_abund$Muntiacus_genus_abundance) # significant and negative
  
  cor(meta_abund$Panthera_pardus_abundance, meta_abund$Macaca_nemestrina_abundance) # weak and negative
  t.test(meta_abund$Panthera_pardus_abundance, meta_abund$Macaca_nemestrina_abundance) # significant and negative
  
  cor(meta_abund$Panthera_pardus_abundance, meta_abund$Macaca_fascicularis_abundance) # strong(ish) and positive 
  t.test(meta_abund$Panthera_pardus_abundance, meta_abund$Macaca_fascicularis_abundance) # significant and negative ???
  
  cor(meta_abund$Panthera_tigris_abundance, meta_abund$Rusa_unicolor_abundance) # weakish and positive 
  t.test(meta_abund$Panthera_tigris_abundance, meta_abund$Rusa_unicolor_abundance) # significant and negative ???
  ### Ignore these, theyre silly
}


## Want to run two tests, predator abundant and prey scarce
cols = names(meta_abund)[grepl("abundance", names(meta_abund))] # grab all species abundance
predators = c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus") # specify predators
res = data.frame("test" = NA, "Landscape" = NA,  "dom_sp" = NA)
res = res[0,] # empty DF for results

# run the loop
for(l in 1:length(cols)){
  
  # choose a species
  c = cols[l]
  sp = paste(strsplit(c, "_")[[1]][1:2], collapse = "_")
  
  # select relevant data 
  m = select(meta_abund, cell_id, Landscape, all_of(c))
  
  # rename col for ddply
  names(m)[3] = "abundance"
  
  # run the ddply 
  d = ddply(m, .(Landscape), summarize,
            mean_abund = mean(abundance[abundance > 0], na.rm = T))
  
  
  ## fill in res
  if(sp %in% predators){
    
    # grab 3rd quartile of abundance
    a = summary(d$mean_abund)[5]
    
    # and the landscapes where they are equal or above that abundance
    lands = paste0(na.omit(unique(d$Landscape[d$mean_abund >= a])), collapse = " & ")
    
    # and save this info! 
    res[l,"test"] = "predator_abundant" # specify the test (top-down)
    res[l,"Landscape"] = lands
    res[l, "dom_sp"] = sp
    
  }else{
    
    # grab 1st quartile of abundance
    a = summary(d$mean_abund)[2]
    
    # and the landscapes where they are equal or less than abundance
    lands = paste0(na.omit(unique(d$Landscape[d$mean_abund <= a])), collapse = " & ")
    
    # and save this info! 
    res[l,"test"] = "prey_scarce" # specify the test (bottom-up)
    res[l,"Landscape"] = lands
    res[l, "dom_sp"] = sp
    
  } # end test conditional 
    
} # end per species
rm(l,c,sp,m,d,lands,a,cols, predators)

## should have 4 predator abundant and the rest prey_scarce
table(res$test) # good! 

```

```{r Ultra loop to generate data bundles for counterfactual test 1} 

## remember, ony making bundles for preferred species pairs here, not the full monty

#create a list to store results
counterf1_bdata_list = list()

for(i in 1:length(umfs)){ # repeat for each first speices
  
  ## select a single species and thier data 
  # 1st species is the dominant! 
  sp_dom = names(umfs)[i]
  sp_dom_dat = umfs[[i]]
  
  ## subset all other species
  all_else = names(umfs)
  all_else = all_else[all_else != sp_dom] # remove first species
  
  ## if the dominant species is a prey,
  if(sp_dom %in% res$dom_sp[res$test == "prey_scarce"]){
    
    ## Then sub_species should only be large carnivores
    all_else = all_else[all_else %in%
                          traits$scientificNameStd[traits$TrophicGuild == "Large_Carnivore"]]
    ## and grab the relevant landscapes, saved as a vector of multiple values
    lands = res$Landscape[res$test == "prey_scarce" & res$dom_sp == sp_dom]
    lands = strsplit(lands, " & ")[[1]]
    
  } # end bottom-up counterfactual condition
  
  ## if the dominant species is a predator 
  if(sp_dom %in% res$dom_sp[res$test == "predator_abundant"]){
    
    ## then sub_species should only be preferred prey
    if(sp_dom == "Cuon_alpinus"){ 
      all_else = all_else[all_else %in%
                            traits$scientificNameStd[traits$dhole_pref == "Yes"]]
      ## and grab the relevant landscapes, saved as a vector of multiple values
      lands = res$Landscape[res$test == "predator_abundant" & res$dom_sp == sp_dom]
      lands = strsplit(lands, " & ")[[1]]
    }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Panthera_tigris"){ 
      all_else = all_else[all_else %in%
                            traits$scientificNameStd[traits$tiger_pref == "Yes"]]
      ## and grab the relevant landscapes, saved as a vector of multiple values
      lands = res$Landscape[res$test == "predator_abundant" & res$dom_sp == sp_dom]
      lands = strsplit(lands, " & ")[[1]]
      }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Panthera_pardus"){ 
      all_else = all_else[all_else %in%
                            traits$scientificNameStd[traits$leopard_pref == "Yes"]]
      
      ## and grab the relevant landscapes, saved as a vector of multiple values
      lands = res$Landscape[res$test == "predator_abundant" & res$dom_sp == sp_dom]
      lands = strsplit(lands, " & ")[[1]]
      }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Neofelis_genus"){ 
      all_else = all_else[all_else %in%
                            traits$scientificNameStd[traits$CL_pref == "Yes"]]
      ## and grab the relevant landscapes, saved as a vector of multiple values
      lands = res$Landscape[res$test == "predator_abundant" & res$dom_sp == sp_dom]
      lands = strsplit(lands, " & ")[[1]]}
    
  } # end top-down counterfactial condition 
  
  temp = list() # store results per sp1 here 
  for(l in 1:length(all_else)){ # repeat for each second species
    
    ## select a second species and their data
    # 2nd species is the subordinate! 
    sp_sub = all_else[l]
    sp_sub_dat = umfs[[sp_sub]]
    
    ### Thin both species data to only include sites from key landscapes 
    # frist grab relevant SUs
    su = meta$cell_id[meta$Landscape %in% lands]
    # then thin
    dom_y = sp_dom_dat$y;dom_obs = sp_dom_dat$ObsCovs;dom_site = sp_dom_dat$siteCovs
    dom_y = dom_y[dom_y$SU %in% su, ];dom_obs = dom_obs[dom_obs$SU %in% su,];dom_site = dom_site[dom_site$cell_id %in% su,]
    # and update
    sp_dom_dat$y = dom_y;sp_dom_dat$ObsCovs = dom_obs;sp_dom_dat$siteCovs = dom_site
    ### repeat for subordinate
    # thin first 
    sub_y = sp_sub_dat$y;sub_obs = sp_sub_dat$ObsCovs;sub_site = sp_sub_dat$siteCovs
    sub_y = sub_y[sub_y$SU %in% su, ];sub_obs = sub_obs[sub_obs$SU %in% su,];sub_site = sub_site[sub_site$cell_id %in% su,]
    # and update
    sp_sub_dat$y = sub_y;sp_sub_dat$ObsCovs = sub_obs;sp_sub_dat$siteCovs = sub_site
    rm(sub_y,sub_obs,sub_site, dom_y, dom_obs,dom_site)
    
    ##### Add a conditional to verify there are a combined total of 100 detections
    ##### shared across the landscapes where BOTH species were detected. 
    
    ## Grab landscapes present for each species
    dom_land = unique(caps$Landscape[caps$Species == sp_dom & caps$Landscape %in% lands])
    sub_land = unique(caps$Landscape[caps$Species == sp_sub & caps$Landscape %in% lands])
    ## and only save the ones where they intersect 
    lands = intersect(dom_land, sub_land)
    rm(dom_land, sub_land)
    
    ## gather number of detections at shared landscapes for both species
    dom_det = nrow(caps[caps$Species == sp_dom &
                          caps$Landscape %in% lands,])
    sub_det = nrow(caps[caps$Species == sp_sub &
                          caps$Landscape %in% lands,])
    
    ## if there are less than 100 detections of both species from shared landscapes, 
    if(dom_det + sub_det < 100){
      # let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have less than 100 detections in shared landscapes! This pair was skipped."))
      # and skip the probelmatic pair
      next
      
    } # end 100 detection conditional 
    rm(dom_det, sub_det)
    ## this conditional will also catch species pairs that dont spatially overlap, 
    ## where a conditonal does that below and isnt needed anymore, but leaving for saftey. 
    
    #
    ##
    ###
    #### Remove landscapes where neither species was detected and convert to matrix
    ###
    ##
    #
    
    ## First, determine which landscapes species are extirpated in
    exp = distinct(select(caps[caps$Species %in% c(sp_sub, sp_dom) & caps$Landscape %in% lands,], Landscape, Species))
    exp$presence = "yes"
    
    ## need to add landscapes where Species was NOT detected via loop 
    exp2 = list()
    # e = unique(exp$Species)[1]
    for(e in unique(exp$Species)){ # repeat for each species
      
      # select the relevant landscapes
      b = unique(caps$Landscape[caps$Species == e & caps$Landscape %in% lands])
      #and make it a dataframe
      a = data.frame("Landscape" = unique(caps$Landscape[!caps$Landscape %in% b & caps$Landscape %in% lands]))
      
      ## if there are no extirpations
      if(nrow(a) == 0){
        next
      }
      
      # save species name and make it a no
      a$Species = e
      a$presence = "no"
      
      exp2[[e]] = a
      
    }
    rm(a, b, e)
    
    # combine list into a df
    exp2 = do.call(rbind, exp2)
    rownames(exp2)= NULL
    
    # rbind presence w/ absences
    exp = rbind(exp, exp2)
    rm(exp2)
    
    ## grab all landscapes where neither species was detected
    non_det_land <- exp %>%
      group_by(Landscape) %>%
      filter(all(presence == "no"))
    
    ## extract all the sampling unit names from these landscapes
    non_det_SU = meta$cell_id[meta$Landscape %in% unique(non_det_land$Landscape)]
    
    ## extract both count matricies
    data.dom<- sp_dom_dat$y #dominant
    data.sub<- sp_sub_dat$y #subordinate
    
    ## and thin both to remove empty SUs
    data.dom = data.dom[!data.dom$SU %in% non_det_SU,]
    data.sub = data.sub[!data.sub$SU %in% non_det_SU,]
    
    ## make sure we match! 
    if(nrow(data.dom) != nrow(data.sub)){
      
      # if we dont, let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have different sized matricies! This pair was skipped."))
      # and skip the probelmatic one
      next
      
    } # end exact matrix size validation
    
    ### Now verify that both species are detected together in at least one landscape
    pres = unique(exp$Landscape[exp$presence == "yes" & exp$Species == unique(exp$Species)[1] &
                                  exp$Landscape %in% exp$Landscape[exp$presence == "yes" & 
                                                                     exp$Species ==  unique(exp$Species)[2]]])
    ## if there are no landscapes where both are detected,
    if(length(pres) == 0){
      
      ## let us know 
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "do not spatially overlap! This pair was skipped."))
      ## and skip the problematic one
      next
      
    }

    
    ## convert matrix dataframes back to matrix 
    rownames(data.dom)=data.dom[,1] #set rownames as sampling units
    data.dom=data.dom[,-1] #remove sampling unit col
    data.dom=as.matrix(data.dom) #turn into a matrix
    # repeat for spp 2
    rownames(data.sub)=data.sub[,1]
    data.sub=data.sub[,-1]
    data.sub=as.matrix(data.sub)
    
    #
    ##
    ###
    #### Site-level covairtes
    ###
    ##
    #

    # Make sure we have the same number of sampling units
    if(dim(sp_dom_dat$siteCovs)[1] != dim(sp_sub_dat$siteCovs)[1]){
      
       # if we dont, let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have a different number of sampling units! This pair was skipped."))
      # and skip the probelmatic one
      next
     
    }else{
      
      # thin to match relevant sampling units
      covs = sp_dom_dat$siteCovs[sp_dom_dat$siteCovs$cell_id %in% rownames(data.dom),]
      
    } # end cov size conditional 

    
    ### Add conditional if no sites match each other
    if(dim(covs)[1] == 0){
      print(paste0(sp_dom, " & ", sp_sub, " dont spatially overlap, no combo created."))
      next
    }# end lack of spatial overlap bypass  
    
    
    #Extract Landscape for random effect, stored as a number
    d = data.frame("Landscape"= sort(unique(covs$Landscape)),
                   "land.num" = seq(from = 1, to = length(unique(covs$Landscape))))
    covs = merge(covs, d, by = "Landscape")

    #Extract years for random effect, stored as a number
    d = data.frame("year"= sort(unique(covs$year)),
                   "year.num" = seq(from = 1, to = length(unique(covs$year))))
    covs = merge(covs, d, by = "year")

     #Extract data source for random effect, stored as a number
    d = data.frame("source"= sort(unique(covs$source)),
                   "source.num" = seq(from = 1, to = length(unique(covs$source))))
    covs = merge(covs, d, by = "source")
    rm(d)
    
    
    ## Make sure covs match order of matrix
    covs = covs[order(match(covs$cell_id, rownames(data.dom))),]
    
    #verify
    # match(covs$cell_id, rownames(data.dom))
    
    #
    ##
    ###
    #### Observation-level covariates
    ###
    ##
    #
    
    # Make sure we have the same sampling units!
    if(length(setdiff(sp_dom_dat$ObsCovs$SU, sp_sub_dat$ObsCovs$SU)) == 0 &
       length(setdiff(sp_sub_dat$ObsCovs$SU, sp_dom_dat$ObsCovs$SU)) == 0){
      
      ## thin obs down to relevant SUs
      obs = sp_dom_dat$ObsCovs[sp_dom_dat$ObsCovs$SU %in% rownames(data.dom),]
      
      ## and convert to a matrix 
      rownames(obs)=obs[,1] #set rownames as sampling units
      obs=obs[,-1] #remove sampling unit col
      obs=as.matrix(obs) #turn into a matrix, but only keep one 
      
    }else{
      
      ## if obs SUs dont match, let us know! 
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                "have a different number of sampling units in ObsCovs! This pair was skipped."))
      # and skip the probelmatic one
      next
    }
    
    ## Ensure obs matches the order of matrix 
    obs = obs[order(match(rownames(obs), rownames(data.dom))),]
    # #verify
    # match(rownames(obs), rownames(data.dom))
    
    #
    ##
    ###
    ####
    ##### Prepare the informed ZIP parameters
    ####
    ###
    ##
    #
    
    ## Establish which sites have extirpated dominant species
    
    # first get max number of counts per row
    p.dom = apply(data.dom, 1, max, na.rm = TRUE)

    # subset for only sampling units with detections
    p.dom = names(p.dom[p.dom >= 1])
    
    ### Add conditional if no sites match each other
    if(length(p.dom) == 0){
      print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
      next
    }# end lack of spatial overlap bypass  
    
    # gather all sampling units in the Landscapes where species was detected
    d.dom = select(covs[covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.dom]), ], 
                   cell_id, Landscape) # only need cell_id and landscape
    
    # gather all sites in landscapes without ANY detections
    e.dom = select(covs[! covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.dom]), ], 
                   cell_id, Landscape)
    
    # this should match number of rows in matrix 
    if((length(e.dom$cell_id) + length(d.dom$cell_id)) != dim(data.dom)[1] &
      (length(e.dom$cell_id) + length(d.dom$cell_id)) != dim(data.sub)[1]){
      
      print(paste0("problem with ", sp_dom, " & ", sp_sub))
      
    } # end data check 
    
    
    ## Add if-else statement for dominant species not extirpated anywhere. 
    if(length(e.dom$cell_id) > 0){
      
      # combine, where 0 = extirpated, and 1 = present. 
      z.dom = rbind(data.frame("occu" = 0, "cell_id" = e.dom$cell_id),
                    data.frame("occu" = 1, "cell_id" = d.dom$cell_id))
      
    }else{
      
      z.dom = data.frame("occu" = 1, "cell_id" = d.dom$cell_id)
      
    } # end no extirpation conditional 
    
    # make sure its in the same order as the matrix
    z.dom = z.dom[order(match(z.dom$cell_id, rownames(data.dom))),]
    

    ## Establish which sites have extirpated subordianate species-
    
    # first get max number of counts per row
    p.sub = apply(data.sub, 1, max, na.rm = TRUE)
    
    # subset for only sites with detections
    p.sub = names(p.sub[p.sub >= 1])
    
    ### Add conditional if no sites match each other
    if(length(p.sub) == 0){
      print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
      next
    }# end lack of spatial overlap bypass  
    
    # gather all sampling units in the Landscapes where species was detected
    d.sub = select(covs[covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.sub]), ], 
                   cell_id, Landscape) # only need cell_id and landscape
    
    # gather all sites in landscapes without ANY detections
    e.sub = select(covs[! covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.sub]), ], 
                   cell_id, Landscape)
    
    
    # this should match number of rows in matrix 
    if((length(e.sub$cell_id) + length(d.sub$cell_id)) != dim(data.dom)[1] &
       (length(e.sub$cell_id) + length(d.sub$cell_id)) != dim(data.sub)[1]){
      
      print(paste0("problem with ", sp1, " & ", sp2))
      
    } # end data check 
   
    # combine, where 0 = extirpated, and 1 = present. 
    ## Add if-else statement for subordinate species not extirpated anywhere. 
    if(length(e.sub$cell_id) > 0){
      
      z.sub = rbind(data.frame("occu" = 0, "cell_id" = e.sub$cell_id),
                    data.frame("occu" = 1, "cell_id" = d.sub$cell_id))
      
    }else{
      
      z.sub = data.frame("occu" = 1, "cell_id" = d.sub$cell_id)
    } # end no extirpation conditional 
    
    # make sure its in the same order as the matrix
    z.sub = z.sub[order(match(z.sub$cell_id, rownames(data.sub))),]
    
    ## verify z.dom and z.sub have the same number of sampling units
    if(nrow(z.sub) != nrow(z.dom)){
      print(paste0(sp1, " & ", sp2, " have differing rows in their iZIP parameter! Inspect here!"))
      
      ## and end the loop, this is critical! 
      break
    }
    
    ####### BUNDLE all the data for the bayesian model 
    
    #these are the variable names used in the bayes mod! 
    bdata = list(y.dom = data.dom,                           # Count history matrix for dominant spp
                 y.sub = data.sub,                           # Count history matrix for subordinate spp
                 Z.dom = z.dom$occu,                         # Dominant presence/absence for all sites
                 Z.sub = z.sub$occu,                         # Subordinate presence/absence for all sites 
                 nsites = dim(data.dom)[1],                  # Number of sampling locations (i.e. rows in matrix)
                 nreps = dim(data.dom)[2],                   # Number of sampling occasions (i.e. cols in matrix)
                 flii = covs$Avg_FLLI_3km,                   # Site covaraite 1
                 hfp = covs$Avg_human_footprint_3km,         # Site covaraite 2
                 elev = covs$Avg_altitude_3km,               # Site covaraite 3
                 cams = obs,                                 # Observation covaraite 1
                 narea = length(unique(covs$land.num)),      # number of levels in landscape RE
                 area = covs$land.num,                       # Landscape random effect (as.num)
                 nsource = length(unique(covs$source)),      # number of levels in source RE 
                 source = covs$source.num,                   # Data source random effect (as.num)
                 nyear = length(unique(covs$year.num)),      # number of levels in year RE
                 year = covs$year.num)                       # year random effect (as.num)
    
    ## save the bundle 
    temp[[l]] = bdata
    names(temp)[l] = paste("SUB-", sp_sub, "~DOM-",sp_dom, sep = "" )
    
  } # end 2nd species
  
  if(length(temp) > 0){
    ## remove null values in list for species combos that dont overlap 
    temp[sapply(temp, is.null)] <- NULL
    
    ## save all results for all species
    counterf1_bdata_list[[i]] = temp
  }else{
    next
  }
  
} # end 1st species
rm(i,l,sp_dom,sp_sub, sp_dom_dat, sp_sub_dat, obs, temp, z.dom, z.sub,
   bdata, all_else, data.dom, data.sub, covs, exp, non_det_SU, 
   non_det_land, lands, pres, t, su)

## flatten the list of lists into a single list
counterf1_bdata_list = unlist(counterf1_bdata_list, recursive = FALSE)
length(counterf1_bdata_list) #46 combos of preferred pred and prey after removing species that have less than 100 detections in shared SPECIAL landscapes 
sort(names(counterf1_bdata_list)) # looks good!
## not a lot of prey options tho! 
## how many bottom-up tests
bu = names(counterf1_bdata_list)[grepl("SUB-Panthera_tigis|SUB-Panthera_pardus|SUB-Neofelis_genus|SUB-Cuon_alpinus", names(counterf1_bdata_list))]
length(bu) # 33 bottom-up 
length(names(counterf1_bdata_list)[!names(counterf1_bdata_list) %in% bu]) # 13 top-down 


```

```{r Save bundled data for counterfactual test 1}

## grab today's date 
day<-str_sub(Sys.Date(),-2)
month<-str_sub(Sys.Date(),-5,-4)
year<-str_sub(Sys.Date(),-10,-7)
date = paste(year,month,day, sep = "")
rm(day,month,year)
 
# Save data 
# saveRDS(counterf1_bdata_list, paste("data_GitHub_CoA_bundles/counterfactual_testing/Bundled_data_for_Bayes_co-abundance_mods_counterfactual1_isolate_abundance_", length(counterf1_bdata_list), "_species_pairs_", date, ".RDS", sep = ""))


```

```{r Ultra loop to generate data bundles for counterfactual test 2}

## this one is much more simple, only select landscapes where both species are present! 
## therefore, iZIP parameter will all be one and essentially pointless in the model 

## store results here
counterf2_bdata_list = list()

for(i in 1:length(umfs)){ # repeat for each first speices
  
  ## select a single species and thier data 
  # 1st species is the dominant! 
  sp_dom = names(umfs)[i]
  sp_dom_dat = umfs[[i]]
  
  ### Thin all_else based upon predator or not
  if(!sp_dom %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
    
    # if its not a predator, then make all_else the predators
    all_else = c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")
    
  } # end non-predator condition
  
  ## if the dominant species is a predator 
  if(sp_dom %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
    
    ## then sub_species should only be preferred prey
    if(sp_dom == "Cuon_alpinus"){ 
      all_else = traits$scientificNameStd[traits$dhole_pref == "Yes"]
    }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Panthera_tigris"){ 
      all_else = traits$scientificNameStd[traits$tiger_pref == "Yes"]
      }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Panthera_pardus"){ 
      all_else = traits$scientificNameStd[traits$leopard_pref == "Yes"]
      }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Neofelis_genus"){ 
      all_else = traits$scientificNameStd[traits$CL_pref == "Yes"]
      }
    
  } # end dominant predator condition
  
  temp = list() # store results per sp1 here 
  for(l in 1:length(all_else)){ # repeat for each second species
    
    ## select a second species and their data
    # 2nd species is the subordinate! 
    sp_sub = all_else[l]
    sp_sub_dat = umfs[[sp_sub]]
    
    ## but verify if this is a preferred prey of one of these predators! 
    if(sp_sub %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
      
      ## then sub_species should only be preferred prey
      if(sp_sub == "Cuon_alpinus" & sp_dom %in% traits$scientificNameStd[traits$dhole_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Panthera_tigris" & sp_dom %in% traits$scientificNameStd[traits$tiger_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Panthera_pardus" & sp_dom %in% traits$scientificNameStd[traits$leopard_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Neofelis_genus" & sp_dom %in% traits$scientificNameStd[traits$CL_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      
    } # end preferred prey conditional 
    
    ##### Add a conditional to verify there are a combined total of 100 detections
    ##### shared across the landscapes where BOTH species were detected. 
    
    ## Grab landscapes present for each species
    dom_land = unique(caps$Landscape[caps$Species == sp_dom])
    sub_land = unique(caps$Landscape[caps$Species == sp_sub])
    ## and only save the ones where they intersect 
    lands = intersect(dom_land, sub_land)
    # lands = unique(dom_land, sub_land)
    rm(dom_land, sub_land)
    
    ## gather number of detections at shared landscapes for both species
    dom_det = nrow(caps[caps$Species == sp_dom &
                          caps$Landscape %in% lands,])
    sub_det = nrow(caps[caps$Species == sp_sub &
                          caps$Landscape %in% lands,])
    
    ## if there are less than 100 detections of both species from shared landscapes, 
    if(dom_det + sub_det < 100){
      # let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have less than 100 detections in shared landscapes! This pair was skipped."))
      # and skip the probelmatic pair
      next
      
    } # end 100 detection conditional 
    rm(dom_det, sub_det)
    ## this conditional will also catch species pairs that dont spatially overlap, 
    ## where a conditonal does that below and isnt needed anymore, but leaving for saftey. 
    
    #
    ##
    ###
    #### Thin data to landscapes where both species were detected and convert to matrix
    ###
    ##
    #
    
    ## grab all sampling units from shared landscapes
    su = meta$cell_id[meta$Landscape %in% lands]
    
    ## extract both count matricies
    data.dom<- sp_dom_dat$y #dominant
    data.sub<- sp_sub_dat$y #subordinate
    
    ## and thin both to present SUs
    data.dom = data.dom[data.dom$SU %in% su,]
    data.sub = data.sub[data.sub$SU %in% su,]
    
    ## make sure we match! 
    if(nrow(data.dom) != nrow(data.sub)){
      
      # if we dont, let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have different sized matricies! This pair was skipped."))
      # and skip the probelmatic one
      next
      
    } # end exact matrix size validation
    
    ## convert matrix dataframes back to matrix 
    rownames(data.dom)=data.dom[,1] #set rownames as sampling units
    data.dom=data.dom[,-1] #remove sampling unit col
    data.dom=as.matrix(data.dom) #turn into a matrix
    # repeat for spp 2
    rownames(data.sub)=data.sub[,1]
    data.sub=data.sub[,-1]
    data.sub=as.matrix(data.sub)
    
    #
    ##
    ###
    #### Site-level covairtes
    ###
    ##
    #

    # Make sure we have the same number of sampling units
    if(dim(sp_dom_dat$siteCovs)[1] != dim(sp_sub_dat$siteCovs)[1]){
      
       # if we dont, let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have a different number of sampling units! This pair was skipped."))
      # and skip the probelmatic one
      next
     
    }else{
      
      # thin to match relevant sampling units
      covs = sp_dom_dat$siteCovs[sp_dom_dat$siteCovs$cell_id %in% rownames(data.dom),]
      
    } # end cov size conditional 

    
    ### Add conditional if no sites match each other
    if(dim(covs)[1] == 0){
      print(paste0(sp_dom, " & ", sp_sub, " dont spatially overlap, no combo created."))
      next
    }# end lack of spatial overlap bypass  
    
    
    #Extract Landscape for random effect, stored as a number
    d = data.frame("Landscape"= sort(unique(covs$Landscape)),
                   "land.num" = seq(from = 1, to = length(unique(covs$Landscape))))
    covs = merge(covs, d, by = "Landscape")

    #Extract years for random effect, stored as a number
    d = data.frame("year"= sort(unique(covs$year)),
                   "year.num" = seq(from = 1, to = length(unique(covs$year))))
    covs = merge(covs, d, by = "year")

    #Extract data source for random effect, stored as a number
    d = data.frame("source"= sort(unique(covs$source)),
                   "source.num" = seq(from = 1, to = length(unique(covs$source))))
    covs = merge(covs, d, by = "source")
    rm(d)
    
    
    ## Make sure covs match order of matrix
    covs = covs[order(match(covs$cell_id, rownames(data.dom))),]
    
    #verify
    # match(covs$cell_id, rownames(data.dom))
    
    #
    ##
    ###
    #### Observation-level covariates
    ###
    ##
    #
    
    # Make sure we have the same sampling units!
    if(length(setdiff(sp_dom_dat$ObsCovs$SU, sp_sub_dat$ObsCovs$SU)) == 0 &
       length(setdiff(sp_sub_dat$ObsCovs$SU, sp_dom_dat$ObsCovs$SU)) == 0){
      
      ## thin obs down to relevant SUs
      obs = sp_dom_dat$ObsCovs[sp_dom_dat$ObsCovs$SU %in% rownames(data.dom),]
      
      ## and convert to a matrix 
      rownames(obs)=obs[,1] #set rownames as sampling units
      obs=obs[,-1] #remove sampling unit col
      obs=as.matrix(obs) #turn into a matrix, but only keep one 
      
    }else{
      
      ## if obs SUs dont match, let us know! 
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                "have a different number of sampling units in ObsCovs! This pair was skipped."))
      # and skip the probelmatic one
      next
    }
    
    ## Ensure obs matches the order of matrix 
    obs = obs[order(match(rownames(obs), rownames(data.dom))),]
    # #verify
    # match(rownames(obs), rownames(data.dom))
    
    #
    ##
    ###
    ####
    ##### Prepare the informed ZIP parameters
    ####
    ###
    ##
    #
    ### But there should be no need, since we've already subetted for relevant landscapes 
    
    ## Establish which sites have extirpated dominant species
    
    # # first get max number of counts per row
    # p = apply(data.dom, 1, max, na.rm = TRUE)
    # 
    # # subset for only sampling units with detections
    # p = names(p[p >= 1])
    # 
    # ### Add conditional if no sites match each other
    # if(length(p) == 0){
    #   print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
    #   next
    # }# end lack of spatial overlap bypass  
    # 
    # # gather all sampling units in the Landscapes where there were detections
    # c = select(covs, cell_id, Landscape)
    # d = c[c$cell_id %in% p,]
    # d = c[c$Landscape %in% d$Landscape, ]
    # 
    # # gather all sites in landscapes without ANY detections
    # e = c[!c$Landscape %in% d$Landscape,]
    # 
    # # this should match number of rows in matrix 
    # if((length(e$cell_id) + length(d$cell_id)) != dim(data.dom)[1] &
    #   (length(e$cell_id) + length(d$cell_id)) != dim(data.sub)[1]){
    #   
    #   print(paste0("problem with ", sp_dom, " & ", sp_sub))
    #   
    # } # end data check 
    
    # ## Add if-else statment for dominant species not extirpated anywhere. 
    # if(length(e$cell_id) > 0){
    #   
    #   # combine, where 0 = extirpated, and 1 = present. 
    #   z.dom = rbind(data.frame("occu" = 0, "cell_id" = e$cell_id),
    #                 data.frame("occu" = 1, "cell_id" = d$cell_id))
    #   
    # }else{
      
    z.dom = data.frame("occu" = 1, "cell_id" = covs$cell_id)
      
    # } # end no extirpation conditional 
    
    # make sure its in the same order as the matrix
    z.dom = z.dom[order(match(z.dom$cell_id, rownames(data.dom))),]
    
    
    ## Establish which sites have extirpated subordianate species-
    
    # # first get max number of counts per row
    # p = apply(data.sub, 1, max, na.rm = TRUE)
    # 
    # # subset for only sites with detections
    # p = names(p[p >= 1])
    # 
    # ### Add conditional if no sites match each other
    # if(length(p) == 0){
    #   print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
    #   next
    # }# end lack of spatial overlap bypass  
    # 
    # # gather all sites in the landscapes where there were detections
    # c = select(covs, cell_id, Landscape)
    # d = c[c$cell_id %in% p,]
    # d = c[c$Landscape %in% d$Landscape, ]
    # 
    # # gather all sites in landscapes without ANY detections
    # e = c[!c$Landscape %in% d$Landscape,]
    # 
    # # this should match number of rows in matrix 
    # if((length(e$cell_id) + length(d$cell_id)) != dim(data.dom)[1] &
    #    (length(e$cell_id) + length(d$cell_id)) != dim(data.sub)[1]){
    #   
    #   print(paste0("problem with ", sp1, " & ", sp2))
    #   
    # } # end data check 
    # 
    # # combine, where 0 = extirpated, and 1 = present. 
    # ## Add if-else statment for subordinate species not extirpated anywhere. 
    # if(length(e$cell_id) > 0){
    #   
    #   z.sub = rbind(data.frame("occu" = 0, "cell_id" = e$cell_id),
    #                 data.frame("occu" = 1, "cell_id" = d$cell_id))
    #   
    # }else{
      
    z.sub = data.frame("occu" = 1, "cell_id" = covs$cell_id)
    # } # end no extirpation conditional 
    
    # make sure its in the same order as the matrix
    z.sub = z.sub[order(match(z.sub$cell_id, rownames(data.sub))),]
    
    # rm(c,d,e,p)
    
    ####### BUNDLE all the data for the bayesian model 
    
    #these are the variable names used in the bayes mod! 
    bdata = list(y.dom = data.dom,                           # Count history matrix for dominant spp
                 y.sub = data.sub,                           # Count history matrix for subordinate spp
                 Z.dom = z.dom$occu,                         # Dominant presence/absence for all sites
                 Z.sub = z.sub$occu,                         # Subordinate presence/absence for all sites 
                 nsites = dim(data.dom)[1],                  # Number of sampling locations (i.e. rows in matrix)
                 nreps = dim(data.dom)[2],                   # Number of sampling occasions (i.e. cols in matrix)
                 # flii = covs$Avg_FLLI_3km,                   # Site covaraite 1
                 # hfp = covs$Avg_human_footprint_3km,         # Site covaraite 2
                 # elev = covs$Avg_altitude_3km,               # Site covaraite 3
                 flii = covs$Avg_FLLI_5km,                   # Site covaraite 1 @ 5km 
                 hfp = covs$Avg_human_footprint_5km,         # Site covaraite 2 @ 5km 
                 elev = covs$Avg_altitude_5km,               # Site covaraite 3 @ 5km 
                 cams = obs,                                 # Observation covaraite 1
                 narea = length(unique(covs$land.num)),      # number of levels in landscape RE
                 area = covs$land.num,                       # Landscape random effect (as.num)
                 nsource = length(unique(covs$source)),      # number of levels in source RE 
                 source = covs$source.num,                   # Data source random effect (as.num)
                 nyear = length(unique(covs$year.num)),      # number of levels in year RE
                 year = covs$year.num)                       # year random effect (as.num)
    
    ## save the bundle 
    temp[[l]] = bdata
    names(temp)[l] = paste("SUB-", sp_sub, "~DOM-",sp_dom, sep = "" )
    
  } # end 2nd species
  
  if(length(temp) > 0){
    
    ## remove null values in list for species combos that dont overlap 
    temp[sapply(temp, is.null)] <- NULL
    
    ## save all results for all species
    counterf2_bdata_list[[i]] = temp
    
  }else{
    next
    
  } # end empty temp condition
  
} # end 1st species
rm(i,l,sp_dom,sp_sub, sp_dom_dat, sp_sub_dat, obs, temp, z.dom, z.sub,
   bdata, all_else, data.dom, data.sub, covs, su,lands)

## flatten the list of lists into a single list
counterf2_bdata_list = unlist(counterf2_bdata_list, recursive = FALSE)
length(counterf2_bdata_list) #66 combos of preferred pred and prey --> perfect match! 
sort(names(counterf2_bdata_list)) # looks correct! 

## verify iZIP is 1 for everyone
for(i in 1:length(counterf2_bdata_list)){
  # grab one bundle 
  c = counterf2_bdata_list[[i]]
  # check iZIP 
  if(unique(c$Z.dom) == 1 & unique(c$Z.sub) == 1){
    next
  }else{
    #let us know if it doesnt match! 
    print(names(counterf2_bdata_list)[i], "has iZIP values of zero when its supposed to be one. Inspect!! ")
  }
}
## no problems! 

## how many dhole dom mods?
names(counterf2_bdata_list)[grepl("DOM-Cuon_alpinus", names(counterf2_bdata_list))] # 5 
## how many sub?
names(counterf2_bdata_list)[grepl("SUB-Cuon_alpinus", names(counterf2_bdata_list))] # Also 5, good! 


## how many bottom-up tests
bu = names(counterf2_bdata_list)[grepl("SUB-Panthera_tigis|SUB-Panthera_pardus|SUB-Neofelis_genus|SUB-Cuon_alpinus", names(counterf2_bdata_list))]
length(bu) # 24 bottom-up 
length(names(counterf2_bdata_list)[!names(counterf2_bdata_list) %in% bu]) # 42 top-down 
## not 33 and 33??
rm(bu)

```

```{r Save bundled data for counterfactual test 2}


## grab today's date 
day<-str_sub(Sys.Date(),-2)
month<-str_sub(Sys.Date(),-5,-4)
year<-str_sub(Sys.Date(),-10,-7)
date = paste(year,month,day, sep = "")
rm(day,month,year)
 
# Save data 
# saveRDS(counterf2_bdata_list, paste("data_GitHub_CoA_bundles/counterfactual_testing/Bundled_data_for_Bayes_co-abundance_mods_counterfactual2_isolate_extirpation_", length(counterf2_bdata_list), "_species_pairs_", date, ".RDS", sep = ""))

```

Ideally, there would be a chunk here that implements some nifty site-matching formulas or at least examines various covariate relationships. However, I am running out of time atm and just want to get a test running. Therefore, I will examine relationships ONLY at the Thai E Forest complex and the three landscapes it is composed of. 

```{r Ultra loop to generate data bundles for counterfactual test 3}

## only examining 3 connected landscapes here! 
# "Thailand_Thai_E_FC_Center","Thailand_Thai_E_FC_East","Thailand_Thai_E_FC_West"     

#create a list to store results
counterf3_bdata_list = list()

for(i in 1:length(umfs)){ # repeat for each first speices
  
  ## select a single species and thier data 
  # 1st species is the dominant! 
  sp_dom = names(umfs)[i]
  sp_dom_dat = umfs[[i]]
  
  ### Thin all_else based upon predator or not
  if(!sp_dom %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
    
    # if its not a predator, then make all_else the predators
    all_else = c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")
    
  } # end non-predator condition
  
  ## if the dominant species is a predator 
  if(sp_dom %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
    
    ## then sub_species should only be preferred prey
    if(sp_dom == "Cuon_alpinus"){ 
      all_else = traits$scientificNameStd[traits$dhole_pref == "Yes"]
    }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Panthera_tigris"){ 
      all_else = traits$scientificNameStd[traits$tiger_pref == "Yes"]
      }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Panthera_pardus"){ 
      all_else = traits$scientificNameStd[traits$leopard_pref == "Yes"]
      }
    ## then sub_species should only be preferred prey
    if(sp_dom == "Neofelis_genus"){ 
      all_else = traits$scientificNameStd[traits$CL_pref == "Yes"]
      }
    
  } # end dominant predator condition
  
  temp = list() # store results per sp1 here 
  for(l in 1:length(all_else)){ # repeat for each second species
    
    ## select a second species and their data
    # 2nd species is the subordinate! 
    sp_sub = all_else[l]
    sp_sub_dat = umfs[[sp_sub]]
    
     ## but verify if this is a preferred prey of one of these predators! 
    if(sp_sub %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
      
      ## then sub_species should only be preferred prey
      if(sp_sub == "Cuon_alpinus" & sp_dom %in% traits$scientificNameStd[traits$dhole_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Panthera_tigris" & sp_dom %in% traits$scientificNameStd[traits$tiger_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Panthera_pardus" & sp_dom %in% traits$scientificNameStd[traits$leopard_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Neofelis_genus" & sp_dom %in% traits$scientificNameStd[traits$CL_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
    } # end preferred prey condition
      
    
    ##### Add a conditional to verify there are a combined total of 100 detections
    ##### shared across the landscapes where BOTH species were detected. 
    ## but this counterfactual test is limiting us to only three landscapes! 
    
    lands = c("Thailand_Thai_E_FC_Center","Thailand_Thai_E_FC_East","Thailand_Thai_E_FC_West" )

    
    ## gather number of detections at shared landscapes for both species
    dom_det = nrow(caps[caps$Species == sp_dom &
                          caps$Landscape %in% lands,])
    sub_det = nrow(caps[caps$Species == sp_sub &
                          caps$Landscape %in% lands,])
    
    ## if there are less than 100 detections of both species from shared landscapes, 
    if(dom_det + sub_det < 100){
      # let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have less than 100 detections in shared landscapes! This pair was skipped."))
      # and skip the probelmatic pair
      next
      
    } # end 100 detection conditional 
    rm(dom_det, sub_det, lands)
    ## this conditional will also catch species pairs that dont spatially overlap, 
    ## where a conditonal does that below and isnt needed anymore, but leaving for saftey. 
    
    #
    ##
    ###
    #### Remove landscapes where neither species was detected and convert to matrix
    ###
    ##
    #
    
    ## First, determine which landscapes species are extirpated in
    exp = distinct(select(caps[caps$Species %in% c(sp_sub, sp_dom),], Landscape, Species))
    exp$presence = "yes"
    
    ## need to add landscapes where Species was NOT detected via loop 
    exp2 = list()
    # e = unique(exp$Species)[1]
    for(e in unique(exp$Species)){ # repeat for each species
      
      # select the relevant landscapes
      b = unique(caps$Landscape[caps$Species == e])
      #and make it a dataframe
      a = data.frame("Landscape" = unique(caps$Landscape[!caps$Landscape %in% b]))
      
      # save species name and make it a no
      a$Species = e
      a$presence = "no"
      
      exp2[[e]] = a
      
    }
    rm(a, b, e)
    
    # combine list into a df
    exp2 = do.call(rbind, exp2)
    rownames(exp2)= NULL
    
    # rbind presence w/ absences
    exp = rbind(exp, exp2)
    rm(exp2)
    
    ## Thin exp to only include relevant landscapes for this test! 
    exp = exp[exp$Landscape %in% c("Thailand_Thai_E_FC_Center","Thailand_Thai_E_FC_East","Thailand_Thai_E_FC_West"),]
    
    ## grab all landscapes where neither species was detected
    non_det_land <- exp %>%
      group_by(Landscape) %>%
      filter(all(presence == "no"))
    
    ## extract all the sampling unit names from these landscapes
    non_det_SU = meta$cell_id[meta$Landscape %in% unique(non_det_land$Landscape)]
    
    ## extract both count matricies
    data.dom<- sp_dom_dat$y #dominant
    data.sub<- sp_sub_dat$y #subordinate
    
    ## and thin both to remove empty SUs
    data.dom = data.dom[!data.dom$SU %in% non_det_SU,]
    data.sub = data.sub[!data.sub$SU %in% non_det_SU,]
    
    ## but also grab the relevant sampling units for our test! 
    su = meta$cell_id[meta$Landscape %in% c("Thailand_Thai_E_FC_Center","Thailand_Thai_E_FC_East","Thailand_Thai_E_FC_West")]
    
    ## and thin both to the relevant SUs
    data.dom = data.dom[data.dom$SU %in% su, ]
    data.sub = data.sub[data.sub$SU %in% su, ]
    
    if(is.null(data.dom)){
      ## let us know 
      # print("Dominant species:", sp_dom, "is not present in counterfactual3 landscapes, pairing with subordinate species:", sp_sub, "is getting skipped!")
      next
    }
    if(is.null(data.sub)){
      ## let us know 
      # print("Subordinate species:", sp_sub, "is not present in counterfactual3 landscapes, pairing with dominant species:", sp_dom, "is getting skipped!")
      next
    }

    
    ## make sure we match! 
    if(nrow(data.dom) != nrow(data.sub)){
      
      # if we dont, let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have different sized matricies! This pair was skipped."))
      # and skip the probelmatic one
      next
      
    } # end exact matrix size validation
    
    ### Now verify that both species are detected together in at least one landscape
    pres = unique(exp$Landscape[exp$presence == "yes" & exp$Species == unique(exp$Species)[1] &
                                  exp$Landscape %in% exp$Landscape[exp$presence == "yes" & 
                                                                     exp$Species ==  unique(exp$Species)[2]]])
    ## if there are no landscapes where both are detected,
    if(length(pres) == 0){
      
      ## let us know 
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "do not spatially overlap! This pair was skipped."))
      ## and skip the problematic one
      next
      
    }

    
    ## convert matrix dataframes back to matrix 
    rownames(data.dom)=data.dom[,1] #set rownames as sampling units
    data.dom=data.dom[,-1] #remove sampling unit col
    data.dom=as.matrix(data.dom) #turn into a matrix
    # repeat for spp 2
    rownames(data.sub)=data.sub[,1]
    data.sub=data.sub[,-1]
    data.sub=as.matrix(data.sub)
    
    #
    ##
    ###
    #### Site-level covairtes
    ###
    ##
    #

    # Make sure we have the same number of sampling units
    if(dim(sp_dom_dat$siteCovs)[1] != dim(sp_sub_dat$siteCovs)[1]){
      
       # if we dont, let us know
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                  "have a different number of sampling units! This pair was skipped."))
      # and skip the probelmatic one
      next
     
    }else{
      
      # thin to match relevant sampling units
      covs = sp_dom_dat$siteCovs[sp_dom_dat$siteCovs$cell_id %in% rownames(data.dom),]
      
    } # end cov size conditional 

    
    ### Add conditional if no sites match each other
    if(dim(covs)[1] == 0){
      print(paste0(sp_dom, " & ", sp_sub, " dont spatially overlap, no combo created."))
      next
    }# end lack of spatial overlap bypass  
    
    
    #Extract Landscape for random effect, stored as a number
    d = data.frame("Landscape"= sort(unique(covs$Landscape)),
                   "land.num" = seq(from = 1, to = length(unique(covs$Landscape))))
    covs = merge(covs, d, by = "Landscape")

    #Extract years for random effect, stored as a number
    d = data.frame("year"= sort(unique(covs$year)),
                   "year.num" = seq(from = 1, to = length(unique(covs$year))))
    covs = merge(covs, d, by = "year")

     #Extract data source for random effect, stored as a number
    d = data.frame("source"= sort(unique(covs$source)),
                   "source.num" = seq(from = 1, to = length(unique(covs$source))))
    covs = merge(covs, d, by = "source")
    rm(d)
    
    
    ## Make sure covs match order of matrix
    covs = covs[order(match(covs$cell_id, rownames(data.dom))),]
    
    #verify
    # match(covs$cell_id, rownames(data.dom))
    
    #
    ##
    ###
    #### Observation-level covariates
    ###
    ##
    #
    
    # Make sure we have the same sampling units!
    if(length(setdiff(sp_dom_dat$ObsCovs$SU, sp_sub_dat$ObsCovs$SU)) == 0 &
       length(setdiff(sp_sub_dat$ObsCovs$SU, sp_dom_dat$ObsCovs$SU)) == 0){
      
      ## thin obs down to relevant SUs
      obs = sp_dom_dat$ObsCovs[sp_dom_dat$ObsCovs$SU %in% rownames(data.dom),]
      
      ## and convert to a matrix 
      rownames(obs)=obs[,1] #set rownames as sampling units
      obs=obs[,-1] #remove sampling unit col
      obs=as.matrix(obs) #turn into a matrix, but only keep one 
      
    }else{
      
      ## if obs SUs dont match, let us know! 
      print(paste("The species combo of", sp_sub, "~", sp_dom, 
                "have a different number of sampling units in ObsCovs! This pair was skipped."))
      # and skip the probelmatic one
      next
    }
    
    ## Ensure obs matches the order of matrix 
    obs = obs[order(match(rownames(obs), rownames(data.dom))),]
    # #verify
    # match(rownames(obs), rownames(data.dom))
    
    #
    ##
    ###
    ####
    ##### Prepare the informed ZIP parameters
    ####
    ###
    ##
    #
    
    ## Establish which sites have extirpated dominant species
    
    # first get max number of counts per row
    p.dom = apply(data.dom, 1, max, na.rm = TRUE)

    # subset for only sampling units with detections
    p.dom = names(p.dom[p.dom >= 1])
    
    ### Add conditional if no sites match each other
    if(length(p.dom) == 0){
      print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
      next
    }# end lack of spatial overlap bypass  
    
    # gather all sampling units in the Landscapes where species was detected
    d.dom = select(covs[covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.dom]), ], 
                   cell_id, Landscape) # only need cell_id and landscape
    
    # gather all sites in landscapes without ANY detections
    e.dom = select(covs[! covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.dom]), ], 
                   cell_id, Landscape)
    
    # this should match number of rows in matrix 
    if((length(e.dom$cell_id) + length(d.dom$cell_id)) != dim(data.dom)[1] &
      (length(e.dom$cell_id) + length(d.dom$cell_id)) != dim(data.sub)[1]){
      
      print(paste0("problem with ", sp_dom, " & ", sp_sub))
      
    } # end data check 
    
    
    ## Add if-else statement for dominant species not extirpated anywhere. 
    if(length(e.dom$cell_id) > 0){
      
      # combine, where 0 = extirpated, and 1 = present. 
      z.dom = rbind(data.frame("occu" = 0, "cell_id" = e.dom$cell_id),
                    data.frame("occu" = 1, "cell_id" = d.dom$cell_id))
      
    }else{
      
      z.dom = data.frame("occu" = 1, "cell_id" = d.dom$cell_id)
      
    } # end no extirpation conditional 
    
    # make sure its in the same order as the matrix
    z.dom = z.dom[order(match(z.dom$cell_id, rownames(data.dom))),]
    

    ## Establish which sites have extirpated subordianate species-
    
    # first get max number of counts per row
    p.sub = apply(data.sub, 1, max, na.rm = TRUE)
    
    # subset for only sites with detections
    p.sub = names(p.sub[p.sub >= 1])
    
    ### Add conditional if no sites match each other
    if(length(p.sub) == 0){
      print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
      next
    }# end lack of spatial overlap bypass  
    
    # gather all sampling units in the Landscapes where species was detected
    d.sub = select(covs[covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.sub]), ], 
                   cell_id, Landscape) # only need cell_id and landscape
    
    # gather all sites in landscapes without ANY detections
    e.sub = select(covs[! covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.sub]), ], 
                   cell_id, Landscape)
    
    
    # this should match number of rows in matrix 
    if((length(e.sub$cell_id) + length(d.sub$cell_id)) != dim(data.dom)[1] &
       (length(e.sub$cell_id) + length(d.sub$cell_id)) != dim(data.sub)[1]){
      
      print(paste0("problem with ", sp1, " & ", sp2))
      
    } # end data check 
   
    # combine, where 0 = extirpated, and 1 = present. 
    ## Add if-else statement for subordinate species not extirpated anywhere. 
    if(length(e.sub$cell_id) > 0){
      
      z.sub = rbind(data.frame("occu" = 0, "cell_id" = e.sub$cell_id),
                    data.frame("occu" = 1, "cell_id" = d.sub$cell_id))
      
    }else{
      
      z.sub = data.frame("occu" = 1, "cell_id" = d.sub$cell_id)
    } # end no extirpation conditional 
    
    # make sure its in the same order as the matrix
    z.sub = z.sub[order(match(z.sub$cell_id, rownames(data.sub))),]
    
    ## verify z.dom and z.sub have the same number of sampling units
    if(nrow(z.sub) != nrow(z.dom)){
      print(paste0(sp1, " & ", sp2, " have differing rows in their iZIP parameter! Inspect here!"))
      
      ## and end the loop, this is critical! 
      break
    }
    
    ####### BUNDLE all the data for the bayesian model 
    
    #these are the variable names used in the bayes mod! 
    bdata = list(y.dom = data.dom,                           # Count history matrix for dominant spp
                 y.sub = data.sub,                           # Count history matrix for subordinate spp
                 Z.dom = z.dom$occu,                         # Dominant presence/absence for all sites
                 Z.sub = z.sub$occu,                         # Subordinate presence/absence for all sites 
                 nsites = dim(data.dom)[1],                  # Number of sampling locations (i.e. rows in matrix)
                 nreps = dim(data.dom)[2],                   # Number of sampling occasions (i.e. cols in matrix)
                 # flii = covs$Avg_FLLI_3km,                   # Site covaraite 1
                 # hfp = covs$Avg_human_footprint_3km,         # Site covaraite 2
                 # elev = covs$Avg_altitude_3km,               # Site covaraite 3
                 flii = covs$Avg_FLLI_5km,                   # Site covaraite 1 @ 5km
                 hfp = covs$Avg_human_footprint_5km,         # Site covaraite 2 @ 5km 
                 elev = covs$Avg_altitude_5km,               # Site covaraite 3 @ 5km 
                 cams = obs,                                 # Observation covaraite 1
                 narea = length(unique(covs$land.num)),      # number of levels in landscape RE
                 area = covs$land.num,                       # Landscape random effect (as.num)
                 nsource = length(unique(covs$source)),      # number of levels in source RE 
                 source = covs$source.num,                   # Data source random effect (as.num)
                 nyear = length(unique(covs$year.num)),      # number of levels in year RE
                 year = covs$year.num)                       # year random effect (as.num)
    
    ## save the bundle 
    temp[[l]] = bdata
    names(temp)[l] = paste("SUB-", sp_sub, "~DOM-",sp_dom, sep = "" )
    
  } # end 2nd species
  
  if(length(temp) > 0){
  
  ## remove null values in list for species combos that dont overlap 
  temp[sapply(temp, is.null)] <- NULL
  
  ## save all results for all species
  counterf3_bdata_list[[i]] = temp  
  } # end temp length condition
  
} # end 1st species
rm(i,l,sp_dom,sp_sub, sp_dom_dat, sp_sub_dat, obs, temp, z.dom, z.sub,
   bdata, all_else, data.dom, data.sub, covs, exp, non_det_SU, 
   non_det_land, pres)


## flatten the list of lists into a single list
counterf3_bdata_list = unlist(counterf3_bdata_list, recursive = FALSE)
length(counterf3_bdata_list) #40 combos! thats so few! 
names(counterf3_bdata_list) # Wont get to test leopards here, but that is ok. 
unique(counterf3_bdata_list$`SUB-Muntiacus_genus~DOM-Panthera_tigris`$Z.dom) # tigers present and absent! Very good! 


```

```{r Save bundled data for counterfactual test 3}

## grab today's date 
day<-str_sub(Sys.Date(),-2)
month<-str_sub(Sys.Date(),-5,-4)
year<-str_sub(Sys.Date(),-10,-7)
date = paste(year,month,day, sep = "")
rm(day,month,year)
 
# # Save data 
# saveRDS(counterf3_bdata_list, paste("data_GitHub_CoA_bundles/counterfactual_testing/Bundled_data_for_Bayes_co-abundance_mods_counterfactual3_site_matching_", length(counterf3_bdata_list), "_species_pairs_", date, ".RDS", sep = ""))

```

```{r Ultra loop to generate data bundles for counterfactual test 4:6}

## first save the three covairates of interest that we will be varying/holding constant
vars = c("Avg_altitude_5km","Avg_FLLI_5km","Avg_human_footprint_5km")

## also calculate landscape averages for these variables 
land_avg = ddply(meta[, c("Landscape", vars)], .(Landscape), summarize, 
                 Avg_altitude_5km = mean(Avg_altitude_5km),
                 Avg_FLLI_5km = mean(Avg_FLLI_5km),
                 Avg_human_footprint_5km = mean(Avg_human_footprint_5km))

## make a list to store all results
counterf_vars_bdata_list = list()

for(v in 1:length(vars)){
  
  # single out the variable 
  var = vars[v]
  # and examine the two other vars
  va = vars[vars != var]
  
  ## Grab the mean and SD of variable of interest 
  mean_var1 = mean(meta[, va[1]])
  mean_var2 = mean(meta[, va[2]])
  sd_var1 = sd(meta[, va[1]])
  sd_var2 = sd(meta[, va[2]])
  
  ## grab landscapes w/ average values for the first variables
  l1 = land_avg[, c("Landscape", va[1])]
  lands_var1 = l1$Landscape[l1[,va[1]] >= (mean_var1 - sd_var1) & l1[,va[1]] <= (mean_var1 + sd_var1)]
  
  ## and repeat for the second variable 
  l2 = land_avg[, c("Landscape", va[2])]
  lands_var2 = l1$Landscape[l2[,va[2]] >= (mean_var2 - sd_var2) & l2[,va[2]] <= (mean_var2 + sd_var2)]
  
  ## and only save the ones that intersect
  lands_var = intersect(lands_var1, lands_var2)
  
  ## but let me know if something went wrong! 
  if(length(lands_var) == 0){
    print(paste("There were no landscape combinations when controlling for", var, 
                "so we are stopping the loop now. Inspect!!"))
    break
  } # end data check condition. 
  
  temp_var = list() # store results per sp1 here
  for(i in 1:length(umfs)){ # repeat for each first species
    
    ## select a single species and thier data 
    # 1st species is the dominant! 
    sp_dom = names(umfs)[i]
    sp_dom_dat = umfs[[i]]
    
    ### Thin all_else based upon predator or not
    if(!sp_dom %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
      
      # if its not a predator, then make all_else the predators
      all_else = c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")
      
    } # end non-predator condition
    
    ## if the dominant species is a predator 
    if(sp_dom %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
      
      ## then sub_species should only be preferred prey
      if(sp_dom == "Cuon_alpinus"){ 
        all_else = traits$scientificNameStd[traits$dhole_pref == "Yes"]
      }
      ## then sub_species should only be preferred prey
      if(sp_dom == "Panthera_tigris"){ 
        all_else = traits$scientificNameStd[traits$tiger_pref == "Yes"]
      }
      ## then sub_species should only be preferred prey
      if(sp_dom == "Panthera_pardus"){ 
        all_else = traits$scientificNameStd[traits$leopard_pref == "Yes"]
      }
      ## then sub_species should only be preferred prey
      if(sp_dom == "Neofelis_genus"){ 
        all_else = traits$scientificNameStd[traits$CL_pref == "Yes"]
      }
      
    } # end dominant predator condition
    
    temp = list() # store results per sp1 here 
    for(l in 1:length(all_else)){ # repeat for each second species
      
      ## select a second species and their data
      # 2nd species is the subordinate! 
      sp_sub = all_else[l]
      sp_sub_dat = umfs[[sp_sub]]
      
       ## but verify if this is a preferred prey of one of these predators! 
    if(sp_sub %in% c("Panthera_tigris", "Panthera_pardus", "Neofelis_genus", "Cuon_alpinus")){
      
      ## then sub_species should only be preferred prey
      if(sp_sub == "Cuon_alpinus" & sp_dom %in% traits$scientificNameStd[traits$dhole_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Panthera_tigris" & sp_dom %in% traits$scientificNameStd[traits$tiger_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Panthera_pardus" & sp_dom %in% traits$scientificNameStd[traits$leopard_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      ## then sub_species should only be preferred prey
      if(sp_sub == "Neofelis_genus" & sp_dom %in% traits$scientificNameStd[traits$CL_pref != "Yes"]){ 
        # skip to the next iterations
        next
      }
      
    } # end preferred prey conditional 
      
      ## First grab relevant SUs from our averaged landscapes above
      su = meta$cell_id[meta$Landscape %in% lands_var]
      
      # then thin
      dom_y = sp_dom_dat$y;dom_obs = sp_dom_dat$ObsCovs;dom_site = sp_dom_dat$siteCovs
      dom_y = dom_y[dom_y$SU %in% su, ];dom_obs = dom_obs[dom_obs$SU %in% su,];dom_site = dom_site[dom_site$cell_id %in% su,]
      # and update
      sp_dom_dat$y = dom_y;sp_dom_dat$ObsCovs = dom_obs;sp_dom_dat$siteCovs = dom_site
      
      ### repeat for subordinate
      # thin first 
      sub_y = sp_sub_dat$y;sub_obs = sp_sub_dat$ObsCovs;sub_site = sp_sub_dat$siteCovs
      sub_y = sub_y[sub_y$SU %in% su, ];sub_obs = sub_obs[sub_obs$SU %in% su,];sub_site = sub_site[sub_site$cell_id %in% su,]
      # and update
      sp_sub_dat$y = sub_y;sp_sub_dat$ObsCovs = sub_obs;sp_sub_dat$siteCovs = sub_site
      rm(sub_y,sub_obs,sub_site, dom_y, dom_obs,dom_site)
      
      ##### Add a conditional to verify there are a combined total of 100 detections
      ##### shared across the landscapes where BOTH species were detected. 
      
      ## Grab landscapes present for each species
      dom_land = unique(caps$Landscape[caps$Species == sp_dom & caps$Landscape %in% lands_var])
      sub_land = unique(caps$Landscape[caps$Species == sp_sub & caps$Landscape %in% lands_var])
      ## and only save the ones where they intersect 
      lands = intersect(dom_land, sub_land)
      rm(dom_land, sub_land)
      
      ## gather number of detections at shared landscapes for both species
      dom_det = nrow(caps[caps$Species == sp_dom &
                            caps$Landscape %in% lands,])
      sub_det = nrow(caps[caps$Species == sp_sub &
                            caps$Landscape %in% lands,])
      
      ## if there are less than 100 detections of both species from shared landscapes, 
      if(dom_det + sub_det < 100){
        # let us know
        print(paste("The species combo of", sp_sub, "~", sp_dom, 
                    "have less than 100 detections in shared landscapes! This pair was skipped."))
        # and skip the probelmatic pair
        next
        
      } # end 100 detection conditional 
      rm(dom_det, sub_det)
      ## this conditional will also catch species pairs that dont spatially overlap, 
      ## where a conditonal does that below and isnt needed anymore, but leaving for saftey. 
      
      #
      ##
      ###
      #### Remove landscapes where neither species was detected and convert to matrix
      ###
      ##
      #
      
      ## First, determine which landscapes species are extirpated in
      exp = distinct(select(caps[caps$Species %in% c(sp_sub, sp_dom) & caps$Landscape %in% lands,], Landscape, Species))
      exp$presence = "yes"
      
      ## need to add landscapes where Species was NOT detected via loop 
      exp2 = list()
      # e = unique(exp$Species)[1]
      for(e in unique(exp$Species)){ # repeat for each species
        
        # select the relevant landscapes
        b = unique(caps$Landscape[caps$Species == e & caps$Landscape %in% lands])
        #and make it a dataframe
        a = data.frame("Landscape" = unique(caps$Landscape[!caps$Landscape %in% b & caps$Landscape %in% lands]))
        
        ## if there are no extirpations
        if(nrow(a) == 0){
          next
        }
        
        # save species name and make it a no
        a$Species = e
        a$presence = "no"
        
        exp2[[e]] = a
        
      }
      rm(a, b, e)
      
      # combine list into a df
      exp2 = do.call(rbind, exp2)
      rownames(exp2)= NULL
      
      # rbind presence w/ absences
      exp = rbind(exp, exp2)
      rm(exp2)
      
      ## grab all landscapes where neither species was detected
      non_det_land <- exp %>%
        group_by(Landscape) %>%
        filter(all(presence == "no"))
      
      ## extract all the sampling unit names from these landscapes
      non_det_SU = meta$cell_id[meta$Landscape %in% unique(non_det_land$Landscape)]
      
      ## extract both count matricies
      data.dom<- sp_dom_dat$y #dominant
      data.sub<- sp_sub_dat$y #subordinate
      
      ## and thin both to remove empty SUs
      data.dom = data.dom[!data.dom$SU %in% non_det_SU,]
      data.sub = data.sub[!data.sub$SU %in% non_det_SU,]
      
      ## make sure we match! 
      if(nrow(data.dom) != nrow(data.sub)){
        
        # if we dont, let us know
        print(paste("The species combo of", sp_sub, "~", sp_dom, 
                    "have different sized matricies! This pair was skipped."))
        # and skip the probelmatic one
        next
        
      } # end exact matrix size validation
      
      ### Now verify that both species are detected together in at least one landscape
      pres = unique(exp$Landscape[exp$presence == "yes" & exp$Species == unique(exp$Species)[1] &
                                    exp$Landscape %in% exp$Landscape[exp$presence == "yes" & 
                                                                       exp$Species ==  unique(exp$Species)[2]]])
      ## if there are no landscapes where both are detected,
      if(length(pres) == 0){
        
        ## let us know 
        print(paste("The species combo of", sp_sub, "~", sp_dom, 
                    "do not spatially overlap! This pair was skipped."))
        ## and skip the problematic one
        next
        
      }
      
      
      ## convert matrix dataframes back to matrix 
      rownames(data.dom)=data.dom[,1] #set rownames as sampling units
      data.dom=data.dom[,-1] #remove sampling unit col
      data.dom=as.matrix(data.dom) #turn into a matrix
      # repeat for spp 2
      rownames(data.sub)=data.sub[,1]
      data.sub=data.sub[,-1]
      data.sub=as.matrix(data.sub)
      
      #
      ##
      ###
      #### Site-level covairtes
      ###
      ##
      #
      
      # Make sure we have the same number of sampling units
      if(dim(sp_dom_dat$siteCovs)[1] != dim(sp_sub_dat$siteCovs)[1]){
        
        # if we dont, let us know
        print(paste("The species combo of", sp_sub, "~", sp_dom, 
                    "have a different number of sampling units! This pair was skipped."))
        # and skip the probelmatic one
        next
        
      }else{
        
        # thin to match relevant sampling units
        covs = sp_dom_dat$siteCovs[sp_dom_dat$siteCovs$cell_id %in% rownames(data.dom),]
        
      } # end cov size conditional 
      
      
      ### Add conditional if no sites match each other
      if(dim(covs)[1] == 0){
        print(paste0(sp_dom, " & ", sp_sub, " dont spatially overlap, no combo created."))
        next
      }# end lack of spatial overlap bypass  
      
      
      #Extract Landscape for random effect, stored as a number
      d = data.frame("Landscape"= sort(unique(covs$Landscape)),
                     "land.num" = seq(from = 1, to = length(unique(covs$Landscape))))
      covs = merge(covs, d, by = "Landscape")
      
      #Extract years for random effect, stored as a number
      d = data.frame("year"= sort(unique(covs$year)),
                     "year.num" = seq(from = 1, to = length(unique(covs$year))))
      covs = merge(covs, d, by = "year")
      
      #Extract data source for random effect, stored as a number
      d = data.frame("source"= sort(unique(covs$source)),
                     "source.num" = seq(from = 1, to = length(unique(covs$source))))
      covs = merge(covs, d, by = "source")
      rm(d)
      
      
      ## Make sure covs match order of matrix
      covs = covs[order(match(covs$cell_id, rownames(data.dom))),]
      
      #verify
      # match(covs$cell_id, rownames(data.dom))
      
      #
      ##
      ###
      #### Observation-level covariates
      ###
      ##
      #
      
      # Make sure we have the same sampling units!
      if(length(setdiff(sp_dom_dat$ObsCovs$SU, sp_sub_dat$ObsCovs$SU)) == 0 &
         length(setdiff(sp_sub_dat$ObsCovs$SU, sp_dom_dat$ObsCovs$SU)) == 0){
        
        ## thin obs down to relevant SUs
        obs = sp_dom_dat$ObsCovs[sp_dom_dat$ObsCovs$SU %in% rownames(data.dom),]
        
        ## and convert to a matrix 
        rownames(obs)=obs[,1] #set rownames as sampling units
        obs=obs[,-1] #remove sampling unit col
        obs=as.matrix(obs) #turn into a matrix, but only keep one 
        
      }else{
        
        ## if obs SUs dont match, let us know! 
        print(paste("The species combo of", sp_sub, "~", sp_dom, 
                    "have a different number of sampling units in ObsCovs! This pair was skipped."))
        # and skip the probelmatic one
        next
      }
      
      ## Ensure obs matches the order of matrix 
      obs = obs[order(match(rownames(obs), rownames(data.dom))),]
      # #verify
      # match(rownames(obs), rownames(data.dom))
      
      #
      ##
      ###
      ####
      ##### Prepare the informed ZIP parameters
      ####
      ###
      ##
      #
      
      ## Establish which sites have extirpated dominant species
      
      # first get max number of counts per row
      p.dom = apply(data.dom, 1, max, na.rm = TRUE)
      
      # subset for only sampling units with detections
      p.dom = names(p.dom[p.dom >= 1])
      
      ### Add conditional if no sites match each other
      if(length(p.dom) == 0){
        print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
        next
      }# end lack of spatial overlap bypass  
      
      # gather all sampling units in the Landscapes where species was detected
      d.dom = select(covs[covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.dom]), ], 
                     cell_id, Landscape) # only need cell_id and landscape
      
      # gather all sites in landscapes without ANY detections
      e.dom = select(covs[! covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.dom]), ], 
                     cell_id, Landscape)
      
      # this should match number of rows in matrix 
      if((length(e.dom$cell_id) + length(d.dom$cell_id)) != dim(data.dom)[1] &
         (length(e.dom$cell_id) + length(d.dom$cell_id)) != dim(data.sub)[1]){
        
        print(paste0("problem with ", sp_dom, " & ", sp_sub))
        
      } # end data check 
      
      
      ## Add if-else statement for dominant species not extirpated anywhere. 
      if(length(e.dom$cell_id) > 0){
        
        # combine, where 0 = extirpated, and 1 = present. 
        z.dom = rbind(data.frame("occu" = 0, "cell_id" = e.dom$cell_id),
                      data.frame("occu" = 1, "cell_id" = d.dom$cell_id))
        
      }else{
        
        z.dom = data.frame("occu" = 1, "cell_id" = d.dom$cell_id)
        
      } # end no extirpation conditional 
      
      # make sure its in the same order as the matrix
      z.dom = z.dom[order(match(z.dom$cell_id, rownames(data.dom))),]
      
      
      ## Establish which sites have extirpated subordianate species-
      
      # first get max number of counts per row
      p.sub = apply(data.sub, 1, max, na.rm = TRUE)
      
      # subset for only sites with detections
      p.sub = names(p.sub[p.sub >= 1])
      
      ### Add conditional if no sites match each other
      if(length(p.sub) == 0){
        print(paste0(sp1, " & ", sp2, " may spatially overlap, but were never detected in the same landscape"))
        next
      }# end lack of spatial overlap bypass  
      
      # gather all sampling units in the Landscapes where species was detected
      d.sub = select(covs[covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.sub]), ], 
                     cell_id, Landscape) # only need cell_id and landscape
      
      # gather all sites in landscapes without ANY detections
      e.sub = select(covs[! covs$Landscape %in% unique(meta$Landscape[meta$cell_id %in% p.sub]), ], 
                     cell_id, Landscape)
      
      
      # this should match number of rows in matrix 
      if((length(e.sub$cell_id) + length(d.sub$cell_id)) != dim(data.dom)[1] &
         (length(e.sub$cell_id) + length(d.sub$cell_id)) != dim(data.sub)[1]){
        
        print(paste0("problem with ", sp1, " & ", sp2))
        
      } # end data check 
      
      # combine, where 0 = extirpated, and 1 = present. 
      ## Add if-else statement for subordinate species not extirpated anywhere. 
      if(length(e.sub$cell_id) > 0){
        
        z.sub = rbind(data.frame("occu" = 0, "cell_id" = e.sub$cell_id),
                      data.frame("occu" = 1, "cell_id" = d.sub$cell_id))
        
      }else{
        
        z.sub = data.frame("occu" = 1, "cell_id" = d.sub$cell_id)
      } # end no extirpation conditional 
      
      # make sure its in the same order as the matrix
      z.sub = z.sub[order(match(z.sub$cell_id, rownames(data.sub))),]
      
      ## verify z.dom and z.sub have the same number of sampling units
      if(nrow(z.sub) != nrow(z.dom)){
        print(paste0(sp1, " & ", sp2, " have differing rows in their iZIP parameter! Inspect here!"))
        
        ## and end the loop, this is critical! 
        break
      }
      
      ####### BUNDLE all the data for the bayesian model 
      
      #these are the variable names used in the bayes mod! 
      bdata = list(y.dom = data.dom,                           # Count history matrix for dominant spp
                   y.sub = data.sub,                           # Count history matrix for subordinate spp
                   Z.dom = z.dom$occu,                         # Dominant presence/absence for all sites
                   Z.sub = z.sub$occu,                         # Subordinate presence/absence for all sites 
                   nsites = dim(data.dom)[1],                  # Number of sampling locations (i.e. rows in matrix)
                   nreps = dim(data.dom)[2],                   # Number of sampling occasions (i.e. cols in matrix)
                   flii = covs$Avg_FLLI_5km,                   # Site covaraite 1 @ 5km res! 
                   hfp = covs$Avg_human_footprint_5km,         # Site covaraite 2 @ 5km res!
                   elev = covs$Avg_altitude_5km,               # Site covaraite 3 @ 5km res!
                   cams = obs,                                 # Observation covaraite 1
                   narea = length(unique(covs$land.num)),      # number of levels in landscape RE
                   area = covs$land.num,                       # Landscape random effect (as.num)
                   nsource = length(unique(covs$source)),      # number of levels in source RE 
                   source = covs$source.num,                   # Data source random effect (as.num)
                   nyear = length(unique(covs$year.num)),      # number of levels in year RE
                   year = covs$year.num)                       # year random effect (as.num)
      
      ## save the bundle 
      temp[[l]] = bdata
      names(temp)[l] = paste("SUB-", sp_sub, "~DOM-",sp_dom, sep = "" )
      
    } # end 2nd species
    
    if(length(temp) > 0){
      ## remove null values in list for species combos that dont overlap 
      temp[sapply(temp, is.null)] <- NULL
      
      # ## flatten the list of lists into a single list
      # temp = unlist(temp, recursive = FALSE)
      
      ## save all results for sp1
      temp_var[[i]] = temp
      # # and save the name too!
      # names(temp_var)[v] = var
      
    }else{
      next
    }
    
  } # end per first species 
  
  ## flatten the list of lists into a single list
  temp_var = unlist(temp_var, recursive = FALSE)
  
  ## so we can save it in another list 
  counterf_vars_bdata_list[[v]] = temp_var
  names(counterf_vars_bdata_list)[v] = var
  
} # end per variable
rm(i,l,sp_dom,sp_sub, sp_dom_dat, sp_sub_dat, obs, temp, z.dom, z.sub,
   bdata, all_else, data.dom, data.sub, covs, exp, non_det_SU, 
   non_det_land, lands, pres, t, su, d.dom, d.sub, e.dom, e.sub, l1, l2, 
   land_avg, temp_var, lands_var, lands_var1, lands_var2, mean_var1, mean_var2, 
   p.dom, p.sub, sd_var1, sd_var2, v, va, var, vars)

## inspect! 
length(counterf_vars_bdata_list$Avg_altitude_5km) # 66 mods here
length(counterf_vars_bdata_list$Avg_FLLI_5km) # 66 mods here
length(counterf_vars_bdata_list$Avg_human_footprint_5km) # 66 mods here

## check one key species pair and make sure it has different number of landscapes per list
counterf_vars_bdata_list$Avg_altitude_5km$`SUB-Panthera_tigris~DOM-Rusa_unicolor`$narea #32 for altitude
counterf_vars_bdata_list$Avg_FLLI_5km$`SUB-Panthera_tigris~DOM-Rusa_unicolor`$narea # 23 for FLII
counterf_vars_bdata_list$Avg_human_footprint_5km$`SUB-Panthera_tigris~DOM-Rusa_unicolor`$narea #20 for HFP
## I think that checks out alright! 


```

```{r Save bundled data for coutnerfactual tests 4:6}

## grab today's date 
day<-str_sub(Sys.Date(),-2)
month<-str_sub(Sys.Date(),-5,-4)
year<-str_sub(Sys.Date(),-10,-7)
date = paste(year,month,day, sep = "")
rm(day,month,year)
 
# # Save altitude
# saveRDS(counterf_vars_bdata_list$Avg_altitude_5km, paste("data_GitHub_CoA_bundles/counterfactual_testing/Bundled_data_for_Bayes_co-abundance_mods_counterfactual4_isolate_altitude_", length(counterf_vars_bdata_list$Avg_altitude_5km), "_species_pairs_", date, ".RDS", sep = ""))
# 
# # Save FLII
# saveRDS(counterf_vars_bdata_list$Avg_FLLI_5km, paste("data_GitHub_CoA_bundles/counterfactual_testing/Bundled_data_for_Bayes_co-abundance_mods_counterfactual5_isolate_FLII_", length(counterf_vars_bdata_list$Avg_FLLI_5km), "_species_pairs_", date, ".RDS", sep = ""))
# 
# # Save HFP
# saveRDS(counterf_vars_bdata_list$Avg_human_footprint_5km, paste("data_GitHub_CoA_bundles/counterfactual_testing/Bundled_data_for_Bayes_co-abundance_mods_counterfactual6_isolate_HFP_", length(counterf_vars_bdata_list$Avg_human_footprint_5km), "_species_pairs_", date, ".RDS", sep = ""))


```



